---
title: "CoastalFires_Flux_Remineralization"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---

#Load necessary Libraries
```{r}
library(dplyr)
library(tidyverse)
```

#set working directory
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires")
```

#load dataset
```{r}
CombinedData <- readr::read_csv("CoastalFires_CombinedData.csv")
```
#Calculate flux of levoglucosan for each sample
```{r}
CombinedData$Q_L_per_s <- CombinedData$Q_m_sq_per_sec * 1000 #convert discharge values from cubic meters per second to liters per second to make it compatible with levoglucosan concentrations that are in micro grams per liter

CombinedData$Levo_Flux_ug_per_s <- CombinedData$levo_micro_gram_per_liter * CombinedData$Q_L_per_s #calculate levoglucosan flux for each sample. Units should be in micro grams per second

CombinedData$Levo_Flux_kg_per_yr <- CombinedData$Levo_Flux_ug_per_s *86400 * 365 / 1000 / 1000 / 1000 #convert levoglucosan flux units to kilograms per year
```

#Average annual levoglucosan flux for each watershed
```{r}
Levo_Flux_Average <- CombinedData %>%
  group_by(watershed) %>%
  summarise(
    mean_levo_flux_kg_per_yr = mean(Levo_Flux_kg_per_yr, na.rm = TRUE), # na.rm = TRUE ignores NA values and mean is the average
    sd_levo_flux_kg_per_yr = sd(Levo_Flux_kg_per_yr, na.rm = TRUE) #to determine the standard deviation associated with the average per watershed
  )

# View the average and standard deviation of annual levoglucosan fluxes for each watershed
print(Levo_Flux_Average)
```

#Calculate annual amount of levoglucosan leaving these 5 watersheds
```{r}
# Add the annual flux of levoglucosan from each watershed together
total_mean_levo_kg_per_yr <- sum(Levo_Flux_Average$mean_levo_flux_kg_per_yr)

# Propagate standard deviation for the summation of the annual levoglucosan flux from all watersheds
total_sd_levo_kg_per_yr <- sqrt(sum(Levo_Flux_Average$sd_levo_flux_kg_per_yr^2))

# Combine results into a new data frame
Annual_Levo_Flux <- data.frame(
  All_Watersheds_levo_kg_per_yr = total_mean_levo_kg_per_yr,
  All_Watersheds_levo_sd = total_sd_levo_kg_per_yr
)

# View the result
print(Annual_Levo_Flux)
```

#Calculate the annual amount of low-temperature PyC that could be leaving these 5 watersheds
```{r}
All_Watersheds_ltPyC_kg_per_yr <- Annual_Levo_Flux$All_Watersheds_levo_kg_per_yr *100 / 1.7 #converts levoglucosan to low-temperature PyC. This is based on the percentage of total DOC leached from woody char created at 250C that is levoglucosan. Essentially for woody char leachates, 1.7% of low-temperature PyC is levoglucosan. See Norwood et al (2013) for further details. 

All_Watersheds_ltPyC_sd <- Annual_Levo_Flux$All_Watersheds_levo_sd *100 / 1.7 #converts the standard deviation as well

# Combine results into a new data frame
Annual_ltPyC_Flux <- data.frame(
  All_Watersheds_ltPyC_kg_per_yr = All_Watersheds_ltPyC_kg_per_yr,
  All_Watersheds_ltPyC_sd = All_Watersheds_ltPyC_sd
)

print(Annual_ltPyC_Flux)
```

#Estimate the emissions of CO2 assuming 100% remineralization of levoglucosan and low-temp PyC
```{r}

```












