---
title: "CoastalFires_CumulativeExport"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---

Packages
```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(patchwork)
library(ggplot2)
```


Get Data
```{r}
#From CUAHSI Hydroshare
DischargeData <- readr::read_csv("https://www.hydroshare.org/resource/26f08ef7fe344fcb9d3da45160b697fe/data/contents/CoastalFires_DischargeData.csv")

#From GitHub
CombinedData <- readr::read_csv("data/CoastalFires_CombinedData.csv")
```

Exclude samples that were taken at a different location. 
```{r}
#Scott Creek flooded during the events of January 27-28th, 2021. Samples were collected at a downstream lagoon site. However, due to this the data collected does not accurately reflect that at the sample site. As such authors of the associated manuscript have excluded it from all data analysis and figure creation.
CombinedData <- CombinedData %>%
  filter(Location_Change != "Y")
```

Make sure dates are right
```{r}
CombinedData <- CombinedData %>%
  mutate(
    DateTime = as.POSIXct(Date, tz = "America/Los_Angeles") +
      if_else(is.na(Time),
              12*60*60,
              as.numeric(Time)),
    DateTime = format(DateTime, "%Y-%m-%d %H:%M")
  )
```

Get data for cumulative export
```{r}
ConcData <- CombinedData %>%
  select(Watershed, DateTime, Levo_micro_mol_C_per_L, stdev_Levo_micro_mol_C_per_L, B6CA_and_B5CA_micro_mol_C_per_L)

#Fix watershed names to match discharge data
ConcData <- ConcData %>%
  mutate(Watershed = recode(Watershed,
                            "Laguna Creek" = "LAG",
                            "San Lorenzo River" = "SAN",
                            "Pescadero Creek" = "PESC",
                            "Majors Creek" = "MAJ",
                            "Scott Creek" = "SCT"))
#Make sure date is correct format
ConcData <- ConcData %>%
  mutate(
    DateTime = as.POSIXct(DateTime,
                          format = "%Y-%m-%d %H:%M",
                          tz = "America/Los_Angeles")
  )
```

Convert concentrations
```{r}
#to get gC from uM-C, multiply by 0.012011
ConcData$Levo_mg_C_per_L <- ConcData$Levo_micro_mol_C_per_L * 0.012011
ConcData$BPCA_mg_C_per_L <- ConcData$B6CA_and_B5CA_micro_mol_C_per_L * 0.012011

#remove unneccessary columns
ConcData <- ConcData %>%
  select(-Levo_micro_mol_C_per_L,
         -B6CA_and_B5CA_micro_mol_C_per_L)
```

Get Discharge set
```{r}
SAN <- DischargeData %>%
  transmute(
    Watershed = "SAN",
    DateTime = parse_date_time(SAN_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = SAN_Flow_cms
  )

PESC <- DischargeData %>%
  transmute(
    Watershed = "PESC",
    DateTime = parse_date_time(PESC_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = PESC_Flow_cms
  )

LAG <- DischargeData %>%
  transmute(
    Watershed = "LAG",
    DateTime = parse_date_time(LAG_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = LAG_Flow_cms
  )

MAJ <- DischargeData %>%
  transmute(
    Watershed = "MAJ",
    DateTime = parse_date_time(MAJ_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = MAJ_Flow_cms
  )

SCT <- DischargeData %>%
  transmute(
    Watershed = "SCT",
    DateTime = parse_date_time(SCT_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = SCT_Flow_cms
  )

FlowData <- bind_rows(SAN, PESC, LAG, MAJ, SCT)
```

Remove NAs from FlowData
```{r}
FlowData <- FlowData %>%
  filter(!is.na(DateTime),
         !is.na(Flow))

#check it worked
FlowData %>%
  group_by(Watershed) %>%
  summarize(
    missing_dates = sum(is.na(DateTime)),
    start = min(DateTime),
    end = max(DateTime),
    n=n()
  )
```

Convert to daily discharge (highest resolution for Scott Creek)
```{r}
# DailyFlow <- FlowData %>%
#   mutate(Date = as.Date(DateTime)) %>%
#   filter(Date >= as.Date("2020-10-07"),
#          Date <= as.Date("2021-05-02")) %>%
#   group_by(Watershed, Date) %>%
#   summarize(
#     Q_mean = mean(Flow, na.rm=TRUE),
#     .groups="drop"
#   )

DailyFlow <- FlowData %>%
  mutate(Date = as.Date(DateTime)) %>%
  filter(Date >= as.Date("2020-10-07"),
         Date <= as.Date("2021-05-02")) %>%
  group_by(Watershed, Date) %>%
  summarize(
    Q_mean = mean(Flow, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Q_log = log10(Q_mean)
  )

#check
DailyFlow %>%
  count(Watershed)
```

Get Daily Concentrations
```{r}
DailyConc <- DailyFlow %>%
  group_by(Watershed) %>%
  group_modify(~{

    ws <- .y$Watershed

    samples <- ConcData %>%
      filter(Watershed == ws) %>%
      arrange(DateTime) %>%
      mutate(Date = as.Date(DateTime))

    if(nrow(samples) < 2){
      .x$Levo_interp <- NA
      .x$BPCA_interp <- NA
      return(.x)
    }

    .x$Levo_interp <-
      approx(
          x = as.numeric(samples$Date),
          y = samples$Levo_mg_C_per_L,
          xout = as.numeric(.x$Date),
          method = "linear",
          rule = 1 #only interpolate within sampling period
      )$y 

    .x$BPCA_interp <-
      approx(
          x = as.numeric(samples$Date),
          y = samples$BPCA_mg_C_per_L,
          xout = as.numeric(.x$Date),
          method = "linear",
          rule = 1 #only interpolate within sampling period
      )$y 
    .x
  })
```

Get Daily Loads
```{r}
DailyLoads <- DailyConc %>%
  filter(!is.na(Levo_interp),
         !is.na(BPCA_interp)) %>%
  mutate(
    Levo_g_day = Q_mean * Levo_interp * 86400,
    BPCA_g_day = Q_mean * BPCA_interp * 86400
  )
```

Get Cumulative Export
```{r}
DailyLoads <- DailyLoads %>%
  group_by(Watershed) %>%
  arrange(Date) %>%
  mutate(
    Levo_cumulative_Mg = cumsum(Levo_g_day)/1e6,
    BPCA_cumulative_Mg = cumsum(BPCA_g_day)/1e6
  )
```

Revised estimate
```{r}
DailyLoads %>%
  group_by(Watershed) %>%
  summarize(
    Levo_export_Mg = max(Levo_cumulative_Mg),
    BPCA_export_Mg = max(BPCA_cumulative_Mg)
  )
```

Redo Plots
```{r}
#Make data for plot in long format
PlotData <- DailyLoads %>%
  group_by(Watershed) %>%
  mutate(
    Levo_fraction = Levo_cumulative_Mg / max(Levo_cumulative_Mg),
    BPCA_fraction = BPCA_cumulative_Mg / max(BPCA_cumulative_Mg)
  ) %>%
  select(Date, Watershed, Levo_fraction, BPCA_fraction) %>%
  pivot_longer(
    cols = c(Levo_fraction, BPCA_fraction),
    names_to = "Compound",
    values_to = "Fraction"
  ) %>%
  mutate(
    Compound = recode(
      Compound,
      Levo_fraction = "A) Levoglucosan",
      BPCA_fraction = "B) B6CA + B5CA"
    )
  )

#Define color scheme
ws_cols <- c(
  SAN  = "#D9C84A", #orignial color = "#FBF9A2"
  PESC = "#FFCF55",
  MAJ  = "#FF7410",
  LAG  = "#EA4412",
  SCT  = "#A00000"
)

#Set order for legend
PlotData <- PlotData %>%
  mutate(
    Watershed = factor(
      Watershed,
      levels = c("SAN", "PESC", "MAJ", "LAG", "SCT")
    )
  )

#Plot
ggplot(PlotData,
       aes(Date, Fraction,
           color = Watershed)) +

  geom_line(linewidth = 1.3) +

  facet_wrap(~Compound,
             ncol = 1,
             axes = "all") +

  scale_color_manual(values = ws_cols) +

  scale_y_continuous(
    limits = c(0,1),
    breaks = seq(0,1,0.25),
    expand = c(0,0)
  ) +

  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b"
  ) +

  labs(
    x = NULL,
    y = "Fraction of cumulative export"
  ) +

  # geom_vline(
  #   xintercept = as.Date(c(
  #     "2020-10-07",
  #     "2020-11-17",
  #     "2020-11-18",
  #     "2020-12-11",
  #     "2020-12-13",
  #     "2021-01-27",
  #     "2021-01-28",
  #     "2021-02-02",
  #     "2021-03-18",
  #     "2021-05-02"
  #   )),
  #   linetype = "dashed",
  #   color = "grey70",
  #   linewidth = 0.5
  # ) +
  
theme_classic(base_size = 13) +

theme(
  panel.background = element_rect(fill = "white", colour = NA),
  plot.background  = element_rect(fill = "white", colour = NA),
  strip.background = element_blank(),
  strip.text = element_text(
    face = "bold",
    hjust = 0
  ),
  axis.text.x = element_text(color = "black"),
  axis.ticks.x = element_line(color = "black")
)
```

Not fraction
```{r}
ggplot(DailyLoads, aes(Date, Levo_cumulative_Mg, color = Watershed)) +
  geom_line() +
  geom_vline(
    xintercept = as.numeric(as.Date(c(
      "2020-11-17",
      "2020-11-18",
      "2021-01-04",
      "2021-01-27",
      "2021-01-28"
    ))),
    linetype = "dashed"
  )
```

Get Export Timing
```{r}
ExportTiming_Levo <- DailyLoads %>%
  group_by(Watershed) %>%
  summarize(
    total_levo = max(Levo_cumulative_Mg),
    day_50 = Date[min(which(Levo_cumulative_Mg >= 0.5*total_levo))],
    day_90 = Date[min(which(Levo_cumulative_Mg >= 0.9*total_levo))]
  )

ExportTiming_BPCA <- DailyLoads %>%
  group_by(Watershed) %>%
  summarize(
    total_BPCA = max(BPCA_cumulative_Mg),
    day_50 = Date[min(which(BPCA_cumulative_Mg >= 0.5*total_BPCA))],
    day_90 = Date[min(which(BPCA_cumulative_Mg >= 0.9*total_BPCA))]
  )
```

