---
title: "Burn Severity - dnbr continuous & classified"
author: "rbb"
date: "2026-08-13"
output: html_document
---

Watershed boundaries were obtained from StreamStats and represent the upstream
contributing areas associated with each sampling location. The CZU Lightning
Complex Fire perimeter was obtained from the ArcGIS feature service. Continuous
dNBR and classified dNBR rasters were obtained from the Monitoring Trends in
Burn Severity (MTBS) dataset.The Albers projection was used because the WGS84
projection was rotated in the MTBS system.

The watershed polygons and fire perimeter are provided in WGS84, whereas the
dNBR rasters are provided in NAD83 / Conus Albers (EPSG:5070). Vector data are
reprojected to the dNBR raster CRS for spatial analysis.

Install packages if necessary - can skip if you already have these packages installed
```{r}
install.packages("terra")
install.packages("tidyr")
install.packages("dplyr")
```

Load libraries
```{r}
library(terra)
library(tidyr)
library(dplyr)
```

Read input data
```{r}
#read watershed polygons
SAN <- vect("data/GIS Files/streamstats/SAN.geojson")
PESC <- vect("data/GIS Files/streamstats/PESC.geojson")
LAG <- vect("data/GIS Files/streamstats/LAG.geojson")
MAJ <- vect("data/GIS Files/streamstats/MAJ.geojson")
SCT <- vect("data/GIS Files/streamstats/SCT.geojson")

#read in fire perimeter
fire_url <- paste0(
  "https://services1.arcgis.com/jJfZghspGKh8J9Jm/",
  "arcgis/rest/services/CZU_Fire_Perimeter/FeatureServer/3/query",
  "?where=1%3D1",
  "&outFields=*",
  "&returnGeometry=true",
  "&f=geojson"
)
fire_perimeter <- vect(fire_url)

#read dNBR raster
dnbr <- rast("data/GIS Files/mtbs-albers-dnbr/mtbs_ca3726212222320200816_10021177_20200724_20210711_dnbr.tif")

#read dnbr classified raster
dnbr_class <- rast("data/GIS Files/mtbs-albers-dnbr/mtbs_ca3726212222320200816_10021177_20200724_20210711_dnbr6.tif")
```

Prepare spatial data
```{r}
# Check coordinate reference systems
# crs(SAN) #WGS84
# crs(PESC) #WGS84
# crs(LAG) #WGS84
# crs(MAJ) #WGS84
# crs(SCT) #WGS84
# crs(fire_perimeter) #WGS84
# crs(dnbr) #Conus Albers

#Reproject vector data to the raster crs
#fire perimeter reprojection
fire_perimeter_5070 <- project(fire_perimeter, crs(dnbr))

#watershed layer reprojection
watersheds <- rbind(
  SAN,
  PESC,
  LAG,
  MAJ,
  SCT
)
watersheds_5070 <- project(watersheds, crs(dnbr))

#add watershed names
watersheds_5070$Watershed <- c(
  "SAN",
  "PESC",
  "LAG",
  "MAJ",
  "SCT"
)
```

Calculate continuous dnbr metrics
```{r}
#step 1: Clip/mask dnbr to fire perimeter and set outside dnbr to 0 (i.e. unburned) ----
#crop dnbr raster to fire perimeter extent
dnbr_crop <- crop(dnbr, fire_perimeter_5070)

#mask dNBR to the fire perimeter
dnbr_fire <- mask(dnbr_crop, fire_perimeter_5070)

#rasterize the fire perimeter onto the dNBR grid
fire_mask <- rasterize(
  fire_perimeter_5070,
  dnbr_fire,
  field = 1,
  background = NA
)

#set pixels outside the fire perimeter to 0
dnbr_fire_zero <- ifel(
  is.na(fire_mask),
  0,
  dnbr_fire
)

#step 2: Calculate watershed level mean dnbr ----
# Calculate mean dNBR for each watershed (areas outside fire perimeter assumed 0)
mean_dnbr <- terra::extract(
  dnbr_fire_zero,
  watersheds_5070,
  fun = mean,
  na.rm = TRUE
)

#clean table
mean_dnbr_df <- data.frame(
  Watershed = watersheds_5070$Watershed,
  Mean_dNBR = mean_dnbr[[2]]
)

#step 3: Calculate watershed burn extent ----
#calculate watershed areas
watersheds_5070$Watershed_area_m2 <- expanse(watersheds_5070)

#determine the number of burned pixels within each watershed
burned_cells <- terra::extract(
  fire_mask,
  watersheds_5070,
  fun = sum,
  na.rm = TRUE
)

#convert burned pixels to an area
cell_area <- prod(res(dnbr))
watersheds_5070$Burned_area_m2 <- burned_cells[[2]] * cell_area

#calculate burn extent of each watershed
watersheds_5070$Percent_burned <-
  watersheds_5070$Burned_area_m2 /
  watersheds_5070$Watershed_area_m2 * 100

#step 4: Compare dnbr within burn area - how severe was the fire (burned area only) ----
#calculate mean dNBR across burned pixels only
mean_dnbr_burned <- terra::extract(
  dnbr_fire,
  watersheds_5070,
  fun = mean,
  na.rm = TRUE
)

#add to watershed data
watersheds_5070$Mean_dNBR_burned <- mean_dnbr_burned[[2]]

#step 5: summarize continuous dnbr metrics ----
# Create summary table
burn_summary <- data.frame(
  Watershed = watersheds_5070$Watershed,
  Percent_burned = watersheds_5070$Percent_burned,
  Mean_dNBR_burned = watersheds_5070$Mean_dNBR_burned,
  Mean_dNBR_watershed = mean_dnbr_df$Mean_dNBR
)

burn_summary
```

Calculate categorical burn-severity extents
```{r}
#step 1: Make a classified raster that covers the entire watershed ----
#extend classified dnbr raster to cover all watersheds
dnbr_class_full <- extend(
  dnbr_class,
  ext(watersheds_5070),
)

#rasterize fire perimeter onto the full classified raster
fire_mask_class <- rasterize(
  fire_perimeter_5070,
  dnbr_class_full,
  field = 1,
  background = NA
)

#set areas outside the fire perimeter to Class 1
dnbr_class_full <- ifel(
  is.na(fire_mask_class),
  1,
  dnbr_class_full
)

#step 2: Calculate the 6 dnbr categories within each watershed ----
#count pixels in each dNBR class within each watershed
# Extract classified dNBR values for each watershed
# Extract classified dNBR values for each watershed
class_values <- terra::extract(
  dnbr_class_full,
  watersheds_5070
)

head(class_values)

#step 3: Count each class and calculate percentages ----
# Rename columns
names(class_values) <- c("ID", "Class")

# Remove NA values
class_values <- class_values[!is.na(class_values$Class), ]

# Count pixels in each class for each watershed
class_counts <- as.data.frame(
  table(
    class_values$ID,
    class_values$Class
  )
)

# Rename columns
names(class_counts) <- c(
  "ID",
  "Class",
  "Count"
)

# Calculate total number of pixels in each watershed
watershed_totals <- aggregate(
  Count ~ ID,
  data = class_counts,
  FUN = sum
)

# Add watershed totals to class table
class_counts <- merge(
  class_counts,
  watershed_totals,
  by = "ID",
  suffixes = c("", "_total")
)

# Calculate percentage of watershed
class_counts$Percent <- 
  class_counts$Count / class_counts$Count_total * 100

# Add watershed names
class_counts$Watershed <- c(
  "SAN",
  "PESC",
  "LAG",
  "MAJ",
  "SCT"
)[class_counts$ID]

# Add MTBS class names
class_counts$Class_name <- c(
  "Unburned to low",
  "Low",
  "Moderate",
  "High",
  "Increased greenness",
  "Non-processing"
)[class_counts$Class]

# Final table
burn_categories <- class_counts[, c(
  "Watershed",
  "Class",
  "Class_name",
  "Count",
  "Percent"
)]

burn_categories$Percent <- round(
  burn_categories$Percent, 
  2
)

burn_categories

#step 4: format classified dnbr table
burn_categories_wide <- burn_categories %>%
  select(Watershed, Class_name, Percent) %>%
  pivot_wider(
    names_from = Class_name,
    values_from = Percent
  )

burn_categories_wide
```

Combine continuous and classified burn metrics and export results
```{r}
# Reshape classified burn extents to one row per watershed
classified_extents <- burn_categories %>%
  select(Watershed, Class_name, Percent) %>%
  pivot_wider(
    names_from = Class_name,
    values_from = Percent
  )

# Combine with the original burn summary
final_burn_summary <- burn_summary %>%
  select(
    Watershed,
    Percent_burned,
    Mean_dNBR_burned,
    Mean_dNBR_watershed
  ) %>%
  left_join(
    classified_extents,
    by = "Watershed"
  )

# Round percentages and dNBR values
final_burn_summary <- final_burn_summary %>%
  mutate(
    Percent_burned = round(Percent_burned, 2),
    Mean_dNBR_burned = round(Mean_dNBR_burned, 2),
    Mean_dNBR_watershed = round(Mean_dNBR_watershed, 2)
  )

#Add Watershed_Name column to make it easier to combine with other dataframes
final_burn_summary <- final_burn_summary %>%
  mutate(
    Watershed_Name = case_when(
      Watershed == "SAN" ~ "San Lorenzo River",
      Watershed == "PESC" ~ "Pescadero Creek",
      Watershed == "LAG" ~ "Laguna Creek",
      Watershed == "MAJ" ~ "Majors Creek",
      Watershed == "SCT" ~ "Scott Creek",
      TRUE ~ Watershed
    )
  )

#Rename columns
final_burn_summary <- final_burn_summary %>%
  rename(
    BE_un_low = `Unburned to low`,
    BE_low = Low,
    BE_mod = Moderate,
    BE_high = High,
    `Inc-greenness` = `Increased greenness`,
    `Non-processing` = `Non-processing`
  )

#Save burn metrics as csv
write.csv(
  final_burn_summary,
  "data/watershed_burn_severity_summary.csv",
  row.names = FALSE
)
```
