---
title: "Coastal Fires Levoglucosan"
author: "RB"
date: "2024-09-13"
output: html_document
---
#load necessary libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(randomForest)
library(caret)
library(vegan)
library(ggrepel)
```

#set working director, load data files, and rename them
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/")
CombinedData <- readr::read_csv("CoastalFires_CombinedData.csv")
DischargeData <- readr::read_csv("CoastalFires_DischargeData.csv")
``` 
Test data for normality using Shapiro-Wilk Test
```{r}
#are all variables normally distributed?
#filter to only numeric data
numeric_columns <- sapply(CombinedData, is.numeric)
Data_numeric <- CombinedData[, numeric_columns]
Data_numeric <- Data_numeric %>% select(-lat, -long)

# Apply Shapiro-Wilk test and create a readable summary
shapiro_results <- sapply(Data_numeric, function(x) {
  test <- shapiro.test(x)
  return(c(p_value = test$p.value, W = test$statistic))
})

# Convert results to a data frame
shapiro_df <- as.data.frame(t(shapiro_results))
shapiro_df$Variable <- rownames(shapiro_df)
rownames(shapiro_df) <- NULL

# Print readable results
print(shapiro_df)
```

#The differences of concentrations and key ratios by watershed
```{r}
#is the concentration of levoglucosan different across watersheds?
kruskal_test_levo <- kruskal.test(levo_micro_gram_per_liter ~ watershed, data = CombinedData)
print(kruskal_test_levo)
kruskal_test_Qnorm_levo <- kruskal.test(Q_norm_levo ~ watershed, data = CombinedData)
print(kruskal_test_Qnorm_levo)
#is the ratio of levoglucosan to bulk DOC different across watersheds?
kruskal_test_levotoDOC <- kruskal.test(Levo_DOC_perc ~ watershed, data = CombinedData)
print(kruskal_test_levotoDOC)
#is the ratio of levoglucosan to DBC different across watersheds?
kruskal_test_levotoDBC <- kruskal.test(Levo_DBC ~ watershed, data = CombinedData)
print(kruskal_test_levotoDBC)
```

#since there is a significant difference between watersheds for Q-norm levoglucosan are significantly different?
```{r}
#Q_norm_levo is the concentration of levoglucosan divided by discharge
wilcoxon_test_Qnorm_levo <- pairwise.wilcox.test(CombinedData$Q_norm_levo, CombinedData$watershed, p.adjust.method = "bonferroni")
print(wilcoxon_test_Qnorm_levo)
```

#Time since fire - Spearman's rank
```{r}
#make a table to showcase the coefficients and p-values
# Perform Spearman's rank correlation using the averaged data
spearman_corr_levo_ppb <- cor.test(CombinedData$time_since_fire_day, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_Qnorm_levo <- cor.test(CombinedData$time_since_fire_day, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_levoDOC <- cor.test(CombinedData$time_since_fire_day, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_DBC_ppm <- cor.test(CombinedData$time_since_fire_day, CombinedData$DBC_mgC_per_L, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_dbcDOC <- cor.test(CombinedData$time_since_fire_day, CombinedData$DBC_DOC_perc, method = "spearman", use = "pairwise.complete.obs")

# Extract p-values and rho values
results <- data.frame(
  Comparison = c("time_since_fire_day vs levo_ppb",
                  "time_since_fire_day vs Qnorm_levo",
                 "time_since_fire_day vs levoDOC",
                  "time_since_fire_day vs DBC",
                  "time_since_fire_day vs dbcDOC"),
  Rho = c(spearman_corr_levo_ppb$estimate,
          spearman_corr_Qnorm_levo$estimate,
          spearman_corr_levoDOC$estimate,
          spearman_corr_DBC_ppm$estimate,
          spearman_corr_dbcDOC$estimate),
  P_Value = c(spearman_corr_levo_ppb$p.value,
              spearman_corr_Qnorm_levo$p.value,
              spearman_corr_levoDOC$p.value,
              spearman_corr_DBC_ppm$p.value,
              spearman_corr_dbcDOC$p.value)
)

# Print the results table
print(results)
```

#Miscellaneous Correlation Significant Tests for Levoglucosan Concentrations
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by discharge
spearman_corr_Q <- cor.test(CombinedData$Q_m_sq_per_sec, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")

# Extract p-values and rho values
results <- data.frame(
  Comparison = c("levo_ppb vs burn_extent",
                  "levo_ppb vs low burn_extent",
                  "levo_ppb vs low-mod burn_extent",
                 "levo_ppb vs mod-high burn extent", 
                 "levo_ppb vs watershed area", 
                 "levo_ppb vs watershed slope",
                 "levo_ppb vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)


```

#Miscellaneous Correlation Significant Tests for Discharge normalized Levoglucosan concentrations
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")


# Extract p-values and rho values
results <- data.frame(
  Comparison = c("Qnorm_levo vs burn_extent",
                  "Qnorm_levo vs low burn_extent",
                  "Qnorm_levo vs low-mod burn_extent",
                 "Qnorm_levo vs mod-high burn extent", 
                 "Qnorm_levo vs watershed area", 
                 "Qnorm_levo vs watershed slope",
                 "Qnorm_levo vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)
```

#Miscellaneous Correlation Significant Tests for Levo:DOC
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by discharge
spearman_corr_Q <- cor.test(CombinedData$Q_m_sq_per_sec, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")

#Extract p-values and rho values
results <- data.frame(
  Comparison = c("Levo_DOC_perc vs burn_extent",
                  "Levo_DOC_perc vs low burn_extent",
                  "Levo_DOC_perc vs low-mod burn_extent",
                 "Levo_DOC_perc vs mod-high burn extent", 
                 "Levo_DOC_perc vs watershed area", 
                 "Levo_DOC_perc vs watershed slope",
                 "Levo_DOC_perc vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)
```

#NMDS
#Prepare data for NMDS
```{r}
#Isolate PyC concentrations and relevant ratios 
PyC_conc_rat <- CombinedData[, c("levo_micro_gram_per_liter", "DOC_mgC_per_L", "DBC_mgC_per_L", "DBC_DOC_perc", "Levo_DOC_perc")]

#put levoglucosan concentrations into the same units as DOC and DBC
PyC_conc_rat$levo_mg_per_L <- PyC_conc_rat$levo_micro_gram_per_liter/1000 #dividing micrograms per liter by 1000 to get levoglucosan concentrations in milligrams per liter

#delete the levoglucosan column with concentrations in micrograms per liter
analyte_data_conc_rat <- PyC_conc_rat[,c("levo_mg_per_L", "DOC_mgC_per_L", "DBC_mgC_per_L", "DBC_DOC_perc", "Levo_DOC_perc")]

View(analyte_data_conc_rat)
```

#Run NMDS
```{r}
nmds_result <- metaMDS(analyte_data_conc_rat, k = 2, trymax = 100)
print(nmds_result)
```


#Plot NMDS Results
```{r}
# Extract NMDS scores for plotting
nmds_scores <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_species<-as.data.frame(scores(nmds_result, display="species"))
nmds_scores$Site <- CombinedData$watershed  # Add Site info for coloring
nmds_scores$time <- CombinedData$time_since_fire_day #this is for coloring the data points by time
nmds_scores$watershed <- CombinedData$watershed #this is for the point shape by watershed

# Environmental variables (adjust to your desired variables)
env_variables <- CombinedData[c("time_since_fire_day", "Area_low_severity_percent", "Q_m_sq_per_sec", "Area_mod_to_high_severity_percent", "Area_square_kilometer")]

#If want ALL environmental variables for sasha, use the code below
#env_variables <- CombinedData[c("time_since_fire_day", "Area_low_severity_percent", "Q_m_sq_per_sec", "Area_square_kilometer", "Average_slope_percent", "Burn_Extent_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent")]

#assign environmental variables to df
envfit_result <- envfit(nmds_result, env_variables, permutations = 999)

#have the environmental variables assigned to arrows for plotting
en_coord_cont = as.data.frame(scores(envfit_result, "vectors")) * ordiArrowMul(envfit_result)

# Shorten labels for plotting species and vectors
nmds_species$ShortLabel <- c("Levo", "DOC", "DBC", "DBC:DOC", "Levo:DOC")  # Replace with your short labels

# Change labels for plotting vectors. Also wrapping text for improving clarity of the plot
en_coord_cont$WrappedLabel <- c(
  "Time Since\nFire", #Time Since FIre
  "Low Severity\nBurn Extent", #Low Seveirty Burn Extent
  "Discharge", #Discharge
  "Mod-High\nBurn Extent", #Moderate to High Burn Extent
  "Watershed\nArea" #Watershed Area
)
  
#NMDS Visual Plot
ggplot(nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  
  #Sites
  geom_point(data = nmds_scores, aes(size = 100, color = time, fill = time, shape = watershed)) +
     scale_shape_manual(values = c(21, 22, 24, 23, 25))+
  scale_fill_gradient(low="#00FFFF", high="navy")+
  scale_color_gradient(low="#00FFFF", high="navy")+
 
  #Species
  geom_point(data = nmds_species, aes(x = NMDS1, y = NMDS2), shape = 8, stroke = 2, size = 3, alpha = 1, colour = "hotpink") +
  
  #vectors
  geom_segment(aes(x = 0, y = 0, xend = NMDS1*0.75, yend = NMDS2*0.75), data = en_coord_cont, linewidth =1, alpha = 1, colour = "black", arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +

#Add repelled labels for species and vectors
  geom_text_repel(data = nmds_species, aes(x = NMDS1, y = NMDS2, label = ShortLabel), colour = "black", fontface = "bold", size = 3, box.padding = 0.2, max.overlaps = Inf) + #species
  geom_text_repel(data = en_coord_cont, aes(x = NMDS1*0.75, y = NMDS2*0.75, label = WrappedLabel), size = 3, box.padding = 0.5, max.overlaps = 50) + #vectors
  
  
  #chart title
  labs(x = "NMDS1",
       y = "NMDS2") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(size = 0.5),  # Add axis lines
    axis.ticks = element_line(size = 0.5),  # Add axis ticks
    axis.ticks.length = unit(3, "pt"),  # Make ticks outward
    axis.line.x = element_line(size = 0.5),
    axis.line.y = element_line(size = 0.5)
  )

#Determine which vectors are statistically significant
envfit_result
```









#Random Forest Models

#prepare data for RFM
```{r}
#remove any missing values
na.omit(CombinedData) 
#ensure categorical variables are factors
CombinedData$watershed <- as.factor(CombinedData$watershed) 
```

#RFM for Levoglucosan concentrations
```{r}
#Split levoglucosan dataset into model and test datasets
set.seed(123)  # For reproducibility
train_idx <- createDataPartition(CombinedData$levo_micro_gram_per_liter, p = 0.7, list = FALSE)
train_data <- CombinedData[train_idx, ]
test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_levoglucosan <- randomForest(
    levo_micro_gram_per_liter ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_levoglucosan)
varImpPlot(rf_levoglucosan)

#Evaluate RFM Performance

#predict of the test set
predictions <- predict(rf_levoglucosan, newdata = test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - test_data$levo_micro_gram_per_liter)^2))
r_squared <- cor(predictions, test_data$levo_micro_gram_per_liter)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

#Analyze varibale importance
importance(rf_levoglucosan)

#visualize variable importance
varImpPlot(rf_levoglucosan)

#Tune the RFM

control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(train_data) - 1, by = 1)  # Tune mtry
)

rf_tuned <- train(
    levo_micro_gram_per_liter ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)

#use best model
levo_best_rf_model <- rf_tuned$finalModel
varImpPlot(levo_best_rf_model)
```
#Plotting the RFM Results for Levoglucosan Concentrations
```{r}
# Extract and format variable importance
var_imp <- as.data.frame(importance(rf_levoglucosan))
var_imp$Variable <- rownames(var_imp)  # Convert rownames to a column

# Rename columns to something more user-friendly if necessary
colnames(var_imp) <- c("MeanDecreaseAccuracy", "MeanDecreaseGini", "Variable")

# Replace variable names with custom labels
custom_labels <- c(
    "Time Since Fire", "Flow Rate (m³/s)", "Area (km²)", "Slope (%)",
    "Burn Extent (%)", "Low Severity Area (%)", 
    "Low-Moderate Severity Area (%)", "Moderate-High Severity Area (%)"
)
var_imp$Variable <- factor(custom_labels, levels = custom_labels)

#Plotting variable importance for visual clarity
ggplot(var_imp, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
    geom_bar(stat = "identity", fill = "pink") +
    coord_flip() +
    labs(
        title = "Levoglucosan - Variable Importance Plot",
        x = "Variables",
        y = "Mean Decrease in Accuracy"
    ) +
    theme_minimal(base_size = 14) +
    theme(
        panel.grid = element_blank(),  # Remove gridlines
        plot.title = element_text(hjust = 0.5),  # Center the title
        axis.line = element_line(),  # Add axis lines
        axis.ticks = element_line(),  # Add axis ticks
        axis.ticks.length = unit(0.2, "cm")  # Customize tick length
    )

#Display key statistical values in a tabular format
# Extract variable importance
levo_importance_data <- importance(levo_best_rf_model)

# Convert to a data frame for better readability
levo_importance_table <- as.data.frame(levo_importance_data)
levo_importance_table$Variable <- rownames(levo_importance_table)
rownames(levo_importance_table) <- NULL

# Print the table
print(levo_importance_table)
```


#RFM for for dissolved black carbon :) 
```{r}
#Split DBC dataset into model and test datasets
set.seed(123)  # For reproducibility
DBC_train_idx <- createDataPartition(CombinedData$DBC_mgC_per_L, p = 0.7, list = FALSE)
DBC_train_data <- CombinedData[train_idx, ]
DBC_test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_DBC <- randomForest(
    DBC_mgC_per_L ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_DBC)
varImpPlot(rf_DBC)

#Evaluate RFM Performance
#predict of the test set
predictions <- predict(rf_DBC, newdata = DBC_test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - DBC_test_data$DBC_mgC_per_L)^2))
r_squared <- cor(predictions, DBC_test_data$DBC_mgC_per_L)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

##Analyze varibale importance

#something
importance(rf_DBC)

#visualize variable importance
varImpPlot(rf_DBC)

#tune the DBC RFM
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(DBC_train_data) - 1, by = 1)  # Tune mtry
)

DBC_rf_tuned <- train(
    DBC_mgC_per_L ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)


#use best model
DBC_best_rf_model <- DBC_rf_tuned$finalModel
varImpPlot(DBC_best_rf_model)

```

#Plotting the RFM Results for DBC Concentrations
```{r}
# Extract and format variable importance
var_imp <- as.data.frame(importance(DBC_best_rf_model))
var_imp$Variable <- rownames(var_imp)  # Convert rownames to a column

# Rename columns to something more user-friendly if necessary
colnames(var_imp) <- c("MeanDecreaseAccuracy", "MeanDecreaseGini", "Variable")

# Replace variable names with custom labels
custom_labels <- c(
    "Time Since Fire", "Discharge (m³/s)", "Area (km²)", "Slope (%)",
    "Burn Extent (%)", "Low Severity Area (%)", 
    "Low-Moderate Severity Area (%)", "Moderate-High Severity Area (%)"
)
var_imp$Variable <- factor(custom_labels, levels = custom_labels)

#Plotting variable importance for visual clarity
ggplot(var_imp, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
    geom_bar(stat = "identity", fill = "pink") +
    coord_flip() +
    labs(
        title = "DBC - Variable Importance Plot",
        x = "Variables",
        y = "Mean Decrease in Accuracy"
    ) +
    theme_minimal(base_size = 14) +
    theme(
        panel.grid = element_blank(),  # Remove gridlines
        plot.title = element_text(hjust = 0.5),  # Center the title
        axis.line = element_line(),  # Add axis lines
        axis.ticks = element_line(),  # Add axis ticks
        axis.ticks.length = unit(0.2, "cm")  # Customize tick length
    )

#Display key statistical values in a tabular format
# Extract variable importance
DBC_importance_data <- importance(DBC_best_rf_model)

# Convert to a data frame for better readability
DBC_importance_table <- as.data.frame(DBC_importance_data)
DBC_importance_table$Variable <- rownames(DBC_importance_table)
rownames(importance_table) <- NULL

# Print the table
print(importance_table)
```
#RFM for Levo:DOC :) 
#```{r}
#Split DBC dataset into model and test datasets
set.seed(123)  # For reproducibility
LevoDOC_train_idx <- createDataPartition(CombinedData$Levo_DOC_perc, p = 0.7, list = FALSE)
LevoDOC_train_data <- CombinedData[train_idx, ]
LevoDOC_test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_LevoDOC <- randomForest(
    Levo_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_LevoDOC)
varImpPlot(rf_LevoDOC)

#Evaluate RFM Performance
#predict of the test set
predictions <- predict(rf_LevoDOC, newdata = LevoDOC_test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - LevoDOC_test_data$Levo_DOC_perc)^2))
r_squared <- cor(predictions, LevoDOC_test_data$Levo_DOC_perc)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

##Analyze varibale importance

#something
importance(rf_LevoDOC)

#visualize variable importance
varImpPlot(rf_LevoDOC)

#tune the DBC RFM
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(LevoDOC_train_data) - 1, by = 1)  # Tune mtry
)

LevoDOC_rf_tuned <- train(
    Levo_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)


#use best model
LevoDOC_best_rf_model <- DBC_rf_tuned$finalModel
varImpPlot(LevoDOC_best_rf_model)

```

#Plotting the RFM Results for Levo:DOC Concentrations
#```{r}
# Extract and format variable importance
var_imp <- as.data.frame(importance(LevoDOC_best_rf_model))
var_imp$Variable <- rownames(var_imp)  # Convert rownames to a column

# Rename columns to something more user-friendly if necessary
colnames(var_imp) <- c("MeanDecreaseAccuracy", "MeanDecreaseGini", "Variable")

# Replace variable names with custom labels
custom_labels <- c(
    "Time Since Fire", "Discharge (m³/s)", "Area (km²)", "Slope (%)",
    "Burn Extent (%)", "Low Severity Area (%)", 
    "Low-Moderate Severity Area (%)", "Moderate-High Severity Area (%)"
)
var_imp$Variable <- factor(custom_labels, levels = custom_labels)

#Plotting variable importance for visual clarity
ggplot(var_imp, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
    geom_bar(stat = "identity", fill = "pink") +
    coord_flip() +
    labs(
        title = "Levo:DOC - Variable Importance Plot",
        x = "Variables",
        y = "Mean Decrease in Accuracy"
    ) +
    theme_minimal(base_size = 14) +
    theme(
        panel.grid = element_blank(),  # Remove gridlines
        plot.title = element_text(hjust = 0.5),  # Center the title
        axis.line = element_line(),  # Add axis lines
        axis.ticks = element_line(),  # Add axis ticks
        axis.ticks.length = unit(0.2, "cm")  # Customize tick length
    )

#Display key statistical values in a tabular format
# Extract variable importance
LevoDOC_importance_data <- importance(LevoDOC_best_rf_model)

# Convert to a data frame for better readability
LevoDOC_importance_table <- as.data.frame(LevoDOC_importance_data)
LevoDOC_importance_table$Variable <- rownames(LevoDOC_importance_table)
rownames(LevoDOC_importance_table) <- NULL

# Print the table
print(LevoDOC_importance_table)
```
#RFM for DBC:DOC :) 
#```{r}
#Split DBC dataset into model and test datasets
set.seed(123)  # For reproducibility
DBCDOC_train_idx <- createDataPartition(CombinedData$DBC_DOC_perc, p = 0.7, list = FALSE)
DBCDOC_train_data <- CombinedData[train_idx, ]
DBCDOC_test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_DBCDOC <- randomForest(
    DBC_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_DBCDOC)
varImpPlot(rf_DBCDOC)

#Evaluate RFM Performance
#predict of the test set
predictions <- predict(rf_DBCDOC, newdata = DBCDOC_test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - DBCDOC_test_data$DBC_DOC_perc)^2))
r_squared <- cor(predictions, DBCDOC_test_data$DBC_DOC_perc)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

##Analyze varibale importance

#something
importance(rf_DBCDOC)

#visualize variable importance
varImpPlot(rf_DBCDOC)

#tune the DBC RFM
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(DBCDOC_train_data) - 1, by = 1)  # Tune mtry
)

DBCDOC_rf_tuned <- train(
    DBC_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)


#use best model
DBCDOC_best_rf_model <- DBC_rf_tuned$finalModel
varImpPlot(DBCDOC_best_rf_model)

```
#Plotting the RFM Results for DBC:DOC Concentrations
#```{r}
# Extract and format variable importance
var_imp <- as.data.frame(importance(DBCDOC_best_rf_model))
var_imp$Variable <- rownames(var_imp)  # Convert rownames to a column

# Rename columns to something more user-friendly if necessary
colnames(var_imp) <- c("MeanDecreaseAccuracy", "MeanDecreaseGini", "Variable")

# Replace variable names with custom labels
custom_labels <- c(
    "Time Since Fire", "Discharge (m³/s)", "Area (km²)", "Slope (%)",
    "Burn Extent (%)", "Low Severity Area (%)", 
    "Low-Moderate Severity Area (%)", "Moderate-High Severity Area (%)"
)
var_imp$Variable <- factor(custom_labels, levels = custom_labels)

#Plotting variable importance for visual clarity
ggplot(var_imp, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
    geom_bar(stat = "identity", fill = "pink") +
    coord_flip() +
    labs(
        title = "Levo:DOC - Variable Importance Plot",
        x = "Variables",
        y = "Mean Decrease in Accuracy"
    ) +
    theme_minimal(base_size = 14) +
    theme(
        panel.grid = element_blank(),  # Remove gridlines
        plot.title = element_text(hjust = 0.5),  # Center the title
        axis.line = element_line(),  # Add axis lines
        axis.ticks = element_line(),  # Add axis ticks
        axis.ticks.length = unit(0.2, "cm")  # Customize tick length
    )

#Display key statistical values in a tabular format
# Extract variable importance
DBCDOC_importance_data <- importance(DBCDOC_best_rf_model)

# Convert to a data frame for better readability
DBCDOC_importance_table <- as.data.frame(DBCDOC_importance_data)
DBCDOC_importance_table$Variable <- rownames(DBCDOC_importance_table)
rownames(DBCDOC_importance_table) <- NULL

# Print the table
print(DBCDOC_importance_table)
```
