---
title: "CoastalFires_CumulativeExport"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---

Packages
```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(patchwork)
```


Get Data
```{r}
#From CUAHSI Hydroshare
DischargeData <- readr::read_csv("https://www.hydroshare.org/resource/26f08ef7fe344fcb9d3da45160b697fe/data/contents/CoastalFires_DischargeData.csv")

#From GitHub
CombinedData <- readr::read_csv("data/CoastalFires_CombinedData.csv")
```

Exclude samples that were taken at a different location. 
```{r}
#Scott Creek flooded during the events of January 27-28th, 2021. Samples were collected at a downstream lagoon site. However, due to this the data collected does not accurately reflect that at the sample site. As such authors of the associated manuscript have excluded it from all data analysis and figure creation.
CombinedData <- CombinedData %>%
  filter(Location_Change != "Y")
```


Make sure dates are right
```{r}
CombinedData <- CombinedData %>%
  mutate(
    DateTime = as.POSIXct(Date, tz = "America/Los_Angeles") +
      if_else(is.na(Time),
              12*60*60,
              as.numeric(Time)),
    DateTime = format(DateTime, "%Y-%m-%d %H:%M")
  )
```

Get data for cumulative export
```{r}
ConcData <- CombinedData %>%
  select(Watershed, DateTime, Levo_micro_mol_C_per_L, stdev_Levo_micro_mol_C_per_L, B6CA_and_B5CA_micro_mol_C_per_L)

#Fix watershed names to match discharge data
ConcData <- ConcData %>%
  mutate(Watershed = recode(Watershed,
                            "Laguna Creek" = "LAG",
                            "San Lorenzo River" = "SAN",
                            "Pescadero Creek" = "PESC",
                            "Majors Creek" = "MAJ",
                            "Scott Creek" = "SCT"))
#Make sure date is correct format
ConcData <- ConcData %>%
  mutate(
    DateTime = as.POSIXct(DateTime,
                          format = "%Y-%m-%d %H:%M",
                          tz = "UTC")
  )
```

Convert concentrations
```{r}
#to get gC from uM-C, multiply by 0.012011
ConcData$Levo_mg_C_per_L <- ConcData$Levo_micro_mol_C_per_L * 0.012011
ConcData$BPCA_mg_C_per_L <- ConcData$B6CA_and_B5CA_micro_mol_C_per_L * 0.012011

#remove unneccessary columns
ConcData <- ConcData %>%
  select(-Levo_micro_mol_C_per_L,
         -B6CA_and_B5CA_micro_mol_C_per_L)
```

Get Discharge set
```{r}
SAN <- DischargeData %>%
  transmute(
    Watershed = "SAN",
    DateTime = parse_date_time(SAN_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = SAN_Flow_cms
  )

PESC <- DischargeData %>%
  transmute(
    Watershed = "PESC",
    DateTime = parse_date_time(PESC_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = PESC_Flow_cms
  )

LAG <- DischargeData %>%
  transmute(
    Watershed = "LAG",
    DateTime = parse_date_time(LAG_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = LAG_Flow_cms
  )

MAJ <- DischargeData %>%
  transmute(
    Watershed = "MAJ",
    DateTime = parse_date_time(MAJ_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = MAJ_Flow_cms
  )

SCT <- DischargeData %>%
  transmute(
    Watershed = "SCT",
    DateTime = parse_date_time(SCT_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = SCT_Flow_cms
  )

FlowData <- bind_rows(SAN, PESC, LAG, MAJ, SCT)
```

Remove NAs from FlowData
```{r}
FlowData <- FlowData %>%
  filter(!is.na(DateTime),
         !is.na(Flow))

#check it worked
FlowData %>%
  group_by(Watershed) %>%
  summarize(
    missing_dates = sum(is.na(DateTime)),
    start = min(DateTime),
    end = max(DateTime),
    n=n()
  )
```


Convert to daily discharge (highest resolution for Scott Creek)
```{r}
DailyFlow <- FlowData %>%
  mutate(Date = as.Date(DateTime)) %>%
  filter(Date >= as.Date("2020-10-07"),
         Date <= as.Date("2021-05-02")) %>%
  group_by(Watershed, Date) %>%
  summarize(
    Q_mean = mean(Flow, na.rm=TRUE),
    .groups="drop"
  )

#check
DailyFlow %>%
  count(Watershed)
```

Get Daily Concentrations
```{r}
DailyConc <- DailyFlow %>%
  group_by(Watershed) %>%
  group_modify(~{

    ws <- .y$Watershed

    samples <- ConcData %>%
      filter(Watershed == ws) %>%
      arrange(DateTime)

    if(nrow(samples) < 2){
      .x$Levo_interp <- NA
      .x$BPCA_interp <- NA
      return(.x)
    }

    .x$Levo_interp <-
      approx(
        x = as.numeric(samples$DateTime),
        y = samples$Levo_mg_C_per_L,
        xout = as.numeric(as.POSIXct(.x$Date)),
        rule = 1
      )$y

    .x$BPCA_interp <-
      approx(
        x = as.numeric(samples$DateTime),
        y = samples$BPCA_mg_C_per_L,
        xout = as.numeric(as.POSIXct(.x$Date)),
        rule = 1
      )$y

    .x
  })
```

Get Daily Loads
```{r}
DailyLoads <- DailyConc %>%
  filter(!is.na(Levo_interp),
         !is.na(BPCA_interp)) %>%
  mutate(
    Levo_g_day = Q_mean * Levo_interp * 86400,
    BPCA_g_day = Q_mean * BPCA_interp * 86400
  )
```

Get Cumulative Export
```{r}
DailyLoads <- DailyLoads %>%
  group_by(Watershed) %>%
  arrange(Date) %>%
  mutate(
    Levo_cumulative_Mg = cumsum(Levo_g_day)/1e6,
    BPCA_cumulative_Mg = cumsum(BPCA_g_day)/1e6
  )
```

Revised estimate
```{r}
DailyLoads %>%
  group_by(Watershed) %>%
  summarize(
    Levo_export_Mg = max(Levo_cumulative_Mg),
    BPCA_export_Mg = max(BPCA_cumulative_Mg)
  )
```

Calc Cum Exp & Plot
```{r}
# LEVO

DailyLoads <- DailyLoads %>%
  group_by(Watershed) %>%
  mutate(
    Levo_fraction = Levo_cumulative_Mg /
      max(Levo_cumulative_Mg),

    BPCA_fraction = BPCA_cumulative_Mg /
      max(BPCA_cumulative_Mg)
  )

ggplot(DailyLoads,
       aes(Date, Levo_fraction, color=Watershed)) +
  geom_line(linewidth=1) +
  labs(
    y="Fraction of cumulative levoglucosan export",
    x=NULL
  )

#BPCA
ggplot(DailyLoads,
       aes(Date, BPCA_fraction, color=Watershed)) +
  geom_line(linewidth=1) +
  labs(
    y="Fraction of cumulative BPCA export",
    x=NULL
  )

```
Not fraction
```{r}
ggplot(DailyLoads, aes(Date, Levo_cumulative_Mg, color = Watershed)) +
  geom_line() +
  geom_vline(
    xintercept = as.numeric(as.Date(c(
      "2020-11-17",
      "2020-11-18",
      "2021-01-04",
      "2021-01-27",
      "2021-01-28"
    ))),
    linetype = "dashed"
  )
```

Get Export Timing
```{r}
ExportTiming_Levo <- DailyLoads %>%
  group_by(Watershed) %>%
  summarize(
    total_levo = max(Levo_cumulative_Mg),
    day_50 = Date[min(which(Levo_cumulative_Mg >= 0.5*total_levo))],
    day_90 = Date[min(which(Levo_cumulative_Mg >= 0.9*total_levo))]
  )

ExportTiming_BPCA <- DailyLoads %>%
  group_by(Watershed) %>%
  summarize(
    total_BPCA = max(BPCA_cumulative_Mg),
    day_50 = Date[min(which(BPCA_cumulative_Mg >= 0.5*total_BPCA))],
    day_90 = Date[min(which(BPCA_cumulative_Mg >= 0.9*total_BPCA))]
  )
```

Trying to Make Fig 2
#get mean error of replicates for error bars
```{r}
mean_levo_sd <- CombinedData %>%
  filter(!is.na(stdev_Levo_micro_mol_C_per_L)) %>%
  summarize(
    mean_sd = mean(stdev_Levo_micro_mol_C_per_L)
  ) %>%
  pull(mean_sd)
mean_levo_sd
```

Create plot dataset
```{r}
# Calculate mean analytical standard deviation from replicate measurements
mean_levo_sd <- CombinedData %>%
  filter(!is.na(stdev_Levo_micro_mol_C_per_L)) %>%
  summarize(
    mean_sd = mean(stdev_Levo_micro_mol_C_per_L, na.rm = TRUE)
  ) %>%
  pull(mean_sd)


# Concentration data for plotting
Levo_plot <- CombinedData %>%
  filter(!is.na(Levo_micro_mol_C_per_L)) %>%
  mutate(
    Date = as.Date(DateTime),
    Levo_error = mean_levo_sd
  ) %>%
  filter(Watershed == "San Lorenzo River")


# Discharge data for plotting
flow_plot <- DailyFlow %>%
  filter(Watershed == "SAN") %>%
  mutate(
    Q_log = log10(Q_mean + 0.001) - min(log10(Q_mean + 0.001), na.rm = TRUE) #normalizes while keeping it positive
  )


# Scaling factor for secondary axis
# Set desired axis ranges
levo_max <- 2.5
q_max <- 100

# Transform the discharge axis limits
q_log_max <- log10(q_max + 0.001) - 
  min(log10(flow_plot$Q_mean + 0.001), na.rm = TRUE)

scale_factor <- levo_max / q_log_max

# Plot
ggplot() +

  # Hydrograph
  geom_line(
    data = flow_plot,
    aes(
      x = Date,
      y = Q_log * scale_factor
    ),
    color = "#156082",
    linewidth = 1.2
  )+

  # Levoglucosan error bars
  geom_errorbar(
    data = Levo_plot,
    aes(
      x = Date,
      ymin = Levo_micro_mol_C_per_L - Levo_error,
      ymax = Levo_micro_mol_C_per_L + Levo_error
    ),
    width = 0,
    linewidth = 0.4,
    color = "black"
  ) +

  # Levoglucosan concentrations
  geom_point(
    data = Levo_plot,
    aes(
      x = Date,
      y = Levo_micro_mol_C_per_L
    ),
    shape = 23,
    size = 4,
    fill = "#FBF9A2",
    color = "black",
    stroke = 0.8
  ) +

  scale_y_continuous(
    limits = c(0, 2.5),
    name = expression(
      "Levoglucosan ("*mu*"mol C L"^{-1}*")"
    ),
    sec.axis = sec_axis(
      ~ . / scale_factor,
      breaks = (
        log10(c(0.01, 0.1, 1, 10, 100) + 0.001) -
          min(log10(flow_plot$Q_mean + 0.001), na.rm = TRUE)
      ) * scale_factor,
      labels = c("0.01", "0.1", "1", "10", "100"),
      name = expression(
        "Discharge (m"^3*" s"^{-1}*")"
      )
    )
  ) +
  #Watershed label
  annotate(
   "text",
   x = as.Date("2020-10-01"),
   y = 2.5,
   label = "A. San Lorenzo River",
   hjust = 0,
   fontface = "bold"
  ) +

  theme_bw(base_size = 12) +

  theme(
    panel.grid.minor = element_blank()
  )
```
Trying to make a levoglucosan plot function
```{r}
make_levo_plot <- function(
  watershed_name,
  flow_name,
  panel_label,
  point_shape,
  point_fill
) {
  
  # Concentration data
  Levo_plot <- CombinedData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.Date(DateTime),
      Levo_error = mean_levo_sd
    )
  
  # Discharge data
  flow_plot <- DailyFlow %>%
    filter(Watershed == flow_name) %>%
    mutate(
      Q_log = log10(Q_mean + 0.001)
    )
  
  # Scaling
  q_min <- 0.01
  q_max <- 100
  
  q_low <- log10(q_min)
  q_high <- log10(q_max)
  
  levo_max <- 2.5
  
  scale_factor <- levo_max / (q_high - q_low)
    
  # Plot
  ggplot() +
    
    # Hydrograph
    geom_line(
      data = flow_plot,
      aes(
        x = Date,
        y = (log10(Q_mean) - q_low) * scale_factor
      ),
      color = "#156082",
      linewidth = 1
    ) +
    
    # Levoglucosan connecting line
    geom_line(
      data = Levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      color = "grey30",
      linewidth = 0.45,
      linetype = "dashed",
      alpha = 0.7
    ) +
    # Error bars
    geom_errorbar(
      data = Levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4,
      color = "black"
    ) +
    
    # Points
    geom_point(
      data = Levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = point_shape,
      size = 3,
      fill = point_fill,
      color = "black",
      stroke = 0.8
    ) +
    
    # Axes
    scale_y_continuous(
      limits = c(0, 2.5),
      name = expression(
        "Levoglucosan ("*mu*"mol C L"^{-1}*")"
      ),
      sec.axis = sec_axis(
        ~ (. / scale_factor) + q_low,
        breaks = log10(c(0.01, 0.1, 1, 10, 100)),
        labels = c("0.01", "0.1", "1", "10", "100"),
        name = expression(
          "Discharge (m"^3*" s"^{-1}*")"
        )
      )
      ) +
    
    # Label
    annotate(
      "text",
      x = as.Date("2020-10-01"),
      y = 2.4,
      label = panel_label,
      hjust = 0,
      fontface = "bold",
      size = 5
    ) +
    
    #Scale x-axis
    scale_x_date(
      limits = as.Date(c("2020-10-01", "2021-05-02")),
      date_breaks = "1 month",
      date_labels = "%b"
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid = element_blank()
    )
}
  
```

Plot with function
```{r}
plot_SAN <- make_levo_plot(
  "San Lorenzo River",
  "SAN",
  "A)",
  point_shape = 23,
  point_fill = "#FBF9A2"
)


plot_PESC <- make_levo_plot(
  "Pescadero Creek",
  "PESC",
  "B)",
  point_shape = 24,
  point_fill = "#FFCF55"
)


plot_MAJ <- make_levo_plot(
  "Majors Creek",
  "MAJ",
  "C)",
  point_shape = 22,
  point_fill = "#FF7410"
)


plot_LAG <- make_levo_plot(
  "Laguna Creek",
  "LAG",
  "D)",
  point_shape = 21,
  point_fill = "#EA4412"
) 


plot_SCT <- make_levo_plot(
  "Scott Creek",
  "SCT",
  "E)",
  point_shape = 25,
  point_fill = "#A00000"
) 

#To remove the X-axis label from some of the plots
plot_SAN <- plot_SAN +
  theme(axis.title.x = element_blank())

plot_PESC <- plot_PESC +
  theme(axis.title.x = element_blank())

plot_LAG <- plot_LAG +
  theme(axis.title.x = element_blank())
```

Trying to plot with patchwork
```{r}
final_fig <- (
  plot_SAN /
  plot_PESC /
  plot_MAJ
) |
(
  plot_LAG /
  plot_SCT /
  plot_spacer()
)

final_fig <- final_fig +
  plot_layout(
    widths = c(1,1)
  ) &
  theme(
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    plot.tag.position = c(0.5,1.1),
    plot.tag = element_text(
      face = "bold",
      size = 12
    )
  )
```

Save
```{r}
ggsave(
  "Figure2_Levo_Hydrographs.png",
  final_fig,
  width = 9,
  height = 9,
  units = "in",
  dpi = 600
)
```


```{r}
plot_data <- DailyFlow %>%
  left_join(
    ConcData %>%
      mutate(Date = as.Date(DateTime)),
    by = c("Watershed", "Date")
  ) %>%
  filter(Watershed == "SAN")

max_conc <- max(plot_data$Levo_mg_C_per_L, na.rm = TRUE)
max_flow <- max(plot_data$Q_mean, na.rm = TRUE)

scale_factor <- max_conc / max_flow
```

