---
title: "CoastalFires_Dataset_Wrangling"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
Load necessary Libraries
```{r}
library(fuzzyjoin)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
```
Set Working Directory
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/")
```

Load datafiles: levoglucosan data, discharge data, watershed and burn severity data, and previously published organic carbon and black carbon data
```{r}
LevoData <- readr::read_csv("CoastalFires_LevoglucosanData.csv")
BlackCarbonData <- readr::read_csv("CoastalFires_BlackCarbonData.csv")
WatershedData <- readr::read_csv("CoastalFires_WatershedData.csv")
DischargeData <- readr::read_csv("CoastalFires_DischargeData.csv")
``` 

Adjust the PublishedData data frame to remove unnecessary information
```{r}
#adjust the published data frame
#remove excess columns and print
BlackCarbonData <- BlackCarbonData %>% select(-TSS_mg_per_L, -POC_mgC_per_L, -PBC_mgC_per_L, -PBC_TSS_perc, -PBC_POC_perc, -pB6CA_nM,-pB5CA_nM, -pB6CA_B5CA)
print(BlackCarbonData)
```

Combine LevoData, WatershedData, and PublishedData Data Frames to one centralized dataframe: CoastalFires_CombinedData
```{r}
#combine data frames
Data <- left_join(LevoData, BlackCarbonData, by = "sample_id")
#exclude samples taken at lagoon dowstream of Scott sampling site
Data <- subset(Data, Data$sample_id != "SCT-7")
Data <- subset(Data, Data$sample_id != "SCT-8")
#add watershed data to table and remove duplicate locations
Data <- left_join(Data, WatershedData, by = "watershed")
CoastalFires_CombinedData <- Data %>% select(-Lat, -Long)
#print new combined data frame to be used for data analysis
print(CoastalFires_CombinedData) 
```

#Calculate levoclucosan ratios
```{r}
#calculate the ratio of levoglucosan to bulk DOC and DBC
CoastalFires_CombinedData$Levo_DOC_perc <- Data$levo_micro_gram_per_liter/1000/Data$DOC_mgC_per_L*100
CoastalFires_CombinedData$Levo_DBC <- Data$levo_micro_gram_per_liter/1000/Data$DBC_mgC_per_L
print(CoastalFires_CombinedData)
```

#Calculate fluxes and yields
```{r}
#discharge normalized levoglucosan concentrations
CoastalFires_CombinedData$Q_norm_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_micro_gram_per_liter / CoastalFires_CombinedData$Q_m_sq_per_sec)

#flux
CoastalFires_CombinedData$flux_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_micro_gram_per_liter * CoastalFires_CombinedData$Q_m_sq_per_sec)
CoastalFires_CombinedData$flux_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DOC_mgC_per_L * CoastalFires_CombinedData$Q_m_sq_per_sec)
CoastalFires_CombinedData$flux_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DBC_mgC_per_L * CoastalFires_CombinedData$Q_m_sq_per_sec)

#yield
CoastalFires_CombinedData$yield_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_levo / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DOC / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DBC / CoastalFires_CombinedData$Area_square_kilometer)

print(CoastalFires_CombinedData)
```
#Calculating a "Time Since Fire" metric
```{r}
#Ensuring that the Date column is correctly classified as a date rather than a character
CoastalFires_CombinedData$date <- as.Date(CoastalFires_CombinedData$date, format = "%d-%b-%y")

#Calculating the time since fire for all samples and ensuring it is a numeric value
fire_date <- as.Date("2020-08-16")
CoastalFires_CombinedData$time_since_fire_day <- as.numeric(CoastalFires_CombinedData$date - fire_date)
print(CoastalFires_CombinedData)
```

Re-arranged columns and change column names
```{r}
# Specify the desired column order
desired_order <- c("sample_id", "watershed", "date", "time","time_since_fire_day", "lat", "long", 
                   "Q_m_sq_per_sec", "flow_cat", "lagoon", "DOC_mgC_per_L", 
                   "levo_micro_gram_per_liter", "std_levo_micro_gram_per_liter", 
                   "DBC_mgC_per_L", "DBC_DOC_perc", "dB6CA_nM", "dB5CA_nM", 
                   "dB6CA_B5CA", "Levo_DOC_perc", "Levo_DBC", "Q_norm_levo", "flux_levo", 
                   "flux_DOC", "flux_DBC", "yield_levo", "yield_DOC", 
                   "yield_DBC", "a254_per_m", "SUVA254_L_per_mgC_per_m", 
                   "Area_square_kilometer", "Average_slope_percent", 
                   "Average_precipitation_milli_meter", "Burn_Extent_percent", 
                   "Area_low_severity_percent", "Area_low_to_mod_severity_percent", 
                   "Area_mod_to_high_severity_percent", "Forest_percent", 
                   "Shrub_percent", "Developed_percent", 
                   "Sandstone_and_sedimentary_percent", "Igneous_percent", 
                   "Suficial_Sediments_percent", "Schist_percent")

# Rearrange the columns
CoastalFires_CombinedData <- CoastalFires_CombinedData[, desired_order]
```

Create a csv file of the combined dataset:CoastalFires_CombinedData
```{r}
write.csv(CoastalFires_CombinedData, file = "CoastalFires_CombinedData.csv")
```