---
title: "CoastalFires_FigureCreation"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
#load necessary libraries
```{r}
library(ggplot2)
library(grid)
library(tidyr)
library(dplyr)
library(lubridate)
```

```{r}
# Calculate average value for each date
average_values <- CombinedData %>%
  group_by(time_since_fire_day) %>%
  summarize(average_values = mean(DBC_DOC_perc, na.rm = TRUE))

print(average_values)

```

#set working directory
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires")
```

#load dataset
```{r}
CombinedData <- readr::read_csv("CoastalFires_CombinedData.csv")
DischargeData <- readr::read_csv("CoastalFires_DischargeData.csv")
```
Print Column names
```{r}
col_names <- colnames(DischargeData)

# Print column names
print(col_names)
```
Make a new column in Combined data for Date_Time
```{r}
# Combine date and time into a new column with the desired format
CombinedData$datetime <- format(
  as.POSIXct(
    paste(CombinedData$date, CombinedData$time),        # Combine date and time
    format = "%m/%d/%Y %H:%M"           # Input format
  ),
  format = "%m/%d/%Y %H:%M:%S"          # Output format
)
```

Standardize DateTime formats for each watershed
```{r}
DischargeData$SAN_DateTime <- mdy_hm(DischargeData$SAN_DateTime)
DischargeData$PESC_DateTime <- mdy_hm(DischargeData$PESC_DateTime)
DischargeData$LAG_DateTime <- mdy_hm(DischargeData$LAG_DateTime)
DischargeData$MAJ_DateTime <- mdy_hm(DischargeData$MAJ_DateTime)
DischargeData$SCT_DateTime <- mdy(DischargeData$SCT_DateTime)  # Only date, no time
#some of the Scott Creek DateTime values will fail to parse. This is because the second water year was recorded using different DateTime formats. However, as this manuscript is only examining the first water year (Oct 2020-May 2021), this loss of discharge values do not impact analyses. Figures will be made only utilizing points prior to May 4th, 2021. 
```
Ensure all DateTime columns are in "POSIXct" rather than character classes
```{r}
CombinedData$datetime <- as.POSIXct(CombinedData$datetime, format = "%m/%d/%Y %H:%M")
DischargeData$LAG_DateTime <- as.POSIXct(DischargeData$LAG_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$MAJ_DateTime <- as.POSIXct(DischargeData$MAJ_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$PESC_DateTime <- as.POSIXct(DischargeData$PESC_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$SAN_DateTime <- as.POSIXct(DischargeData$SAN_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$SCT_DateTime <- as.POSIXct(DischargeData$SCT_DateTime, format = "%m/%d/%Y %H:%M")

class(CombinedData$datetime)
class(DischargeData$LAG_DateTime)
class(DischargeData$MAJ_DateTime)
class(DischargeData$PESC_DateTime)
class(DischargeData$SAN_DateTime)
class(DischargeData$SCT_DateTime)
```


Figure 2: The discharge (blue lines) and concentration of levoglucosan (yellow to dark red markers) from (A) the San Lorenzo River (B) Pescadero Creek (C) Majors Creek, (D) Laguna Creek, and (E) Scott Creek. The CZU Lightning Complex Fires (indicated by fire symbol) burned from August 16th until September 21st.
```{r}
# Log-transform the discharge data
DischargeData <- DischargeData$PESC_Flow_cms_log10 = log10(DischargeData$PESC_Flow_cms)

# Calculate the scaling factor
primary_max <- 70  # Maximum of the primary axis
secondary_max <- max(DischargeData$PESC_Flow_cms_log10, na.rm = TRUE)  # Log-transformed max of secondary axis
scaling_factor <- primary_max / secondary_max

# Plot
ggplot() +
  # Levoglucosan data (primary y-axis)
  geom_point(data = CombinedData %>% filter(watershed == "Pescadero Creek"), 
             aes(x = datetime, y = levo_micro_gram_per_liter), 
             color = "blue", alpha = 0.7, size = 2) +
  geom_line(data = CombinedData %>% filter(watershed == "Pescadero Creek"), 
            aes(x = datetime, y = levo_micro_gram_per_liter), 
            color = "blue", linewidth = 1) +
  # Discharge data (secondary y-axis with log-transformed and scaled values)
  geom_line(data = DischargeData, 
            aes(x = PESC_DateTime, y = PESC_Flow_cms_log10 * scaling_factor), 
            color = "red", linewidth = 1) +
  # Primary axis
  scale_y_continuous(
    name = "Levoglucosan (ppb)", 
    limits = c(0, 70), 
    breaks = seq(0, 70, by = 10), 
    sec.axis = sec_axis(
      ~ . / scaling_factor,  # Reverse the scaling for the secondary axis
      name = "Discharge (m³/s, log scale)"  # Let ggplot handle breaks and labels dynamically
    )
  ) +
  # Common x-axis
  scale_x_datetime(name = "Date") +
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red"),
    panel.grid.minor = element_blank(),
    axis.text.y.right = element_text(color = "red")
  ) +
  labs(title = "Discharge (Log Scale) and Levoglucosan Concentration Over Time")


```

```{r}
# Create Fig. 2A - Pescadero Creek

# Rescale the levoglucosan data and discharge data to the same range
# Calculate the scaling factor
primary_max <- 70  # Maximum of primary axis
PESC_secondary_max <- max(DischargeData$PESC_Flow_cms, na.rm = TRUE)  # Max of secondary axis data
PESC_scaling_factor <- primary_max / PESC_secondary_max

# Plot
ggplot() +
  # Levoglucosan data (primary y-axis)
  geom_point(data = CombinedData %>% filter(watershed == "Pescadero Creek"), 
             aes(x = datetime, y = levo_micro_gram_per_liter), 
             color = "blue", alpha = 0.7, size = 2) +
  geom_line(data = CombinedData %>% filter(watershed == "Pescadero Creek"), 
            aes(x = datetime, y = levo_micro_gram_per_liter), 
            color = "blue", linewidth = 1) +
  # Discharge data (secondary y-axis with rescaled values and log transformation)
  geom_line(data = DischargeData, 
            aes(x = PESC_DateTime, y = log10(rescaled_discharge + 1)),  # Apply log10 with +1 to avoid log(0)
            color = "red", linewidth = 1) +
  # Secondary axis for discharge (log scale)
  scale_y_continuous(
    name = "Levoglucosan (ppb)", 
    sec.axis = sec_axis(
      trans = ~10^(.) - 1,  # Reverse log scaling on the secondary axis
      name = "Discharge (m³/s, log scale)"
    )
  ) +
  # Common x-axis
  scale_x_datetime(name = "Date") +
  # Customize appearance
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red"),
    panel.grid.minor = element_blank()
  ) +
  labs(title = "Discharge (Log Scale) and Levoglucosan Concentration Over Time")

  
```

```{r}
ggplot(DischargeData, aes(x = PESC_DateTime, y = PESC_Flow_cms)) +
  geom_line(color = "red") +
  scale_x_datetime(name = "Date") +
  scale_y_continuous(name = "Discharge (m³/s)") +
  theme_minimal() +
  labs(title = "Discharge Over Time")
```
