---
title: "Coastal_Fires_DataAnalysis"
author: "RB and AC"
date: "2025-04-13"
output: html_document
---
To utilize this script it is necessary to have run "CosatalFires_DataFrameWrangling.Rmd" ending with the creation of "CoastalFires_CombinedData.csv". If you have not completed this, please do so before continuing. 

Here, this script is utilized for data analysis and preliminary figure creation.

Install necessary packages (run once if necessary)
```{r}
#install.packages("gt")
#install.packages("Cairo")
#install.packages("ggrepel")
#install.packages("vegan")
#install.packages("tidyverse")
#install.packages("vegan")
#install.packages("tidyverse")
```

Load necessary libraries
```{r}
library(gt)
library(Cairo)
library(ggrepel)
library(vegan)
library(tidyverse)
library(dplyr)
library(purrr)
library(cowplot)
```

Read data file
```{r}
CombinedData <- readr::read_csv(file.path("data/CoastalFires_CombinedData.csv"))
``` 

Exclude samples that were taken at a different location. 
```{r}
#Scott Creek flooded during the events of January 27-28th, 2021. Samples were collected at a downstream lagoon site. However, due to this the data collected does not accurately reflect that at the sample site. As such authors of the associated manuscript have excluded it from all data analysis and figure creation.
CombinedData <- CombinedData %>%
  filter(Location_Change != "Y")
```

#Calculating a "discharge-normalized" levoglucosan concentration
```{r}
CombinedData$Q_norm_levo <- CombinedData$Levo_micro_mol_C_per_L / CombinedData$Discharge_cms #units are in uM-C s per m^(-3)
```

Range, Average, Median
```{r}
# List of variables to summarize
vars_to_summarize <- c("DOC_micro_mol_C_per_L", "Levo_micro_mol_C_per_L", "B6CA_and_B5CA_micro_mol_C_per_L", 
                       "Levo_DOC_percent", "BPCA_DOC_percent", "SUVA254_L_per_mgC_per_m", "a254_per_m")

# Function to calculate summary stats for one variable
summary_stats <- function(var) {
  x <- CombinedData[[var]]
  x <- x[!is.na(x)]  # remove NAs
  data.frame(
    Variable = var,
    Min = min(x),
    Max = max(x),
    Mean = mean(x),
    SD = sd(x),
    Median = median(x)
  )
}

# Apply function to all variables
summary_table <- do.call(rbind, lapply(vars_to_summarize, summary_stats))

# View the summary
print(summary_table)

```

#Shapiro-Wilk: Test data for normality
```{r}
#are all variables normally distributed?
#filter to only numeric data
numeric_columns <- sapply(CombinedData, is.numeric)
Data_numeric <- CombinedData[, numeric_columns]
Data_numeric <- Data_numeric %>% select(-Latitude, -Longitude)

# Apply Shapiro-Wilk test and create a readable summary
shapiro_results <- sapply(Data_numeric, function(x) {
  test <- shapiro.test(x)
  return(c(p_value = test$p.value, W = test$statistic))
})

# Convert results to a data frame
shapiro_df <- as.data.frame(t(shapiro_results))
shapiro_df$Variable <- rownames(shapiro_df)
rownames(shapiro_df) <- NULL

# Print readable results
print(shapiro_df)
```

#Kruskal Wallis: Test for differences between watersheds
```{r}
# Ensure Watershed is a factor
CombinedData$Watershed <- as.factor(CombinedData$Watershed)

# Filter to numeric data (excluding lat/long again)
numeric_columns <- sapply(CombinedData, is.numeric)
Data_numeric <- CombinedData[, numeric_columns]
Data_numeric <- Data_numeric %>% select(-Latitude, -Longitude)

# Run Kruskal-Wallis test for each numeric variable by Watershed
kruskal_results <- sapply(names(Data_numeric), function(var) {
  formula <- as.formula(paste(var, "~ Watershed"))
  test <- kruskal.test(formula, data = CombinedData)
  return(c(p_value = test$p.value, chi_squared = test$statistic))
})

# Convert to readable data frame
kruskal_df <- as.data.frame(t(kruskal_results))
kruskal_df$Variable <- rownames(kruskal_df)
rownames(kruskal_df) <- NULL

# Print the results
print(kruskal_df)
```

#Wilcoxon: Which watersheds have significantly different concentrations?
```{r}
#According to the above Kruskal Wallis tests, the following variables are significantly different between watersheds: Q_norm_levo, BPCA_DOC_percent, and dB6CA_B5CA
# Make sure Watershed is a factor
CombinedData$Watershed <- as.factor(CombinedData$Watershed)

# Variables to test
vars_to_test <- c("Q_norm_levo", "BPCA_DOC_percent", "dB6CA_B5CA")

# Get all unique pairwise combinations of watersheds
watershed_pairs <- combn(levels(CombinedData$Watershed), 2, simplify = FALSE)

# Initialize an empty list to store results
wilcox_results <- list()

# Loop through each variable and watershed pair
for (var in vars_to_test) {
  for (pair in watershed_pairs) {
    
    # Extract and unlist group data
    group1 <- CombinedData[CombinedData$Watershed == pair[1], var]
    group2 <- CombinedData[CombinedData$Watershed == pair[2], var]
    
    # Unlist and convert to numeric, removing NA values
    group1 <- as.numeric(unlist(group1))
    group2 <- as.numeric(unlist(group2))
    
    group1 <- group1[!is.na(group1)]
    group2 <- group2[!is.na(group2)]
    
    # Only test if both groups have data
    if (length(group1) > 0 & length(group2) > 0) {
      test <- wilcox.test(group1, group2)
      
      wilcox_results[[length(wilcox_results) + 1]] <- data.frame(
        Variable = var,
        Watershed_1 = pair[1],
        Watershed_2 = pair[2],
        p_value = test$p.value,
        W_statistic = test$statistic
      )
    }
  }
}

# Combine results into a single data frame
wilcox_df <- do.call(rbind, wilcox_results)

# Print
print(wilcox_df)
```

Spearman's Rank: Time since fire
```{r}
# Variables to test against time_since_fire
vars_to_test_time_since_fire <- c("DOC_micro_mol_C_per_L", "Levo_micro_mol_C_per_L", "B6CA_and_B5CA_micro_mol_C_per_L", "Levo_DOC_percent", "BPCA_DOC_percent", "Q_norm_levo", "SUVA254_L_per_mgC_per_m", "a254_per_m")

# Initialize a list to store results
spearman_results_time_since_fire <- lapply(vars_to_test_time_since_fire, function(var) {
  
  # Extract x and y, remove rows with NA
  df <- CombinedData[, c("time_since_fire_day", var)]
  df <- na.omit(df)
  
  # Run Spearman correlation test
  test <- cor.test(df$time_since_fire_day, df[[var]], method = "spearman")
  
  # Return results as a row in a data frame
  data.frame(
    Variable = var,
    Spearman_rho = test$estimate,
    p_value = test$p.value,
    stringsAsFactors = FALSE
  )
})

# Combine into a single data frame
spearman_df_time_since_fire <- do.call(rbind, spearman_results_time_since_fire)

# Print results
print(spearman_df_time_since_fire)
```

Figure 3: time-since-fire relationships
```{r}
#step 0: define order and style of watersheds -----
#Watershed colors and shapes
CombinedData <- CombinedData %>%
  mutate(
    Watershed = recode(
      Watershed,
      "San Lorenzo River" = "SAN",
      "Pescadero Creek" = "PESC",
      "Majors Creek" = "MAJ",
      "Laguna Creek" = "LAG",
      "Scott Creek" = "SCT"
    )
  )
ws_cols <- c(
  SAN  = "#FBF9A2",
  PESC = "#FFCF55",
  MAJ  = "#FF7410",
  LAG  = "#EA4412",
  SCT  = "#A00000"
)

ws_shapes <- c(
  SAN  = 23,
  PESC = 24,
  MAJ  = 22,
  LAG  = 21,
  SCT  = 25
)

#to determine legend order
CombinedData <- CombinedData %>%
  mutate(
    Watershed = factor(
      Watershed,
      levels = c("SAN", "PESC", "MAJ", "LAG", "SCT")
    )
  )

#step 1: make a function for plotting figure 3 ----- 
make_trend_plot <- function(y_var, y_lab, panel = NULL, y_limits = NULL, show_legend = FALSE) {

  p <- ggplot(
    CombinedData,
    aes(
      x = time_since_fire_day,
      y = {{ y_var }},
      fill = Watershed,
      shape = Watershed
    )
  ) +

    geom_smooth(
      data = CombinedData,
      mapping = aes(
        x = time_since_fire_day,
        y = {{ y_var }},
        group = 1
      ),
      method = "gam",
      formula = y ~ s(x, k = 5),
      color = "black",
      fill = "grey60",
      alpha = 0.25,
      linewidth = 1,
      se = TRUE,
      inherit.aes = FALSE
    ) +

    geom_point(
      size = 4,
      color = "black",
      stroke = 0.8
    ) +

    scale_fill_manual(values = ws_cols) +
    scale_shape_manual(values = ws_shapes) +

    scale_y_continuous(
      limits = y_limits
    ) +
  
  labs(
      x = "Time since fire (days)",
      y = y_lab
    ) +

    theme_classic(base_size = 13) +

    theme(
      legend.position = if(show_legend) "right" else "none",
      axis.text = element_text(color = "black")
    )

  if (!is.null(panel)) {
    p <- p +
      annotate(
        "text",
        x = -Inf,
        y = Inf,
        label = panel,
        fontface = "bold",
        hjust = -0.3,
        vjust = 1.3,
        size = 6
      )
  }

  p
}

#step 2: make each sub-plot ----- 
p1 <- make_trend_plot(
  Levo_micro_mol_C_per_L,
  expression(Levoglucosan~(mu*M-C)),
  panel = "A)",
  y_limits = c(0, 3)
)

p2 <- make_trend_plot(
  B6CA_and_B5CA_micro_mol_C_per_L,
  expression(BPCA~(mu*M-C)),
  panel = "B)",
  y_limits = c(0, 12)
)

p3 <- make_trend_plot(
  Levo_DOC_percent,
  "Levo:DOC (%)",
  panel = "C)",
  y_limits = c(0, 1.5)
)

p4 <- make_trend_plot(
  BPCA_DOC_percent,
  "BPCA:DOC (%)",
  panel = "D)",
  y_limits = c(0, 1.5)
)

legend <- get_legend(
  make_trend_plot(
    Levo_micro_mol_C_per_L,
    expression(Levoglucosan~(mu*mol~C~L^{-1})),
    show_legend = TRUE
  )
) #just to get the legend for putting them all together

#step 3: combine sub-plots -----
#Combine plots
plots <- plot_grid(
  p1, p2,
  p3, p4,
  ncol = 2,
  align = "hv",
  axis = "tblr"
)

#add legend
final_plot <- plot_grid(
  plots,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.15)
)

#step 4: save as tiff -----
ggsave(
  "figures/Fig3_TimeSinceFire.tiff",
  final_plot,
  width = 8,
  height = 7,
  units = "in",
  dpi = 600,
  compression = "lzw",
  bg = "white"
)
```

Spearman's Rank: Discharge
```{r}
# Variables to test against discharge
vars_to_test_Q <- c("DOC_micro_mol_C_per_L", "Levo_micro_mol_C_per_L", "B6CA_and_B5CA_micro_mol_C_per_L", "Levo_DOC_percent", "BPCA_DOC_percent", "SUVA254_L_per_mgC_per_m", "a254_per_m")

# Initialize a list to store results
spearman_results_Q <- lapply(vars_to_test_Q, function(var) {
  
  # Extract x and y, remove rows with NA
  df <- CombinedData[, c("Discharge_cms", var)]
  df <- na.omit(df)
  
  # Run Spearman correlation test
  test <- cor.test(df$Discharge_cms, df[[var]], method = "spearman")
  
  # Return results as a row in a data frame
  data.frame(
    Variable = var,
    Spearman_rho = test$estimate,
    p_value = test$p.value,
    stringsAsFactors = FALSE
  )
})

# Combine into a single data frame
spearman_df_Q <- do.call(rbind, spearman_results_Q)

# Print results
print(spearman_df_Q)
```

Spearman's Rank: Levo versus watershed and fire characteristics 
```{r}
# Variables to test against Levo_micro_mol_C_per_L
vars_to_test_levo_conc <- c("Area_square_kilometer", "Average_slope_percent", "Average_precipitation_milli_meter", "Burn_Extent_percent", "Area_low_severity_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent", "Forest_percent", "Shrub_percent", "Developed_percent", "Sandstone_and_sedimentary_percent", "Igneous_percent", "Suficial_Sediments_percent", "Schist_percent")

# Initialize a list to store results
spearman_results_levo_conc <- lapply(vars_to_test_levo_conc, function(var) {
  
  # Extract x and y, remove rows with NA
  df <- CombinedData[, c("Levo_micro_mol_C_per_L", var)]
  df <- na.omit(df)
  
  # Run Spearman correlation test
  test <- cor.test(df$Levo_micro_mol_C_per_L, df[[var]], method = "spearman")
  
  # Return results as a row in a data frame
  data.frame(
    Variable = var,
    Spearman_rho = test$estimate,
    p_value = test$p.value,
    stringsAsFactors = FALSE
  )
})

# Combine into a single data frame
spearman_df_levo_conc <- do.call(rbind, spearman_results_levo_conc)

# Print results
print(spearman_df_levo_conc)
```

Spearman's Rank: Levo-DOC versus watershed and fire characteristics 
```{r}
# Variables to test against Levo_DOC_percent
vars_to_test_levo_DOC <- c("Area_square_kilometer", "Average_slope_percent", "Average_precipitation_milli_meter", "Burn_Extent_percent", "Area_low_severity_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent", "Forest_percent", "Shrub_percent", "Developed_percent", "Sandstone_and_sedimentary_percent", "Igneous_percent", "Suficial_Sediments_percent", "Schist_percent")

# Initialize a list to store results
spearman_results_levo_DOC <- lapply(vars_to_test_levo_DOC, function(var) {
  
  # Extract x and y, remove rows with NA
  df <- CombinedData[, c("Levo_DOC_percent", var)]
  df <- na.omit(df)
  
  # Run Spearman correlation test
  test <- cor.test(df$Levo_DOC_percent, df[[var]], method = "spearman")
  
  # Return results as a row in a data frame
  data.frame(
    Variable = var,
    Spearman_rho = test$estimate,
    p_value = test$p.value,
    stringsAsFactors = FALSE
  )
})

# Combine into a single data frame
spearman_df_levo_DOC <- do.call(rbind, spearman_results_levo_DOC)

# Print results
print(spearman_df_levo_DOC)
```

NMDS
Prepare data for NMDS
```{r}
#delete rows with NA
CombinedData_clean <- na.omit(CombinedData)

#Isolate PyC concentrations and relevant ratios 
analyte_data_PyC_conc_rat <- CombinedData_clean[, c("DOC_micro_mol_C_per_L", "Levo_micro_mol_C_per_L", "B6CA_and_B5CA_micro_mol_C_per_L", "Levo_DOC_percent", "BPCA_DOC_percent")]
```

Run NMDS
```{r}
nmds_result <- metaMDS(analyte_data_PyC_conc_rat, k = 2, trymax = 100)
print(nmds_result)
```

Plot NMDS Results - Full NMDS 
```{r}
# Extract NMDS scores for plotting
nmds_scores <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_species<-as.data.frame(scores(nmds_result, display="species"))
nmds_scores$Site <- CombinedData_clean$Watershed  # Add Site info for coloring
nmds_scores$time <- CombinedData_clean$time_since_fire_day #this is for coloring the data points by time
nmds_scores$watershed <- CombinedData_clean$Watershed #this is for the point shape by watershed

# Environmental variables 
env_variables <- CombinedData_clean[c("time_since_fire_day", "Area_low_severity_percent", "Discharge_cms", "Area_square_kilometer", "Average_slope_percent", "Burn_Extent_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent")]

#assign environmental variables to df
envfit_result <- envfit(nmds_result, env_variables, permutations = 999)

#have the environmental variables assigned to arrows for plotting
en_coord_cont = as.data.frame(scores(envfit_result, "vectors")) * ordiArrowMul(envfit_result)

# Shorten labels for plotting species and vectors
nmds_species$ShortLabel <- c("DOC", "Levo","BPCA","Levo:DOC", "BPCA:DOC")  # Replace with your short labels

# Change labels for plotting vectors. Also wrapping text for improving clarity of the plot
en_coord_cont$WrappedLabel <- c(
  "Time Since\nFire", #Time Since Fire
  "Low Severity\nBurn Extent", #Low Seveirty Burn Extent
  "Discharge", #Discharge
  "Watershed\nArea",
  "Slope",
  "Burn Extent",
  "Low-Mod\nBurn Extent",
  "Mod-High\nBurn Extent" #Moderate to High Burn Extent
)

#NMDS Visual Plot
ggplot(nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  
  #Sites
  geom_point(data = nmds_scores, aes(size = 100, color = time, fill = time, shape = watershed)) +
     scale_shape_manual(values = c(21, 22, 24, 23, 25))+
  scale_fill_gradient(low="#00FFFF", high="navy")+
  scale_color_gradient(low="#00FFFF", high="navy")+
 
  #Species
  geom_point(data = nmds_species, aes(x = NMDS1, y = NMDS2), shape = 8, stroke = 2, size = 3, alpha = 1, colour = "hotpink") +
  
  geom_text_repel(
  data = nmds_species,
  aes(x = NMDS1, y = NMDS2, label = ShortLabel),
  colour = "black", fontface = "bold", size = 3,
  box.padding = 0.3,
  point.padding = 0.25,
  segment.color = NA,
  max.overlaps = Inf,
  force = 1
) +

geom_text_repel(
  data = en_coord_cont,
  aes(x = NMDS1 * 0.75, y = NMDS2 * 0.75, label = WrappedLabel),
  size = 3,
  box.padding = 0.5,
  point.padding = 0.25,
  segment.color = NA,
  max.overlaps = Inf,
  force = 1
) +
  
  #vectors
  geom_segment(aes(x = 0, y = 0, xend = NMDS1*0.75, yend = NMDS2*0.75), data = en_coord_cont, linewidth =1, alpha = 1, colour = "black", arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +
  
  #chart title
  labs(x = "NMDS1",
       y = "NMDS2") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(size = 0.5),  # Add axis lines
    axis.ticks = element_line(size = 0.5),  # Add axis ticks
    axis.ticks.length = unit(3, "pt"),  # Make ticks outward
    axis.line.x = element_line(size = 0.5),
    axis.line.y = element_line(size = 0.5)
  )

#Determine which vectors are statistically significant
envfit_result
```

Print a labeled NMDS figure
```{r}
# Begin saving to file to output folder
CairoPNG(filename = file.path("NMDS_labeled.png"),
         width = 7, height = 4, units = "in", res = 300)

#NMDS Visual Plot
ggplot(nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  
  #Sites
  geom_point(data = nmds_scores, aes(size = 100, color = time, fill = time, shape = watershed)) +
     scale_shape_manual(values = c(21, 22, 24, 23, 25))+
  scale_fill_gradient(low="#00FFFF", high="navy")+
  scale_color_gradient(low="#00FFFF", high="navy")+
 
  #Species
  geom_point(data = nmds_species, aes(x = NMDS1, y = NMDS2), shape = 8, stroke = 2, size = 3, alpha = 1, colour = "hotpink") +
  
  geom_text_repel(
  data = nmds_species,
  aes(x = NMDS1, y = NMDS2, label = ShortLabel),
  colour = "black", fontface = "bold", size = 3,
  box.padding = 0.3,
  point.padding = 0.25,
  segment.color = NA,
  max.overlaps = Inf,
  force = 1
) +

geom_text_repel(
  data = en_coord_cont,
  aes(x = NMDS1 * 0.75, y = NMDS2 * 0.75, label = WrappedLabel),
  size = 3,
  box.padding = 0.5,
  point.padding = 0.25,
  segment.color = NA,
  max.overlaps = Inf,
  force = 1
) +
  
  #vectors
  geom_segment(aes(x = 0, y = 0, xend = NMDS1*0.75, yend = NMDS2*0.75), data = en_coord_cont, linewidth =1, alpha = 1, colour = "black", arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +
  
  #chart title
  labs(x = "NMDS1",
       y = "NMDS2") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(size = 0.5),  # Add axis lines
    axis.ticks = element_line(size = 0.5),  # Add axis ticks
    axis.ticks.length = unit(3, "pt"),  # Make ticks outward
    axis.line.x = element_line(size = 0.5),
    axis.line.y = element_line(size = 0.5)
  )

# Finish saving the NMDS figure file
dev.off()
```

Print Unlabeled NMDS for figure making
```{r}
# Begin saving to file to output folder
CairoPNG(filename = file.path("NMDS_unlabeled.png"), 
         width = 7, height = 4, units = "in", res = 300)

# Draw the plot (your label-free version)
ggplot(nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  geom_point(data = nmds_scores, aes(size = 100, color = time, fill = time, shape = watershed)) +
  scale_shape_manual(values = c(21, 22, 24, 23, 25)) +
  scale_fill_gradient(low = "#00FFFF", high = "navy") +
  scale_color_gradient(low = "#00FFFF", high = "navy") +
  geom_point(data = nmds_species, aes(x = NMDS1, y = NMDS2), 
             shape = 8, stroke = 2, size = 3, alpha = 1, colour = "hotpink") +
  geom_segment(data = en_coord_cont, 
               aes(x = 0, y = 0, xend = NMDS1 * 0.75, yend = NMDS2 * 0.75),
               linewidth = 1, alpha = 1, colour = "black",
               arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +
  labs(x = "NMDS1", y = "NMDS2") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(size = 0.5),
    axis.ticks = element_line(size = 0.5),
    axis.ticks.length = unit(3, "pt"),
    axis.line.x = element_line(size = 0.5),
    axis.line.y = element_line(size = 0.5)
  )

# Finish saving the NMDS figure file
dev.off()
```

#Calculate fluxes of levoglucosan in the first year post-fire
```{r}
#Convert discharge values from cubic meters per second (cms) to liters per second
CombinedData$Discharge_liter_per_sec <- CombinedData$Discharge_cms * 1000

#Calculate sample-specific flux values and convert to units of kg per year
CombinedData$Levo_flux_kilo_gram_C_per_year <- CombinedData$Levo_micro_mol_C_per_L * CombinedData$Discharge_liter_per_sec * 31536000 * 12 / 1000000000

#Average flux samples by watershed
avg_flux_per_watershed <- CombinedData %>%
  group_by(Watershed) %>%
  summarise(
    mean_flux = mean(Levo_flux_kilo_gram_C_per_year, na.rm = TRUE),
    sd_flux = sd(Levo_flux_kilo_gram_C_per_year, na.rm = TRUE),
    n = sum(!is.na(Levo_flux_kilo_gram_C_per_year))
  )

#Calculate sum of fluxes from each watershed
total_flux_all_watersheds <- sum(avg_flux_per_watershed$mean_flux)
total_flux_se <- sqrt(sum((avg_flux_per_watershed$sd_flux)^2 / avg_flux_per_watershed$n))

#Print value with standard deviation carried over
print(total_flux_all_watersheds) #flux
print(total_flux_se) #standard error
```
