---
title: "Figure 1: CZU Fires Map"
author: "rbb"
date: "2026-06-03"
output: html_document
---
Before starting, make sure to read the README found on the GitHub repository. This script is utilized for creating a Map of the CZU lightning complex fire and sampling sites associated with the manuscript.

Install package if necessary
```{r}
# install.packages("arcgislayers")
# install.packages("sf")
# install.packages("tmap")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("terra")
# install.packages("tigris")
# install.packages("tidyterra")
# install.packages("ggspatial")
# install.packages("cowplot")
```

load package(s)
```{r}
library(arcgislayers)
library(sf)
library(tmap)
library(ggplot2)
library(dplyr)
library(terra)
library(tigris)
library(tidyterra)
library(ggspatial)
library(cowplot)
```

find and download shapefiles/rasters
```{r}
#California
  #from repo
    ca <- states(cb = TRUE) |>
      dplyr::filter(STUSPS == "CA")
    
#Streams
  #open file as GeoJSON
    streams <- st_read("data/GIS Files/Streams clipped to watersheds.gpkg") 
    
#Watershed Boundaries
  #open file as GeoJSON
    watershed_boundary <- st_read("data/GIS Files/CZU Watersheds.gpkg") 
    
#read in fire perimeter
fire_url <- paste0(
  "https://services1.arcgis.com/jJfZghspGKh8J9Jm/",
  "arcgis/rest/services/CZU_Fire_Perimeter/FeatureServer/3/query",
  "?where=1%3D1",
  "&outFields=*",
  "&returnGeometry=true",
  "&f=geojson"
)
burn_area <- vect(fire_url)


#Burn severity from MTBS, downloaded on May-26-2026. Filters - Region: California, Year: 2020
  burn_severity <- rast("data/GIS Files/mtbs-albers-dnbr/mtbs_ca3726212222320200816_10021177_20200724_20210711_dnbr.tif") #had to download from email
  
#Sites
  sites <- read.csv("data/GIS Files/Sample Locations.csv")
```

Assign and/or transform CRS's to UTM
```{r}
#crs UTM downloaded from mtbs
burn_utm <- terra::project(
  burn_severity,
  "EPSG:32610",
  method = "near"
)

#convert sites from WGS84 to UTM
sites_sf <- st_as_sf(
  sites,
  coords = c("Long", "Lat"),
  crs = 4326
)
sites_utm <- st_transform(sites_sf, 32610)

#streams is in Web Mercator, assign the correct CRS and transform
streams <- st_set_crs(streams, 3857)
streams_utm <- st_transform(streams, 32610)

#basic transformation
burn_area_utm <- terra::project(burn_area, "EPSG:32610")
watershed_utm <- st_transform(watershed_boundary, 32610)
ca_utm <- st_transform(ca, 32610)
```

Snip burn severity raster to burn extent
```{r}
# rename
burn_vect <- burn_area_utm

# Crop to the extent of the fire perimeter
burn_severity_crop <- crop(burn_utm, burn_vect)

# Mask to the actual fire perimeter
burn_severity_crop <- mask(burn_severity_crop, burn_vect)

#convert to dNBR
burn_dnbr <- burn_severity_crop / 1000
```

Main map
```{r}
#set extent to be watersheds with some padding around it
bb <- st_bbox(watershed_utm)

left_pad  <- (bb$xmax - bb$xmin) * 0.2
right_pad <- (bb$xmax - bb$xmin) * 0.5
bottom_pad <- (bb$ymax - bb$ymin) * 0.2
top_pad    <- (bb$ymax - bb$ymin) * 0.10

main_map <- ggplot() +

  geom_sf(data = ca_utm, fill = "white", color = NA) +
  geom_spatraster(data = burn_dnbr) +
  scale_fill_gradientn(
    name = "dNBR",
    colors = c(
      "#2166AC",  # negative
      "#F7F7F7",  # near zero
      "#FEE08B",  # low
      "#F46D43",  # moderate
      "#A50026"   # high
    ),
    na.value = "transparent",
    guide = guide_colorbar(
      frame.colour = "black",
      frame.linewidth = 0.25,
      ticks = TRUE,
      ticks.colour = "black"
    )
  ) +


  geom_sf(data = ca_utm, fill = NA, color = "black", linewidth = 0.75) +

  geom_sf(data = watershed_utm, fill = NA, color = "black", linewidth = 0.5) +

  geom_sf(data = streams_utm, color = "blue", linewidth = 0.3) +

  geom_sf(data = sites_utm, shape = 21, fill = "black", color = "black", size = 2.5) +

  coord_sf(
    xlim = c(
      bb$xmin - left_pad,
      bb$xmax + right_pad
    ),
    ylim = c(
      bb$ymin - bottom_pad,
      bb$ymax + top_pad
    ),
    expand = FALSE
  ) +

  annotation_scale(location = "bl", width_hint = 0.3) +

  annotation_north_arrow(
    location = "tl",
    style = north_arrow_minimal(),
    height = unit(1, "cm"),
    width = unit(1, "cm")
  ) +

  theme_bw() +
  theme(
      legend.position = c(0.89, 0.45),
      legend.background = element_rect(
        fill = NA
        ),
      legend.key.height = unit(0.3, "cm"),
      legend.key.width = unit(0.3, "cm"),
      legend.text = element_text(size = 7),
      legend.title = element_text(size = 8)
      )
```

Inset map
```{r}
#make black box around study site
study_box <- st_as_sfc(st_bbox(watershed_boundary))
western_states <- states(cb = TRUE) |> filter(STUSPS %in% c("CA","OR","WA","NV","ID","MT","WY","UT","AZ","CO","NM"))

#inset map
inset_map <- ggplot() +

  geom_sf(data = western_states,
          fill = "#EFEFEF",
          color = "black",
          linewidth = 0.3) +

  geom_sf(data = study_box,
          fill = NA,
          color = "red",
          linewidth = 0.4) +

  coord_sf() +

  theme_void() +
  
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )
```

Final Map
```{r}
final_map <- ggdraw() +
  draw_plot(main_map) +
  draw_plot(
    inset_map,
    x = 0.66,
    y = 0.61,
    width = 0.35,
    height = 0.35
  ) +
  theme(
  panel.background = element_rect(
    fill = "white",
    color = "black",
    linewidth = 0.8
  )
)
```

Print Final Map
```{r}
ggsave(
  filename = "figures/CZU_fire_map.tif",
  plot = final_map,
  width = 5,
  height = 3.5,
  units = "in",
  dpi = 600,
  compression = "lzw"
)
```
