---
title: "CoastalFires_Dataset_Wrangling"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
Load necessary Libraries
```{r}
library(fuzzyjoin)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
```
Set Working Directory
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/Input Files")
```

Load datafiles: levoglucosan data, discharge data, watershed and burn severity data, and previously published organic carbon and black carbon data
```{r}
LevoData <- readr::read_csv("CoastalFires_LevoglucosanData.csv")
BlackCarbonData <- readr::read_csv("CoastalFires_BlackCarbonData.csv")
WatershedData <- readr::read_csv("CoastalFires_WatershedData.csv")
DischargeData <- readr::read_csv("CoastalFires_DischargeData.csv")
``` 

Adjust LevoData by averaging replicates and removing unnecessary information
```{r}
#delete excess columns
LevoData <- LevoData %>% select(-relative_area_levo, -relative_amount_levo, -levoglucosan_area, -C13_levoglucosan_area, -c13_levo_ppb)
#rearrange columns to get sample_id and replicate number
library(stringr)
LevoData$sample_id <- str_extract(LevoData$header, "(?<=-).*")
LevoData$replicate <- str_extract(LevoData$header, "(?<=R)[0-2]+")
#remove non-data (i.e., standards)
LevoData <- LevoData %>% filter(!is.na(replicate))
#remove any naming errors
LevoData[LevoData == 'SAN-7DUP'] <- 'SAN-7'
#average replicates
LevoData <- LevoData %>% group_by(sample_id) %>% summarize(levo_ppb=mean(levo_ppb))
print(LevoData)
```

Adjust the PublishedData data frame to remove unnecessary information
```{r}
#adjust the published data frame
#remove excess columns and print
PublishedData <- PublishedData %>% select(-TSS_mg_per_L, -POC_mgC_per_L, -PBC_mgC_per_L, -PBC_TSS_perc, -PBC_POC_perc, -pB6CA_nM,-pB5CA_nM, -pB6CA_B5CA)
print(PublishedData)
```

Combine LevoData, WatershedData, and PublishedData Data Frames to one centralized dataframe: CoastalFires_CombinedData
```{r}
#combine data frames
Data <- left_join(LevoData, PublishedData, by = "sample_id")
#rearrange column order to move levo_ppb after DOC
Data <- Data %>% select(sample_id, watershed, date, time, lat, long, Q_m_sq_per_sec, flow_cat, lagoon, DOC_mgC_per_L, levo_ppb, DBC_mgC_per_L, everything())
#exclude samples taken at lagoon dowstream of Scott sampling site
Data <- subset(Data, Data$sample_id != "SCT-7")
Data <- subset(Data, Data$sample_id != "SCT-8")
#add watershed data to table and remove duplicate locations
Data <- left_join(Data, WatershedData, by = "watershed")
CoastalFires_CombinedData <- Data %>% select(-Lat, -Long)
#print new combined data frame to be used for data analysis
print(CoastalFires_CombinedData) 
```

#Calculate levoclucosan ratios
```{r}
#calculate the ratio of levoglucosan to bulk DOC and DBC
CoastalFires_CombinedData$Levo_DOC_perc <- Data$levo_ppb/1000/Data$DOC_mgC_per_L*100
CoastalFires_CombinedData$Levo_DBC <- Data$levo_ppb/1000/Data$DBC_mgC_per_L
print(CoastalFires_CombinedData)
```

#Calculate fluxes and yields
```{r}
#flux
CoastalFires_CombinedData$flux_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_ppb * CoastalFires_CombinedData$Q_m_sq_per_sec)
CoastalFires_CombinedData$flux_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DOC_mgC_per_L * CoastalFires_CombinedData$Q_m_sq_per_sec)
CoastalFires_CombinedData$flux_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DBC_mgC_per_L * CoastalFires_CombinedData$Q_m_sq_per_sec)

#yield
CoastalFires_CombinedData$yield_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_levo / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DOC / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DBC / CoastalFires_CombinedData$Area_square_kilometer)

print(CoastalFires_CombinedData)
```

Re-arranged columns and change column names
```{r}


```


Create a csv file of the combined dataset:CoastalFires_CombinedData
```{r}
write.csv(CoastalFires_CombinedData, file = "CoastalFires_CombinedData.csv")
```