---
title: "Coastal Fires Levoglucosan"
author: "RB"
date: "2024-09-13"
output: html_document
---
To utilize this script it is necessary to have run "CosatalFires_DataFrameWrangling.Rmd" ending with the creation of "CoastalFires_CombinedData.csv". If you have not completed this, please do so before continuing. 

Here, DESCRIBE WHAT THIS IS. 

Install necessary packages
```{r}
install.packages("ggrepel")
install.packages("vegan")
install.packages("tidyverse")
```

Load necessary libraries
```{r}
library(ggrepel)
library(vegan)
library(tidyverse)
```

Set Input and Output Directory, load data files, and rename them
```{r}
#Input_Files folder should contains the following files: CoastalFires_DischargeData.csv
input_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/Input_Files" # Change this to match your data frame location. Make sure it ends in "/Input_Files"

#Output Files folder should contains the following files:
#From ESS-DIVE: CoastalFires_DataAnalysis.Rmd
working_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/"  # Change this to match your scripts location

CombinedData <- readr::read_csv(file.path(working_folder,"CoastalFires_CombinedData.csv"))
DischargeData <- readr::read_csv(file.path(input_folder, "CoastalFires_DischargeData.csv"))
``` 

#Unit Conversions: converting units to molar-C in order to better compare between levoglucosan and black cabron
```{r}
#dissolved black carbon and dissolved organic carbon are in micro grams of carbon per liter. To get both values into micro moles of carbon per liter, divide by number of C atoms in molecule to convert units into mM-C. Then multiply by 1000 to convert from mM-C to uM-C.
CombinedData$DOC_micro_mol_C_per_L <- CombinedData$DOC_mgC_per_L / 12 * 1000
CombinedData$DBC_micro_mol_C_per_L <- CombinedData$DBC_mgC_per_L / 12 * 1000

#individual BPCAs are in micro moles per liter. Multiply by number of C atoms in molecule to convert units to nM-C.Then divide by 1000 to convert from nM-C to uM-C.
CombinedData$B6CA_micro_mol_C_per_L <- CombinedData$dB6CA_nano_mol_per_L * 12 / 1000 
CombinedData$B5CA_micro_mol_C_per_L <- CombinedData$dB5CA_nano_mol_per_L * 11 / 1000 

#Levoglucosan is in micro moles per liter. Divide by molecular weight of levoglucosan to get units into uM-levoglucosan. Then multiply by number of C atoms in molecule to convert units to uM-C.
CombinedData$Levo_micro_mol_C_per_L <- CombinedData$levo_micro_gram_per_liter / 162.14 * 6
```

#Calculate levoclucosan-dissolved organic carbon and [B6CA+B5CA]-dissolved organic carbon ratios
```{r}
#Add a new column that calculates the concentration of both B6CA and B5CA molecular markers. To avoid bias (due to conversion factors associated with calculating black carbon concentrations with BPCAs) when comparing black carbon and levoglucosan, we use the combined concentration of B6CA and B5CA. 
CombinedData$B6CA_and_B5CA_micro_mol_per_L <- CombinedData$B6CA_micro_mol_C_per_L + CombinedData$B5CA_micro_mol_C_per_L

#calculate the ratio of levoglucosan to bulk DOC and DBC
CombinedData$Levo_DOC_percent <- CombinedData$Levo_micro_mol_C_per_L / CombinedData$DOC_micro_mol_C_per_L * 100
CombinedData$BPCA_DOC_percent <- CombinedData$B6CA_and_B5CA_micro_mol_per_L / CombinedData$DOC_mgC_per_L * 100
```

#Calculating a "Time Since Fire" metric
```{r}
#Ensuring that the Date column is correctly classified as a date rather than a character
CombinedData$Date <- as.Date(CombinedData$Date, format = "%d-%b-%y")

#Calculating the time since fire for all samples and ensuring it is a numeric value
fire_date <- as.Date("2020-08-16")
CombinedData$time_since_fire_day <- as.numeric(CombinedData$Date - fire_date)
```

Test data for normality using Shapiro-Wilk Test
```{r}
#are all variables normally distributed?
#filter to only numeric data
numeric_columns <- sapply(CombinedData, is.numeric)
Data_numeric <- CombinedData[, numeric_columns]
Data_numeric <- Data_numeric %>% select(-Latitude, -Longitude)

# Apply Shapiro-Wilk test and create a readable summary
shapiro_results <- sapply(Data_numeric, function(x) {
  test <- shapiro.test(x)
  return(c(p_value = test$p.value, W = test$statistic))
})

# Convert results to a data frame
shapiro_df <- as.data.frame(t(shapiro_results))
shapiro_df$Variable <- rownames(shapiro_df)
rownames(shapiro_df) <- NULL

# Print readable results
print(shapiro_df)
```

The differences of concentrations and key ratios by watershed
```{r}
#is the concentration of levoglucosan different across watersheds?
kruskal_test_levo <- kruskal.test(levo_micro_gram_per_liter ~ Watershed, data = CombinedData)
print(kruskal_test_levo)
kruskal_test_Qnorm_levo <- kruskal.test(Q_norm_levo ~ Watershed, data = CombinedData)
print(kruskal_test_Qnorm_levo)
#is the ratio of levoglucosan to bulk DOC different across watersheds?
kruskal_test_levotoDOC <- kruskal.test(Levo_DOC_percent ~ Watershed, data = CombinedData)
print(kruskal_test_levotoDOC)
#is the ratio of levoglucosan to DBC different across watersheds?
kruskal_test_levotoDBC <- kruskal.test(Levo_DBC ~ Watershed, data = CombinedData)
print(kruskal_test_levotoDBC)
```

Is a significant difference between watersheds for Q-norm levoglucosan?
```{r}
#Q_norm_levo is the concentration of levoglucosan divided by discharge
wilcoxon_test_Qnorm_levo <- pairwise.wilcox.test(CombinedData$Q_norm_levo, CombinedData$Watershed, p.adjust.method = "bonferroni")
print(wilcoxon_test_Qnorm_levo)
```

Time since fire - Spearman's rank
```{r}
#make a table to showcase the coefficients and p-values
# Perform Spearman's rank correlation using the averaged data
spearman_corr_levo_ppb <- cor.test(CombinedData$time_since_fire_day, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_Qnorm_levo <- cor.test(CombinedData$time_since_fire_day, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_levoDOC <- cor.test(CombinedData$time_since_fire_day, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_DBC_ppm <- cor.test(CombinedData$time_since_fire_day, CombinedData$DBC_mgC_per_L, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_dbcDOC <- cor.test(CombinedData$time_since_fire_day, CombinedData$DBC_DOC_percent, method = "spearman", use = "pairwise.complete.obs")

# Extract p-values and rho values
results <- data.frame(
  Comparison = c("time_since_fire_day vs levo_ppb",
                  "time_since_fire_day vs Qnorm_levo",
                 "time_since_fire_day vs levoDOC",
                  "time_since_fire_day vs DBC",
                  "time_since_fire_day vs dbcDOC"),
  Rho = c(spearman_corr_levo_ppb$estimate,
          spearman_corr_Qnorm_levo$estimate,
          spearman_corr_levoDOC$estimate,
          spearman_corr_DBC_ppm$estimate,
          spearman_corr_dbcDOC$estimate),
  P_Value = c(spearman_corr_levo_ppb$p.value,
              spearman_corr_Qnorm_levo$p.value,
              spearman_corr_levoDOC$p.value,
              spearman_corr_DBC_ppm$p.value,
              spearman_corr_dbcDOC$p.value)
)

# Print the results table
print(results)
```

Miscellaneous Correlation Significant Tests for Levoglucosan Concentrations
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by discharge
spearman_corr_Q <- cor.test(CombinedData$Discharge_cms, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")

# Extract p-values and rho values
results <- data.frame(
  Comparison = c("levo_ppb vs burn_extent",
                  "levo_ppb vs low burn_extent",
                  "levo_ppb vs low-mod burn_extent",
                 "levo_ppb vs mod-high burn extent", 
                 "levo_ppb vs watershed area", 
                 "levo_ppb vs watershed slope",
                 "levo_ppb vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)


```

Miscellaneous Correlation Significant Tests for Discharge normalized Levoglucosan concentrations
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$Q_norm_levo, method = "spearman", use = "pairwise.complete.obs")


# Extract p-values and rho values
results <- data.frame(
  Comparison = c("Qnorm_levo vs burn_extent",
                  "Qnorm_levo vs low burn_extent",
                  "Qnorm_levo vs low-mod burn_extent",
                 "Qnorm_levo vs mod-high burn extent", 
                 "Qnorm_levo vs watershed area", 
                 "Qnorm_levo vs watershed slope",
                 "Qnorm_levo vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)
```

Miscellaneous Correlation Significant Tests for Levo:DOC
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")
#by discharge
spearman_corr_Q <- cor.test(CombinedData$Discharge_cms, CombinedData$Levo_DOC_percent, method = "spearman", use = "pairwise.complete.obs")

#Extract p-values and rho values
results <- data.frame(
  Comparison = c("Levo_DOC_perc vs burn_extent",
                  "Levo_DOC_perc vs low burn_extent",
                  "Levo_DOC_perc vs low-mod burn_extent",
                 "Levo_DOC_perc vs mod-high burn extent", 
                 "Levo_DOC_perc vs watershed area", 
                 "Levo_DOC_perc vs watershed slope",
                 "Levo_DOC_perc vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)
```

NMDS
Prepare data for NMDS
```{r}
#Isolate PyC concentrations and relevant ratios 
PyC_conc_rat <- CombinedData[, c("levo_micro_gram_per_liter", "DOC_mgC_per_L", "DBC_mgC_per_L", "DBC_DOC_percent", "Levo_DOC_percent")]

#put levoglucosan concentrations into the same units as DOC and DBC
PyC_conc_rat$levo_mg_per_L <- PyC_conc_rat$levo_micro_gram_per_liter/1000 #dividing micrograms per liter by 1000 to get levoglucosan concentrations in milligrams per liter

#delete the levoglucosan column with concentrations in micrograms per liter
analyte_data_conc_rat <- PyC_conc_rat[,c("levo_mg_per_L", "DOC_mgC_per_L", "DBC_mgC_per_L", "DBC_DOC_percent", "Levo_DOC_percent")]

View(analyte_data_conc_rat)
```

Run NMDS
```{r}
nmds_result <- metaMDS(analyte_data_conc_rat, k = 2, trymax = 100)
print(nmds_result)
```


Plot NMDS Results - Full NMDS
```{r}
# Extract NMDS scores for plotting
nmds_scores <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_species<-as.data.frame(scores(nmds_result, display="species"))
nmds_scores$Site <- CombinedData$Watershed  # Add Site info for coloring
nmds_scores$time <- CombinedData$time_since_fire_day #this is for coloring the data points by time
nmds_scores$watershed <- CombinedData$Watershed #this is for the point shape by watershed

# Environmental variables 
env_variables <- CombinedData[c("time_since_fire_day", "Area_low_severity_percent", "Discharge_cms", "Area_square_kilometer", "Average_slope_percent", "Burn_Extent_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent")]

#assign environmental variables to df
envfit_result <- envfit(nmds_result, env_variables, permutations = 999)

#have the environmental variables assigned to arrows for plotting
en_coord_cont = as.data.frame(scores(envfit_result, "vectors")) * ordiArrowMul(envfit_result)

# Shorten labels for plotting species and vectors
nmds_species$ShortLabel <- c("Levo", "DOC", "DBC", "DBC:DOC", "Levo:DOC")  # Replace with your short labels

# Change labels for plotting vectors. Also wrapping text for improving clarity of the plot
en_coord_cont$WrappedLabel <- c(
  "Time Since\nFire", #Time Since Fire
  "Low Severity\nBurn Extent", #Low Seveirty Burn Extent
  "Discharge", #Discharge
  "Watershed\nArea",
  "Slope",
  "Burn Extent",
  "Low-Mod\nBurn Extent",
  "Mod-High\nBurn Extent" #Moderate to High Burn Extent
)
  
#NMDS Visual Plot
ggplot(nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  
  #Sites
  geom_point(data = nmds_scores, aes(size = 100, color = time, fill = time, shape = watershed)) +
     scale_shape_manual(values = c(21, 22, 24, 23, 25))+
  scale_fill_gradient(low="#00FFFF", high="navy")+
  scale_color_gradient(low="#00FFFF", high="navy")+
 
  #Species
  geom_point(data = nmds_species, aes(x = NMDS1, y = NMDS2), shape = 8, stroke = 2, size = 3, alpha = 1, colour = "hotpink") +
  
  #vectors
  geom_segment(aes(x = 0, y = 0, xend = NMDS1*0.75, yend = NMDS2*0.75), data = en_coord_cont, linewidth =1, alpha = 1, colour = "black", arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +

#Add repelled labels for species and vectors
  geom_text_repel(data = nmds_species, aes(x = NMDS1, y = NMDS2, label = ShortLabel), colour = "black", fontface = "bold", size = 3, box.padding = 0.2, max.overlaps = Inf) + #species
  geom_text_repel(data = en_coord_cont, aes(x = NMDS1*0.75, y = NMDS2*0.75, label = WrappedLabel), size = 3, box.padding = 0.5, max.overlaps = 50) + #vectors
  
  
  #chart title
  labs(x = "NMDS1",
       y = "NMDS2") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(size = 0.5),  # Add axis lines
    axis.ticks = element_line(size = 0.5),  # Add axis ticks
    axis.ticks.length = unit(3, "pt"),  # Make ticks outward
    axis.line.x = element_line(size = 0.5),
    axis.line.y = element_line(size = 0.5)
  )

#Determine which vectors are statistically significant
envfit_result
```

Plot NMDS Results - Simplified NMDS Figure
```{r}
# Extract NMDS scores for plotting
nmds_scores <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_species<-as.data.frame(scores(nmds_result, display="species"))
nmds_scores$Site <- CombinedData$Watershed  # Add Site info for coloring
nmds_scores$time <- CombinedData$time_since_fire_day #this is for coloring the data points by time
nmds_scores$watershed <- CombinedData$Watershed #this is for the point shape by watershed

# Environmental variables (adjust to your desired variables)
env_variables <- CombinedData[c("time_since_fire_day", "Area_low_severity_percent", "Discharge_cms", "Area_mod_to_high_severity_percent", "Area_square_kilometer")]

#If want ALL environmental variables for sasha, use the code below
#env_variables <- CombinedData[c("time_since_fire_day", "Area_low_severity_percent", "Discharge_cms", "Area_square_kilometer", "Average_slope_percent", "Burn_Extent_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent")]

#assign environmental variables to df
envfit_result <- envfit(nmds_result, env_variables, permutations = 999)

#have the environmental variables assigned to arrows for plotting
en_coord_cont = as.data.frame(scores(envfit_result, "vectors")) * ordiArrowMul(envfit_result)

# Shorten labels for plotting species and vectors
nmds_species$ShortLabel <- c("Levo", "DOC", "DBC", "DBC:DOC", "Levo:DOC")  # Replace with your short labels

# Change labels for plotting vectors. Also wrapping text for improving clarity of the plot
en_coord_cont$WrappedLabel <- c(
  "Time Since\nFire", #Time Since FIre
  "Low Severity\nBurn Extent", #Low Seveirty Burn Extent
  "Discharge", #Discharge
  "Mod-High\nBurn Extent", #Moderate to High Burn Extent
  "Watershed\nArea" #Watershed Area
)
  
#NMDS Visual Plot
ggplot(nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  
  #Sites
  geom_point(data = nmds_scores, aes(size = 100, color = time, fill = time, shape = watershed)) +
     scale_shape_manual(values = c(21, 22, 24, 23, 25))+
  scale_fill_gradient(low="#00FFFF", high="navy")+
  scale_color_gradient(low="#00FFFF", high="navy")+
 
  #Species
  geom_point(data = nmds_species, aes(x = NMDS1, y = NMDS2), shape = 8, stroke = 2, size = 3, alpha = 1, colour = "hotpink") +
  
  #vectors
  geom_segment(aes(x = 0, y = 0, xend = NMDS1*0.75, yend = NMDS2*0.75), data = en_coord_cont, linewidth =1, alpha = 1, colour = "black", arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +

#Add repelled labels for species and vectors
  geom_text_repel(data = nmds_species, aes(x = NMDS1, y = NMDS2, label = ShortLabel), colour = "black", fontface = "bold", size = 3, box.padding = 0.2, max.overlaps = Inf) + #species
  geom_text_repel(data = en_coord_cont, aes(x = NMDS1*0.75, y = NMDS2*0.75, label = WrappedLabel), size = 3, box.padding = 0.5, max.overlaps = 50) + #vectors
  
  
  #chart title
  labs(x = "NMDS1",
       y = "NMDS2") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(size = 0.5),  # Add axis lines
    axis.ticks = element_line(size = 0.5),  # Add axis ticks
    axis.ticks.length = unit(3, "pt"),  # Make ticks outward
    axis.line.x = element_line(size = 0.5),
    axis.line.y = element_line(size = 0.5)
  )

#Determine which vectors are statistically significant
envfit_result
```

#Calculate fluxes and yields
```{r}
#discharge normalized levoglucosan concentrations
CoastalFires_CombinedData$Q_norm_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_micro_gram_per_liter / CoastalFires_CombinedData$Discharge_cms)

#flux
CoastalFires_CombinedData$flux_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_micro_gram_per_liter * CoastalFires_CombinedData$Discharge_cms)
CoastalFires_CombinedData$flux_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DOC_mgC_per_L * CoastalFires_CombinedData$Discharge_cms)
CoastalFires_CombinedData$flux_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DBC_mgC_per_L * CoastalFires_CombinedData$Discharge_cms)

#yield
CoastalFires_CombinedData$yield_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_levo / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DOC / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DBC / CoastalFires_CombinedData$Area_square_kilometer)
```
