---
title: "CoastalFires_Dataset_Wrangling"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
Before starting, make sure to read the "Manuscript_Workflow_Instructions.pdf" associated with this script. Before you begin, you must find and download necessary files:

Levoglucosan data from ESS-DIVE:
  CoastalFires_LevoglucosanData.csv

Black carbon data from CUAHSI Hydroshare:
Barton, R., C. Richardson, E. Pae, M. Montalvo, M. Redmond, M. Zimmer, S. Wagner (2025). Hydrology, rather than wildfire burn extent, determines post-fire organic and black carbon export from mountain rivers in central coastal California [Dataset], HydroShare, https://doi.org/10.4211/hs.421c0226bb38460c8393d67fe0c4f802 
  Organic and Black carbon data: CoastalFires_BiogeochemData.csv
  Discharge data: CoastalFires_DischargeData.csv
  Watershed Metadata: CoastalFires_WatershedData.csv
  

Install package if necessary
```{r}
install.packages("tidyverse")
```

Load library
```{r}
library(tidyverse)
```

Set directory for input/output files and working directory
```{r}
#Input files folder should contains the following files: CoastalFires_LevoglucosanData.csv, CoastalFires_BiogeochemData.csv, CoastalFires_DischargeData.csv, CoastalFires_WatershedData.csv
input_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/Input" # Change this to match your directory. Make sure it ends in "/Input"

#working folder should contain: ReadMe.txt, CoastalFires_DataFrameWrangling.Rmd, and CoastalFires_DataAnalysis.Rmd
working_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/"  # Change this to match your directory.

#out put folder should not contain any files yet
output_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/Output" #Change this to match your directory. Make sure it ends in "/Output"
```

Load datafiles from "Input" folder: levoglucosan data, watershed and burn severity data, and previously published organic carbon and black carbon data
```{r}
#From ESS-DIVE
LevoData <- readr::read_csv(file.path(input_folder, "CoastalFires_LevoglucosanData.csv"))

#From CUAHSI Hydroshare
BlackCarbonData <- readr::read_csv(file.path(input_folder, "CoastalFires_BiogeochemData.csv"))
WatershedData <- readr::read_csv(file.path(input_folder, "CoastalFires_WatershedData.csv"))
``` 

Combine LevoData, WatershedData, and BlackCarbonData data frames to one centralized dataframe: CombinedData
```{r}
#combine data frames
CombinedData <- left_join(LevoData, BlackCarbonData, by = "Sample_ID")

#add watershed metadata and remove duplicated columns
CombinedData <- left_join(CombinedData, WatershedData, by = "Watershed")
CombinedData <- CombinedData %>% select(-Lat, -Long) #delete extra location columns
```

Unit Conversions: converting units to micro moles of C per liter (uM-C) in order to better compare between levoglucosan and black carbon and DOC
```{r}
#dissolved black carbon (DBC) and dissolved organic carbon (DOC) are in micro grams of carbon per liter. To get both values into micro moles of carbon per liter, divide by mass of C atom to convert units into milli moles of C per liter (mM-C). Then multiply by 1000 to convert from mM-C to uM-C.
CombinedData$DOC_micro_mol_C_per_L <- CombinedData$DOC_mgC_per_L / 12 * 1000
CombinedData$DBC_micro_mol_C_per_L <- CombinedData$DBC_mgC_per_L / 12 * 1000

#individual BPCAs are in nano moles per liter (nM). Multiply by number of C atoms in molecule to convert units to nM-C.Then divide by 1000 to convert from nM-C to uM-C.
CombinedData$B6CA_micro_mol_C_per_L <- CombinedData$dB6CA_nano_mol_per_L * 12 / 1000 #B6CA, or benzenehexacarboxylic acid, has 12 C atoms
CombinedData$B5CA_micro_mol_C_per_L <- CombinedData$dB5CA_nano_mol_per_L * 11 / 1000 #B5CA, or benzenepentacarboxylic acid, has 12 C atoms

#Levoglucosan is in micro grams per liter. Divide by molecular weight of levoglucosan to get units into micro moles of levoglucosan per liter (uM). Then multiply by number of C atoms in the levoglucosan molecule to convert units to uM-C.
CombinedData$Levo_micro_mol_C_per_L <- CombinedData$levo_micro_gram_per_liter / 162.14 * 6 #levoglucosan has six C atoms in its structure
```

PyC-to-DOC ratios: calculate levoclucosan-dissolved organic carbon and [B6CA+B5CA]-dissolved organic carbon ratios
```{r}
#Add a new column that calculates the concentration of both B6CA and B5CA molecular markers. To avoid bias (due to conversion factors associated with calculating black carbon concentrations with BPCAs) when comparing black carbon and levoglucosan, we use the combined concentration of B6CA and B5CA. 
CombinedData$B6CA_and_B5CA_micro_mol_C_per_L <- CombinedData$B6CA_micro_mol_C_per_L + CombinedData$B5CA_micro_mol_C_per_L

#calculate the ratio of levoglucosan to bulk DOC and BPCA to DOC
CombinedData$Levo_DOC_percent <- CombinedData$Levo_micro_mol_C_per_L / CombinedData$DOC_micro_mol_C_per_L * 100
CombinedData$BPCA_DOC_percent <- CombinedData$B6CA_and_B5CA_micro_mol_C_per_L / CombinedData$DOC_micro_mol_C_per_L * 100
```

Calculating a "Time Since Fire" metric
```{r}
#Ensuring that the Date column is correctly classified as a date rather than a character
CombinedData$Date <- as.Date(CombinedData$Date, format = "%d-%b-%y")

#Calculating the time since fire for all samples and ensuring it is a numeric value
fire_date <- as.Date("2020-08-16")
CombinedData$time_since_fire_day <- as.numeric(CombinedData$Date - fire_date)
```

Clean up the csv: remove duplicate columns
```{r}
#Remove unnecessary columns. The columns that are being removed with the following script are associated with the particulate phase (i.e., particulate organic carbon and particulate black carbon). As levoglucosan concentrations were determined only for the dissolved phase, the following data analysis will also only focus on the dissolved phase. 
CombinedData <- CombinedData %>%
  select(-TSS_mg_per_L, -POC_mgC_per_L, -`PBC_mgC_per_L`, -PBC_TSS_percent, 
         -PBC_POC_percent, -pB6CA_nano_mol_per_L, -pB5CA_nano_mol_per_L, -pB6CA_B5CA, -Flow_category)

#Remove levoglucosan, DBC, BPCA, and DOC columns with old units
CombinedData <- CombinedData %>%
  select(-DOC_mgC_per_L, -DBC_mgC_per_L, -dB6CA_nano_mol_per_L, -dB5CA_nano_mol_per_L, 
         -levo_micro_gram_per_liter, -std_levo_micro_gram_per_liter)
```

Re-arrange column order
```{r}
# Specify the desired column order
desired_order <- c("Sample_ID", "Watershed", "Date", "Time", "time_since_fire_day", "Latitude", "Longitude", "Location_Change",
                   "Discharge_cms", "DOC_micro_mol_C_per_L", "Levo_micro_mol_C_per_L", "B6CA_micro_mol_C_per_L",
                   "B5CA_micro_mol_C_per_L", "B6CA_and_B5CA_micro_mol_C_per_L", "dB6CA_B5CA",
                   "Levo_DOC_percent", "BPCA_DOC_percent", "DBC_micro_mol_C_per_L", "DBC_DOC_percent",
                   "a254_per_m", "SUVA254_L_per_mgC_per_m", 
                   "Area_square_kilometer", "Average_slope_percent", 
                   "Average_precipitation_milli_meter", "Burn_Extent_percent", 
                   "Area_low_severity_percent", "Area_low_to_mod_severity_percent", 
                   "Area_mod_to_high_severity_percent", "Forest_percent", 
                   "Shrub_percent", "Developed_percent", 
                   "Sandstone_and_sedimentary_percent", "Igneous_percent", 
                   "Suficial_Sediments_percent", "Schist_percent")

# Rearrange the columns
CombinedData <- CombinedData[, desired_order]
```

Create a csv file of the combined dataset:CoastalFires_CombinedData
```{r}
write.csv(CombinedData, 
          file = file.path(output_folder, "CoastalFires_CombinedData.csv"), 
          row.names = FALSE)  #Prevents writing row numbers
```