---
title: "Figure 2"
author: "rbb"
date: "2026-07-09"
output: html_document
---
Before starting, make sure to read the README found on the GitHub repository. This script is utilized for data analysis and preliminary figure creation.

Install necessary packages (run once if necessary)
```{r}
# install.packages("dplyr")
# install.packages("lubridate")
# install.packages("tidyr")
# install.packages("patchwork")
# install.packages("cowplot")
# install.packages("grid")
# install.packages("ggplot2")
```

Packages
```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(patchwork)
library(cowplot)
library(grid)
library(ggplot2)
```

Get Data
```{r}
#From CUAHSI Hydroshare
#DischargeData <- readr::read_csv("https://www.hydroshare.org/resource/26f08ef7fe344fcb9d3da45160b697fe/data/contents/CoastalFires_DischargeData.csv")

#flow data was rounded...so for now it is here
DischargeData <- readr::read_csv("data/CoastalFires_FlowData.csv") #use this instead since the LAG values are rounded on CUAHSI

#From GitHub
CombinedData <- readr::read_csv("data/CoastalFires_CombinedData.csv")
```

Exclude samples that were taken at a different location. 
```{r}
#Scott Creek flooded during the events of January 27-28th, 2021. Samples were collected at a downstream lagoon site. However, due to this the data collected does not accurately reflect that at the sample site. As such authors of the associated manuscript have excluded it from all data analysis and figure creation.
CombinedData <- CombinedData %>%
  filter(Location_Change != "Y")
```

Create DateTime
```{r}
CombinedData <- CombinedData %>%
  mutate(
    Time = if_else(is.na(Time), "12:00:00", as.character(Time)), #assume noon if no time specified
    DateTime = as.POSIXct(
      paste(Date, Time),
      format = "%Y-%m-%d %H:%M:%S",
      tz = "America/Los_Angeles"
    )
  )
```

Get standard error of replicates for error bars
```{r}
CombinedData$SE_Levo <- CombinedData$stdev_Levo_micro_mol_C_per_L / sqrt(2)
```

Simplify Dataframe
```{r}
ConcData <- CombinedData %>%
  select(Watershed, DateTime, Levo_micro_mol_C_per_L, stdev_Levo_micro_mol_C_per_L, B6CA_and_B5CA_micro_mol_C_per_L)

#Fix watershed names to match discharge data
ConcData <- CombinedData %>%
  select(
    Watershed,
    DateTime,
    Levo_micro_mol_C_per_L,
    stdev_Levo_micro_mol_C_per_L,
    SE_Levo,
    B6CA_and_B5CA_micro_mol_C_per_L
  ) %>%
  mutate(
    Watershed = recode(
      Watershed,
      "Laguna Creek" = "LAG",
      "San Lorenzo River" = "SAN",
      "Pescadero Creek" = "PESC",
      "Majors Creek" = "MAJ",
      "Scott Creek" = "SCT"
    )
  )
```

Alter discharge dataframe for easier plotting
```{r}
SAN <- DischargeData %>%
  transmute(
    Watershed = "SAN",
    DateTime = parse_date_time(SAN_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = SAN_Flow_cms
  )

PESC <- DischargeData %>%
  transmute(
    Watershed = "PESC",
    DateTime = parse_date_time(PESC_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = PESC_Flow_cms
  )

LAG <- DischargeData %>%
  transmute(
    Watershed = "LAG",
    DateTime = parse_date_time(LAG_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = LAG_Flow_cms
  )

MAJ <- DischargeData %>%
  transmute(
    Watershed = "MAJ",
    DateTime = parse_date_time(MAJ_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = MAJ_Flow_cms
  )

SCT <- DischargeData %>%
  transmute(
    Watershed = "SCT",
    DateTime = parse_date_time(SCT_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = SCT_Flow_cms
  )

FlowData <- bind_rows(SAN, PESC, LAG, MAJ, SCT)

#Remove NAs from FlowData
FlowData <- FlowData %>%
  filter(!is.na(DateTime),
         !is.na(Flow))

#Cut data to prefered date range
FlowData <- FlowData %>%
  filter(as.Date(DateTime) >= as.Date("2020-10-01"),
         as.Date(DateTime) <= as.Date("2021-05-2"))

#Remove zero flows before averaging/log transformation
FlowData <- FlowData %>%
  filter(Flow > 0)

#check it worked
FlowData %>%
  group_by(Watershed) %>%
  summarize(
    missing_dates = sum(is.na(DateTime)),
    start = min(DateTime),
    end = max(DateTime),
    n=n()
  )
```

Scaling factor
```{r}
# Prepare flow for plotting
FlowData <- FlowData %>%
  mutate(
    log_flow = log10(Flow)
  )

# Common axis scaling for ALL watersheds
levo_axis_max <- 2.5

# Force common discharge axis:
# 0.01 to 100 m3/s
flow_axis_min <- log10(0.01)
flow_axis_max <- log10(100)

scale_factor <- levo_axis_max /
  (flow_axis_max - flow_axis_min)

scale_factor
```
Color scheme
```{r}
ws_cols <- c(
  SAN  = "#FBF9A2",
  PESC = "#FFCF55",
  MAJ  = "#FF7410",
  LAG  = "#EA4412",
  SCT  = "#A00000"
)

ws_shapes <- c(
  SAN  = 23,
  PESC = 24,
  MAJ  = 22,
  LAG  = 21,
  SCT  = 25
)
```

Function for making plots
```{r}
make_levo_flow_plot <- function(watershed_name,
                                point_fill,
                                point_shape,
                                panel_label,
                                show_x_axis = TRUE) {
  
  levo_plot <- ConcData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.POSIXct(DateTime)
    )
  
  
  flow_plot <- FlowData %>%
    filter(
      Watershed == watershed_name
    )
  
  
  ggplot() +
    
    geom_line(
      data = flow_plot,
      aes(
        x = DateTime,
        y = (log_flow - flow_axis_min) * scale_factor
      ),
      color = "#156082",
      linewidth = 1
    ) +
    
    
    geom_errorbar(
      data = levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - SE_Levo,
        ymax = Levo_micro_mol_C_per_L + SE_Levo
      ),
      width = 0,
      linewidth = 0.4
    ) +
    
    geom_line(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      linetype = "dotted",
      linewidth = 0.5,
      color = "black"
    ) +
    
    geom_point(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = point_shape,
      size = 4,
      fill = point_fill,
      color = "black",
      stroke = 0.8
    ) +
    
    
    scale_y_continuous(
      limits = c(0, levo_axis_max),
      
      name = expression(
        Levoglucosan~("μmol-C"~L^{-1})),

      sec.axis = sec_axis(
        ~ . / scale_factor + flow_axis_min,
        breaks = log10(c(0.01, 0.1, 1, 10, 100)),
        labels = c(
          "0.01",
          "0.1",
          "1",
          "10",
          "100"
        ),
        name = expression(
          "Discharge (m"^3*" s"^{-1}*")"
        )
      )
    ) +
    
    
    scale_x_datetime(
      name = NULL,
      date_breaks = "1 month",
      date_labels = "%b"
    ) +
    
    annotate(
      "text",
      x = -Inf,
      y = Inf,
      label = panel_label,
      hjust = -0.5,
      vjust = 1.5,
      fontface = "bold",
      size = 5
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid = element_blank()
    )
}
```

Make a plot for each watershed
```{r}
SAN_plot <- make_levo_flow_plot(
  "SAN",
  point_fill = ws_cols["SAN"],
  point_shape = ws_shapes["SAN"],
  panel_label = "A)",
  show_x_axis = TRUE
)

PESC_plot <- make_levo_flow_plot(
  "PESC",
  point_fill = ws_cols["PESC"],
  point_shape = ws_shapes["PESC"],
  panel_label = "B)",
  show_x_axis = TRUE
)

MAJ_plot <- make_levo_flow_plot(
  "MAJ",
  point_fill = ws_cols["MAJ"],
  point_shape = ws_shapes["MAJ"],
  panel_label = "C)",
  show_x_axis = TRUE
)

LAG_plot <- make_levo_flow_plot(
  "LAG",
  point_fill = ws_cols["LAG"],
  point_shape = ws_shapes["LAG"],
  panel_label = "D)",
  show_x_axis = TRUE
)

SCT_plot <- make_levo_flow_plot(
  "SCT",
  point_fill = ws_cols["SCT"],
  point_shape = ws_shapes["SCT"],
  panel_label = "E)",
  show_x_axis = TRUE
)
```

Combine the watershed plots into a final plot
```{r}
#combine plots
final_plot <- plot_grid(
  SAN_plot,  LAG_plot,
  PESC_plot, SCT_plot,
  MAJ_plot,  NULL,
  ncol = 2,
  align = "hv",
  axis = "tblr"
)

#Add Date label for x-axis
final_plot <- plot_grid(
  final_plot,
  ggdraw() + draw_label("Date", size = 12),
  ncol = 1,
  rel_heights = c(1, 0.05)
)
```

Save Final Plot - no legend
```{r}
ggsave(
  filename = "figures/Fig2_Levo_StreamFlow.tiff",
  plot = final_plot,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 10,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```

Create final plot with legend
```{r}
legend_labels <- c(
  "San Lorenzo River (SAN)",
  "Pescadero Creek (PESC)",
  "Majors Creek (MAJ)",
  "Laguna Creek (LAG)",
  "Scott Creek (SCT)"
)

# Watershed legend rows
legend_rows <- lapply(seq_along(legend_labels), function(i) {
  
  ggdraw() +
    draw_grob(
      pointsGrob(
        x = 0.08,
        y = 0.5,
        pch = ws_shapes[i],
        gp = gpar(
          fill = ws_cols[i],
          col = "black",
          lwd = 1.5
        )
      )
    ) +
    draw_label(
      legend_labels[i],
      x = 0.22,
      y = 0.5,
      hjust = 0,
      size = 10
    )
})


# Discharge legend row
discharge_row <- ggdraw() +
  draw_grob(
    segmentsGrob(
      x0 = 0.04,
      x1 = 0.12,
      y0 = 0.5,
      y1 = 0.5,
      gp = gpar(
        col = "#156082",
        lwd = 3
      )
    )
  ) +
  draw_label(
    "Discharge",
    x = 0.22,
    y = 0.5,
    hjust = 0,
    size = 10
  )


# Legend title
legend_title <- ggdraw() +
  draw_label(
    "Watershed",
    x = 0.05,
    y = 0.5,
    hjust = 0,
    fontface = "bold",
    size = 11
  )

# Compact legend title + legend body
legend_body <- plot_grid(
  plotlist = c(
    legend_rows,
    list(discharge_row)
  ),
  ncol = 1,
  align = "v"
)

custom_legend <- plot_grid(
  legend_body,
  ncol = 1,
  rel_heights = c(0.08, 1)
)

# Add legend to figure

final_with_legend <- ggdraw() +
  draw_plot(final_plot) +
  draw_plot(
    custom_legend,
    x = 0.62,
    y = 0.05,
    width = 0.32,
    height = 0.27
  )

# display
final_with_legend

#Save
ggsave(
  filename = "figures/Fig2_wLegend.tiff",
  plot = final_with_legend,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 8,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```
