---
title: "Coastal Fires Levoglucosan"
author: "RB"
date: "2024-09-13"
output: html_document
---
#load necessary libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
```

#set working director, load data files, and rename them
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/")
CombinedData <- readr::read_csv("CoastalFires_CombinedData.csv")
DischargeData <- readr::read_csv("CoastalFires_DischargeData.csv")
``` 
Test data for normality using Shapiro-Wilk Test
```{r}
#are all variables normally distributed?
#filter to only numeric data
numeric_columns <- sapply(CombinedData, is.numeric)
Data_numeric <- CombinedData[, numeric_columns]
Data_numeric <- Data_numeric %>% select(-lat, -long)

# Apply Shapiro-Wilk test and create a readable summary
shapiro_results <- sapply(Data_numeric, function(x) {
  test <- shapiro.test(x)
  return(c(p_value = test$p.value, W = test$statistic))
})

# Convert results to a data frame
shapiro_df <- as.data.frame(t(shapiro_results))
shapiro_df$Variable <- rownames(shapiro_df)
rownames(shapiro_df) <- NULL

# Print readable results
print(shapiro_df)
```
#The differences of concentrations and key ratios by watershed
```{r}
#is the concentration of levoglucosan different across watersheds?
kruskal_test_levo <- kruskal.test(levo_micro_gram_per_liter ~ watershed, data = CombinedData)
print(kruskal_test_levo)
kruskal_test_flux_levo <- kruskal.test(flux_levo ~ watershed, data = CombinedData)
print(kruskal_test_flux_levo)
kruskal_test_yield_levo <- kruskal.test(yield_levo ~ watershed, data = CombinedData)
print(kruskal_test_yield_levo)
#is the ratio of levoglucosan to bulk DOC different across watersheds?
kruskal_test_levotoDOC <- kruskal.test(Levo_DOC_perc ~ watershed, data = CombinedData)
print(kruskal_test_levotoDOC)
#is the ratio of levoglucosan to DBC different across watersheds?
kruskal_test_levotoDBC <- kruskal.test(Levo_DBC ~ watershed, data = CombinedData)
print(kruskal_test_levotoDBC)
```

#since there is a significant difference between watersheds for levoglucosan flux and levoglucosan yield, what watersheds are significantly different?
```{r}
wilcoxon_test_flux_levo <- pairwise.wilcox.test(CombinedData$flux_levo, CombinedData$watershed, p.adjust.method = "bonferroni")
print(wilcoxon_test_flux_levo)

wilcoxon_test_yield_levo <- pairwise.wilcox.test(CombinedData$yield_levo, CombinedData$watershed, p.adjust.method = "bonferroni")
print(wilcoxon_test_yield_levo)
```

#Time since fire - Spearman's rank
```{r}
#make a table to showcase the coefficients and p-values
# Perform Spearman's rank correlation using the averaged data
spearman_corr_levo_ppb <- cor.test(CombinedData$time_since_fire_day, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_flux_levo <- cor.test(CombinedData$time_since_fire_day, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_yield_levo <- cor.test(CombinedData$time_since_fire_day, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_levoDOC <- cor.test(CombinedData$time_since_fire_day, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_DBC_ppm <- cor.test(CombinedData$time_since_fire_day, CombinedData$DBC_mgC_per_L, method = "spearman", use = "pairwise.complete.obs")

spearman_corr_dbcDOC <- cor.test(CombinedData$time_since_fire_day, CombinedData$DBC_DOC_perc, method = "spearman", use = "pairwise.complete.obs")

# Extract p-values and rho values
results <- data.frame(
  Comparison = c("time_since_fire_day vs levo_ppb",
                  "time_since_fire_day vs flux_levo",
                 "time_since_fire_day vs yield_levo",
                 "time_since_fire_day vs levoDOC",
                  "time_since_fire_day vs DBC",
                  "time_since_fire_day vs dbcDOC"),
  Rho = c(spearman_corr_levo_ppb$estimate,
          spearman_corr_flux_levo$estimate,
          spearman_corr_yield_levo$estimate,
          spearman_corr_levoDOC$estimate,
          spearman_corr_DBC_ppm$estimate,
          spearman_corr_dbcDOC$estimate),
  P_Value = c(spearman_corr_levo_ppb$p.value,
              spearman_corr_flux_levo$p.value,
              spearman_corr_yield_levo$p.value,
              spearman_corr_levoDOC$p.value,
              spearman_corr_DBC_ppm$p.value,
              spearman_corr_dbcDOC$p.value)
)

# Print the results table
print(results)
```
#Miscellaneous Correlation Significant Tests for Levoglucosan Concentrations
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")
#by discharge
spearman_corr_Q <- cor.test(CombinedData$Q_m_sq_per_sec, CombinedData$levo_micro_gram_per_liter, method = "spearman", use = "pairwise.complete.obs")

# Extract p-values and rho values
results <- data.frame(
  Comparison = c("levo_ppb vs burn_extent",
                  "levo_ppb vs low burn_extent",
                  "levo_ppb vs low-mod burn_extent",
                 "levo_ppb vs mod-high burn extent", 
                 "levo_ppb vs watershed area", 
                 "levo_ppb vs watershed slope",
                 "levo_ppb vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)


```

#Miscellaneous Correlation Significant Tests for Levoglucosan Flux
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$flux_levo, method = "spearman", use = "pairwise.complete.obs")


# Extract p-values and rho values
results <- data.frame(
  Comparison = c("flux_levo vs burn_extent",
                  "flux_levo vs low burn_extent",
                  "flux_levo vs low-mod burn_extent",
                 "flux_levo vs mod-high burn extent", 
                 "flux_levo vs watershed area", 
                 "flux_levo vs watershed slope",
                 "flux_levo vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)


```
#Miscellaneous Correlations with Levoglucosan Yield
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$yield_levo, method = "spearman", use = "pairwise.complete.obs")


# Extract p-values and rho values
results <- data.frame(
  Comparison = c("yield_levo vs burn_extent",
                  "yield_levo vs low burn_extent",
                  "yield_levo vs low-mod burn_extent",
                 "yield_levo vs mod-high burn extent", 
                 "yield_levo vs watershed area", 
                 "yield_levo vs watershed slope",
                 "yield_levo vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)


```

#Miscellaneous Correlation Significant Tests for Levo:DOC
``` {r}
#by burn extent
spearman_corr_BE <- cor.test(CombinedData$Burn_Extent_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by low severity burn extent
spearman_corr_BElow <- cor.test(CombinedData$Area_low_severity_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by low to moderate severity burn extent
spearman_corr_BElowtomod <- cor.test(CombinedData$Area_low_to_mod_severity_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by moderate to high severity burn extent
spearman_corr_BEmodtohigh <- cor.test(CombinedData$Area_mod_to_high_severity_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by watershed area
spearman_corr_area <- cor.test(CombinedData$Area_square_kilometer, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by watershed slope
spearman_corr_slope <- cor.test(CombinedData$Average_slope_percent, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")
#by discharge
spearman_corr_Q <- cor.test(CombinedData$Q_m_sq_per_sec, CombinedData$Levo_DOC_perc, method = "spearman", use = "pairwise.complete.obs")

#Extract p-values and rho values
results <- data.frame(
  Comparison = c("Levo_DOC_perc vs burn_extent",
                  "Levo_DOC_perc vs low burn_extent",
                  "Levo_DOC_perc vs low-mod burn_extent",
                 "Levo_DOC_perc vs mod-high burn extent", 
                 "Levo_DOC_perc vs watershed area", 
                 "Levo_DOC_perc vs watershed slope",
                 "Levo_DOC_perc vs discharge"),
  Rho = c(spearman_corr_BE$estimate,
          spearman_corr_BElow$estimate,
          spearman_corr_BElowtomod$estimate,
          spearman_corr_BEmodtohigh$estimate,
          spearman_corr_area$estimate,
          spearman_corr_slope$estimate,
          spearman_corr_Q$estimate),
  P_Value = c(spearman_corr_BE$p.value,
              spearman_corr_BElow$p.value,
              spearman_corr_BElowtomod$p.value,
              spearman_corr_BEmodtohigh$p.value,
              spearman_corr_area$p.value,
              spearman_corr_slope$p.value,
              spearman_corr_Q$p.value)
)

# Print the results table
print(results)


```

#Random Forest Models
```{r}
install.packages("randomForest")
install.packages("caret")  # Optional, for additional tools like cross-validation
library(randomForest)
library(caret)
```

#prepare data for RFM
```{r}
#remove any missing values
na.omit(CombinedData) 
#ensure categorical variables are factors
CombinedData$watershed <- as.factor(CombinedData$watershed) 
```

#RFM for Levo
```{r}
#Split levoglucosan dataset into model and test datasets
set.seed(123)  # For reproducibility
train_idx <- createDataPartition(CombinedData$levo_micro_gram_per_liter, p = 0.7, list = FALSE)
train_data <- CombinedData[train_idx, ]
test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_levoglucosan <- randomForest(
    levo_micro_gram_per_liter ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_levoglucosan)
varImpPlot(rf_levoglucosan)

#Evaluate RFM Performance

#predict of the test set
predictions <- predict(rf_levoglucosan, newdata = test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - test_data$levo_micro_gram_per_liter)^2))
r_squared <- cor(predictions, test_data$levo_micro_gram_per_liter)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

#Analyze varibale importance
#something
importance(rf_levoglucosan)

#visualize variable importance
varImpPlot(rf_levoglucosan)

#Tune the RFM

control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(train_data) - 1, by = 1)  # Tune mtry
)

rf_tuned <- train(
    levo_micro_gram_per_liter ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)

#use best model
levo_best_rf_model <- rf_tuned$finalModel
varImpPlot(levo_best_rf_model)
```
#RFM for for dissolved black carbon :) 
```{r}
#Split DBC dataset into model and test datasets
set.seed(123)  # For reproducibility
DBC_train_idx <- createDataPartition(CombinedData$DBC_mgC_per_L, p = 0.7, list = FALSE)
DBC_train_data <- CombinedData[train_idx, ]
DBC_test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_DBC <- randomForest(
    DBC_mgC_per_L ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_DBC)
varImpPlot(rf_DBC)

#Evaluate RFM Performance
#predict of the test set
predictions <- predict(rf_DBC, newdata = DBC_test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - DBC_test_data$DBC_mgC_per_L)^2))
r_squared <- cor(predictions, DBC_test_data$DBC_mgC_per_L)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

##Analyze varibale importance

#something
importance(rf_DBC)

#visualize variable importance
varImpPlot(rf_DBC)

#tune the DBC RFM
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(DBC_train_data) - 1, by = 1)  # Tune mtry
)

DBC_rf_tuned <- train(
    DBC_mgC_per_L ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)


#use best model
DBC_best_rf_model <- DBC_rf_tuned$finalModel
varImpPlot(DBC_best_rf_model)

```

#RFM for Levo:DOC :) 
```{r}
#Split DBC dataset into model and test datasets
set.seed(123)  # For reproducibility
LevoDOC_train_idx <- createDataPartition(CombinedData$Levo_DOC_perc, p = 0.7, list = FALSE)
LevoDOC_train_data <- CombinedData[train_idx, ]
LevoDOC_test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_LevoDOC <- randomForest(
    Levo_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_LevoDOC)
varImpPlot(rf_LevoDOC)

#Evaluate RFM Performance
#predict of the test set
predictions <- predict(rf_LevoDOC, newdata = LevoDOC_test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - LevoDOC_test_data$Levo_DOC_perc)^2))
r_squared <- cor(predictions, LevoDOC_test_data$Levo_DOC_perc)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

##Analyze varibale importance

#something
importance(rf_LevoDOC)

#visualize variable importance
varImpPlot(rf_LevoDOC)

#tune the DBC RFM
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(LevoDOC_train_data) - 1, by = 1)  # Tune mtry
)

LevoDOC_rf_tuned <- train(
    Levo_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)


#use best model
LevoDOC_best_rf_model <- DBC_rf_tuned$finalModel
varImpPlot(LevoDOC_best_rf_model)

```
#RFM for DBC:DOC :) 
```{r}
#Split DBC dataset into model and test datasets
set.seed(123)  # For reproducibility
DBCDOC_train_idx <- createDataPartition(CombinedData$DBC_DOC_perc, p = 0.7, list = FALSE)
DBCDOC_train_data <- CombinedData[train_idx, ]
DBCDOC_test_data <- CombinedData[-train_idx, ]

#Fit the RFM to the levoglucosan dataset

# Build the model on the training data
rf_DBCDOC <- randomForest(
    DBC_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    importance = TRUE,
    ntree = 500
)

# Print a summary of the model
print(rf_DBCDOC)
varImpPlot(rf_LevoDOC)

#Evaluate RFM Performance
#predict of the test set
predictions <- predict(rf_DBCDOC, newdata = DBCDOC_test_data)

#assess model accuracy
# Calculate RMSE and R²
rmse <- sqrt(mean((predictions - DBCDOC_test_data$DBC_DOC_perc)^2))
r_squared <- cor(predictions, DBCDOC_test_data$DBC_DOC_perc)^2

cat("RMSE:", rmse, "\n")
cat("R²:", r_squared, "\n")

##Analyze varibale importance

#something
importance(rf_DBCDOC)

#visualize variable importance
varImpPlot(rf_DBCDOC)

#tune the DBC RFM
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

tune_grid <- expand.grid(
    .mtry = seq(2, ncol(DBCDOC_train_data) - 1, by = 1)  # Tune mtry
)

DBCDOC_rf_tuned <- train(
    DBC_DOC_perc ~ time_since_fire_day + Q_m_sq_per_sec + Area_square_kilometer + Average_slope_percent + Burn_Extent_percent + Area_low_severity_percent + Area_low_to_mod_severity_percent + Area_mod_to_high_severity_percent,
    data = train_data,
    method = "rf",
    tuneGrid = tune_grid,
    trControl = control,
    importance = TRUE
)


#use best model
DBCDOC_best_rf_model <- DBC_rf_tuned$finalModel
varImpPlot(DBCDOC_best_rf_model)

```

