---
title: "Figure 2"
author: "rbb"
date: "2026-07-09"
output: html_document
---

Install necessary packages (run once if necessary)
```{r}
# install.packages("dplyr")
# install.packages("lubridate")
# install.packages("tidyr")
# install.packages("cowplot")
# install.packages("ggplot2")
```

Packages
```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(patchwork)
library(cowplot)
library(grid)
library(ggplot2)
```

Get Data
```{r}
#From CUAHSI Hydroshare
DischargeData <- readr::read_csv("https://www.hydroshare.org/resource/26f08ef7fe344fcb9d3da45160b697fe/data/contents/CoastalFires_DischargeData.csv")

#From GitHub
CombinedData <- readr::read_csv("data/CoastalFires_CombinedData.csv")
```

Exclude samples that were taken at a different location. 
```{r}
#Scott Creek flooded during the events of January 27-28th, 2021. Samples were collected at a downstream lagoon site. However, due to this the data collected does not accurately reflect that at the sample site. As such authors of the associated manuscript have excluded it from all data analysis and figure creation.
CombinedData <- CombinedData %>%
  filter(Location_Change != "Y")
```

Make sure dates are right
```{r}
CombinedData <- CombinedData %>%
  mutate(
    DateTime = if_else(
      is.na(Time),
      paste(Date, "12:00:00"),
      paste(Date, Time)
    ),
    DateTime = as.POSIXct(
      DateTime,
      format = "%Y-%m-%d %H:%M:%S",
      tz = "America/Los_Angeles"
    )
  )
```

Get mean error of replicates for error bars
```{r}
mean_levo_sd <- CombinedData %>%
  filter(!is.na(stdev_Levo_micro_mol_C_per_L)) %>%
  summarize(
    mean_sd = mean(stdev_Levo_micro_mol_C_per_L)
  ) %>%
  pull(mean_sd)
mean_levo_sd
```

Simplify Dataframe
```{r}
ConcData <- CombinedData %>%
  select(Watershed, DateTime, Levo_micro_mol_C_per_L, stdev_Levo_micro_mol_C_per_L, B6CA_and_B5CA_micro_mol_C_per_L)

#Fix watershed names to match discharge data
ConcData <- CombinedData %>%
  select(
    Watershed,
    DateTime,
    Levo_micro_mol_C_per_L,
    stdev_Levo_micro_mol_C_per_L,
    B6CA_and_B5CA_micro_mol_C_per_L
  ) %>%
  mutate(
    Watershed = recode(
      Watershed,
      "Laguna Creek" = "LAG",
      "San Lorenzo River" = "SAN",
      "Pescadero Creek" = "PESC",
      "Majors Creek" = "MAJ",
      "Scott Creek" = "SCT"
    )
  )
```

Alter discharge dataframe for easier plotting
```{r}
SAN <- DischargeData %>%
  transmute(
    Watershed = "SAN",
    DateTime = parse_date_time(SAN_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = SAN_Flow_cms
  )

PESC <- DischargeData %>%
  transmute(
    Watershed = "PESC",
    DateTime = parse_date_time(PESC_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = PESC_Flow_cms
  )

LAG <- DischargeData %>%
  transmute(
    Watershed = "LAG",
    DateTime = parse_date_time(LAG_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = LAG_Flow_cms
  )

MAJ <- DischargeData %>%
  transmute(
    Watershed = "MAJ",
    DateTime = parse_date_time(MAJ_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = MAJ_Flow_cms
  )

SCT <- DischargeData %>%
  transmute(
    Watershed = "SCT",
    DateTime = parse_date_time(SCT_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "America/Los_Angeles"),
    Flow = SCT_Flow_cms
  )

FlowData <- bind_rows(SAN, PESC, LAG, MAJ, SCT)

#Remove NAs from FlowData
FlowData <- FlowData %>%
  filter(!is.na(DateTime),
         !is.na(Flow))

#Cut data to prefered date range
FlowData <- FlowData %>%
  filter(as.Date(DateTime) >= as.Date("2020-10-01"),
         as.Date(DateTime) <= as.Date("2021-05-2"))

#Remove zero flows before averaging/log transformation
FlowData <- FlowData %>%
  filter(Flow > 0)

#check it worked
FlowData %>%
  group_by(Watershed) %>%
  summarize(
    missing_dates = sum(is.na(DateTime)),
    start = min(DateTime),
    end = max(DateTime),
    n=n()
  )
```

Scaling factor
```{r}
# Prepare flow for plotting
FlowData <- FlowData %>%
  mutate(
    log_flow = log10(Flow)
  )

# Common axis scaling for ALL watersheds
levo_axis_max <- 2.5

# Force common discharge axis:
# 0.01 to 100 m3/s
flow_axis_min <- log10(0.01)
flow_axis_max <- log10(100)

scale_factor <- levo_axis_max /
  (flow_axis_max - flow_axis_min)

scale_factor
```
Color scheme
```{r}
ws_cols <- c(
  SAN  = "#FBF9A2",
  PESC = "#FFCF55",
  MAJ  = "#FF7410",
  LAG  = "#EA4412",
  SCT  = "#A00000"
)

ws_shapes <- c(
  SAN  = 23,
  PESC = 24,
  MAJ  = 22,
  LAG  = 21,
  SCT  = 25
)
```


Function
```{r}
make_levo_flow_plot <- function(watershed_name,
                                point_fill,
                                point_shape,
                                panel_label,
                                show_x_axis = TRUE) {
  
  levo_plot <- ConcData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.POSIXct(DateTime),
      Levo_error = mean_levo_sd
    )
  
  
  flow_plot <- FlowData %>%
    filter(
      Watershed == watershed_name
    )
  
  
  ggplot() +
    
    geom_line(
      data = flow_plot,
      aes(
        x = DateTime,
        y = (log_flow - flow_axis_min) * scale_factor
      ),
      color = "#156082",
      linewidth = 1
    ) +
    
    
    geom_errorbar(
      data = levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4
    ) +
    
    geom_line(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      linetype = "dotted",
      linewidth = 0.5,
      color = "black"
    ) +
    
    geom_point(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = point_shape,
      size = 4,
      fill = point_fill,
      color = "black",
      stroke = 0.8
    ) +
    
    
    scale_y_continuous(
      limits = c(0, levo_axis_max),
      
      name = expression(
        "Levoglucosan ("*mu*"M-C)"
      ),
      
      sec.axis = sec_axis(
        ~ . / scale_factor + flow_axis_min,
        breaks = log10(c(0.01, 0.1, 1, 10, 100)),
        labels = c(
          "0.01",
          "0.1",
          "1",
          "10",
          "100"
        ),
        name = expression(
          "Stream Flow (m"^3*" s"^{-1}*")"
        )
      )
    ) +
    
    
    scale_x_datetime(
      name = NULL,
      date_breaks = "1 month",
      date_labels = "%b"
    ) +
    
    annotate(
      "text",
      x = -Inf,
      y = Inf,
      label = panel_label,
      hjust = -0.5,
      vjust = 1.5,
      fontface = "bold",
      size = 5
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid = element_blank()
    )
}
```

Plots
```{r}
SAN_plot <- make_levo_flow_plot(
  "SAN",
  point_fill = ws_cols["SAN"],
  point_shape = ws_shapes["SAN"],
  panel_label = "A)",
  show_x_axis = TRUE
)

PESC_plot <- make_levo_flow_plot(
  "PESC",
  point_fill = ws_cols["PESC"],
  point_shape = ws_shapes["PESC"],
  panel_label = "B)",
  show_x_axis = TRUE
)

MAJ_plot <- make_levo_flow_plot(
  "MAJ",
  point_fill = ws_cols["MAJ"],
  point_shape = ws_shapes["MAJ"],
  panel_label = "C)",
  show_x_axis = TRUE
)

LAG_plot <- make_levo_flow_plot(
  "LAG",
  point_fill = ws_cols["LAG"],
  point_shape = ws_shapes["LAG"],
  panel_label = "D)",
  show_x_axis = TRUE
)

SCT_plot <- make_levo_flow_plot(
  "SCT",
  point_fill = ws_cols["SCT"],
  point_shape = ws_shapes["SCT"],
  panel_label = "E)",
  show_x_axis = TRUE
)
```

Final Plot
```{r}
#combine plots
final_plot <- plot_grid(
  SAN_plot,  LAG_plot,
  PESC_plot, SCT_plot,
  MAJ_plot,  NULL,
  ncol = 2,
  align = "hv",
  axis = "tblr"
)

#Add Date label for x-axis
final_plot <- plot_grid(
  final_plot,
  ggdraw() + draw_label("Date", size = 12),
  ncol = 1,
  rel_heights = c(1, 0.05)
)
```

Save Final Plot
```{r}
ggsave(
  filename = "figures/Fig2_Levo_StreamFlow.tiff",
  plot = final_plot,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 10,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```

Playing with legend
```{r}
legend_df <- data.frame(
  Watershed = c(
    "San Lorenzo River (SAN)",
    "Pescadero Creek (PESC)",
    "Majors Creek (MAJ)",
    "Laguna Creek (LAG)",
    "Scott Creek (SCT)"
  ),
  x = 1:5,
  y = 1
)

legend_cols <- c(
  "San Lorenzo River (SAN)" = "#FBF9A2",
  "Pescadero Creek (PESC)" = "#FFCF55",
  "Majors Creek (MAJ)" = "#FF7410",
  "Laguna Creek (LAG)" = "#EA4412",
  "Scott Creek (SCT)" = "#A00000"
)

legend_shapes <- c(
  "San Lorenzo River (SAN)" = 23,
  "Pescadero Creek (PESC)" = 24,
  "Majors Creek (MAJ)" = 22,
  "Laguna Creek (LAG)" = 21,
  "Scott Creek (SCT)" = 25
)

legend_plot <- ggplot(
  legend_df,
  aes(
    x = x,
    y = y,
    fill = Watershed,
    shape = Watershed
  )
) +

  # watershed symbols
  geom_point(
    size = 5,
    color = "black",
    stroke = 0.8
  ) +

  # discharge line example
  annotate(
    "segment",
    x = 0.4,
    xend = 0.8,
    y = 0.5,
    yend = 0.5,
    color = "#156082",
    linewidth = 1.2
  ) +

  annotate(
    "text",
    x = 1.2,
    y = 0.5,
    label = "Discharge",
    hjust = 0,
    size = 4
  ) +

  scale_fill_manual(
    values = legend_cols,
    name = NULL
  ) +

  scale_shape_manual(
    values = legend_shapes,
    name = NULL
  ) +

  theme_void() +

  theme(
    legend.position = "right"
  )

#extract legend
legend <- cowplot::get_legend(legend_plot)

#final plot with legend
final_with_legend <- ggdraw(final_plot) +

  draw_grob(
    legend,
    x = 0.55,   # horizontal position
    y = 0.05,   # vertical position
    width = 0.35,
    height = 0.3
  )

ggsave(
  filename = "figures/Fig2_wLegend.tiff",
  plot = final_with_legend,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 10,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```

Fig 2 with legend
```{r}
legend_df <- data.frame(
  Watershed = c(
    "San Lorenzo River (SAN)",
    "Pescadero Creek (PESC)",
    "Majors Creek (MAJ)",
    "Laguna Creek (LAG)",
    "Scott Creek (SCT)"
  )
)

# force legend order to match color scheme
legend_df$Watershed <- factor(
  legend_df$Watershed,
  levels = c(
    "San Lorenzo River (SAN)",
    "Pescadero Creek (PESC)",
    "Majors Creek (MAJ)",
    "Laguna Creek (LAG)",
    "Scott Creek (SCT)"
  )
)

legend_cols <- c(
  "San Lorenzo River (SAN)" = "#FBF9A2",
  "Pescadero Creek (PESC)" = "#FFCF55",
  "Majors Creek (MAJ)" = "#FF7410",
  "Laguna Creek (LAG)" = "#EA4412",
  "Scott Creek (SCT)" = "#A00000"
)

legend_shapes <- c(
  "San Lorenzo River (SAN)" = 23,
  "Pescadero Creek (PESC)" = 24,
  "Majors Creek (MAJ)" = 22,
  "Laguna Creek (LAG)" = 21,
  "Scott Creek (SCT)" = 25
)


# create legend plot
watershed_legend_plot <- ggplot(
  legend_df,
  aes(
    x = 1,
    y = Watershed,
    fill = Watershed,
    shape = Watershed
  )
) +
  geom_point(
    size = 4,
    color = "black",
    stroke = 0.8
  ) +
  scale_fill_manual(
    values = legend_cols,
    name = "Watershed"
  ) +
  scale_shape_manual(
    values = legend_shapes,
    name = "Watershed"
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10)
  )


# extract legend
watershed_legend <- get_legend(watershed_legend_plot)

# Add watershed legend as inset
final_with_legend <- ggdraw() +
  
  # original figure
  draw_plot(final_plot) +
  
  # movable watershed legend
  draw_plot(
    watershed_legend,
    x = 0.60,   # move left/right
    y = 0.1,   # move up/down
    width = 0.30,
    height = 0.25
  )


# display
final_with_legend

#save
ggsave(
  filename = "figures/Fig2_wLegend.tiff",
  plot = final_with_legend,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 8,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```


Trouble shooting laguna
```{r}

```






















