---
title: "CoastalFires_FigureCreation"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
#load necessary libraries
```{r}
library(ggplot2)
library(grid)
library(extrafont)
library(tidyr)
library(dplyr)
library(lubridate)
library(scales)
```

#set working directory
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires")
```

#load dataset
```{r}
CombinedData <- readr::read_csv("CoastalFires_CombinedData.csv")
DischargeData <- readr::read_csv("CoastalFires_DischargeData.csv")
```
Print Column names
```{r}
col_names <- colnames(DischargeData)

# Print column names
print(col_names)
```
Make a new column in Combined data for Date_Time
```{r}
# Combine date and time into a new column with the desired format
CombinedData$datetime <- format(
  as.POSIXct(
    paste(CombinedData$date, CombinedData$time),        # Combine date and time
    format = "%m/%d/%Y %H:%M"           # Input format
  ),
  format = "%m/%d/%Y %H:%M:%S"          # Output format
)
```

Standardize DateTime formats for each watershed
```{r}
DischargeData$SAN_DateTime <- mdy_hm(DischargeData$SAN_DateTime)
DischargeData$PESC_DateTime <- mdy_hm(DischargeData$PESC_DateTime)
DischargeData$LAG_DateTime <- mdy_hm(DischargeData$LAG_DateTime)
DischargeData$MAJ_DateTime <- mdy_hm(DischargeData$MAJ_DateTime)
DischargeData$SCT_DateTime <- mdy(DischargeData$SCT_DateTime)  # Only date, no time
#some of the Scott Creek DateTime values will fail to parse. This is because the second water year was recorded using different DateTime formats. However, as this manuscript is only examining the first water year (Oct 2020-May 2021), this loss of discharge values do not impact analyses. Figures will be made only utilizing points prior to May 4th, 2021. 
```
Ensure all DateTime columns are in "POSIXct" rather than character classes
```{r}
CombinedData$datetime <- as.POSIXct(CombinedData$datetime, format = "%m/%d/%Y %H:%M")
DischargeData$LAG_DateTime <- as.POSIXct(DischargeData$LAG_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$MAJ_DateTime <- as.POSIXct(DischargeData$MAJ_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$PESC_DateTime <- as.POSIXct(DischargeData$PESC_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$SAN_DateTime <- as.POSIXct(DischargeData$SAN_DateTime, format = "%m/%d/%Y %H:%M")
DischargeData$SCT_DateTime <- as.POSIXct(DischargeData$SCT_DateTime, format = "%m/%d/%Y %H:%M")

class(CombinedData$datetime)
class(DischargeData$LAG_DateTime)
class(DischargeData$MAJ_DateTime)
class(DischargeData$PESC_DateTime)
class(DischargeData$SAN_DateTime)
class(DischargeData$SCT_DateTime)
```

Figure 2: The discharge (blue lines) and concentration of levoglucosan (yellow to dark red markers) from (A) the San Lorenzo River (B) Pescadero Creek (C) Majors Creek, (D) Laguna Creek, and (E) Scott Creek. The CZU Lightning Complex Fires (indicated by fire symbol) burned from August 16th until September 21st.
```{r}
# Create Fig. 2A - Pescadero Creek
library(scales)

# Rescale the levoglucosan data and discharge data to the same range
rescaled_discharge <- rescale(DischargeData$PESC_Flow_cms, to = c(min(CombinedData$levo_micro_gram_per_liter, na.rm = TRUE), 
                                                                  max(CombinedData$levo_micro_gram_per_liter, na.rm = TRUE)))

# Plot
ggplot() +
  # Levoglucosan data (primary y-axis)
  geom_point(data = CombinedData %>% filter(watershed == "Pescadero Creek"), 
             aes(x = datetime, y = levo_micro_gram_per_liter), 
             color = "blue", alpha = 0.7, size = 2) +
  geom_line(data = CombinedData %>% filter(watershed == "Pescadero Creek"), 
            aes(x = datetime, y = levo_micro_gram_per_liter), 
            color = "blue", linewidth = 1) +
  # Discharge data (secondary y-axis with rescaled values and log transformation)
  geom_line(data = DischargeData, 
            aes(x = PESC_DateTime, y = log10(rescaled_discharge + 1)),  # Apply log10 with +1 to avoid log(0)
            color = "red", linewidth = 1) +
  # Secondary axis for discharge (log scale)
  scale_y_continuous(
    name = "Levoglucosan (ppb)", 
    sec.axis = sec_axis(
      trans = ~10^(.) - 1,  # Reverse log scaling on the secondary axis
      name = "Discharge (m³/s, log scale)"
    )
  ) +
  # Common x-axis
  scale_x_datetime(name = "Date") +
  # Customize appearance
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "blue"),
    axis.title.y.right = element_text(color = "red"),
    panel.grid.minor = element_blank()
  ) +
  labs(title = "Discharge (Log Scale) and Levoglucosan Concentration Over Time")

  
```

```{r}
ggplot(DischargeData, aes(x = PESC_DateTime, y = PESC_Flow_cms)) +
  geom_line(color = "red") +
  scale_x_datetime(name = "Date") +
  scale_y_continuous(name = "Discharge (m³/s)") +
  theme_minimal() +
  labs(title = "Discharge Over Time")
```
