---
title: "CoastalFires_Dataset_Wrangling"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
Before starting, find and download necessary files:

Levoglucosan data from ESS-DIVE:
  CoastalFires_LevoglucosanData.csv

Black carbon data from CUAHSI Hydroshare:
Barton, R., C. Richardson, E. Pae, M. Montalvo, M. Redmond, M. Zimmer, S. Wagner (2025). Hydrology, rather than wildfire burn extent, determines post-fire organic and black carbon export from mountain rivers in central coastal California [Dataset], HydroShare, https://www.hydroshare.org/resource/421c0226bb38460c8393d67fe0c4f802/
  Organic and Black carbon data: CoastalFires_BiogeochemData.csv
  Discharge data: CoastalFires_DischargeData.csv
  Watershed Metadata: CoastalFires_WatershedData.csv
  

Install Packages
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyverse")
```

Load necessary Libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
```

Set Input and Output Directory
```{r}
#Input_Files folder should contains the following files: CoastalFires_LevoglucosanData.csv, CoastalFires_BiogeochemData.csv, CoastalFires_DischargeData.csv, CoastalFires_WatershedData.csv
input_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/Input_Files" # Change this to match your data frame location. Make sure it ends in "/Input_Files"

#Output Files folder should contains the following files:
#From ESS-DIVE: CoastalFires_DataFrameWrangling.Rmd
output_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/"  # Change this to match your scripts location
```

Load datafiles: levoglucosan data, discharge data, watershed and burn severity data, and previously published organic carbon and black carbon data
```{r}
LevoData <- readr::read_csv(file.path(input_folder, "CoastalFires_LevoglucosanData.csv"))
BlackCarbonData <- readr::read_csv(file.path(input_folder, "CoastalFires_BiogeochemData.csv"))
WatershedData <- readr::read_csv(file.path(input_folder, "CoastalFires_WatershedData.csv"))
DischargeData <- readr::read_csv(file.path(input_folder, "CoastalFires_DischargeData.csv"))
``` 

Adjust the black carbon data frame to remove unnecessary information
```{r}
#adjust the published data frame
#remove excess columns and print. The columns that are being removed are associated with the particulate phase (i.e., particulate organic carbon and particulate black carbon). As levoglucosan concentrations were determined only for the dissolved phase, the following data analysis will also only focus on the dissolved phase. 
BlackCarbonData <- BlackCarbonData %>%
  select(-TSS_mg_per_L, -POC_mgC_per_L, -`PBC_mgC_per_L`, -PBC_TSS_percent, 
         -PBC_POC_percent, -pB6CA_nano_mol_per_L, -pB5CA_nano_mol_per_L, -pB6CA_B5CA)
```

Combine LevoData, WatershedData, and BlackCarbonData Data Frames to one centralized dataframe: CoastalFires_CombinedData
```{r}
#combine data frames
Data <- left_join(LevoData, BlackCarbonData, by = "Sample_ID")
#exclude samples taken at lagoon dowstream of Scott sampling site
Data <- subset(Data, Data$Sample_ID != "SCT-7")
Data <- subset(Data, Data$Sample_ID != "SCT-8")
#add watershed data to table and remove duplicate locations
Data <- left_join(Data, WatershedData, by = "Watershed")
CoastalFires_CombinedData <- Data %>% select(-Lat, -Long)
```

#Calculate levoclucosan ratios
```{r}
#calculate the ratio of levoglucosan to bulk DOC and DBC
CoastalFires_CombinedData$Levo_DOC_percent <- Data$levo_micro_gram_per_liter/1000/Data$DOC_mgC_per_L*100
CoastalFires_CombinedData$Levo_DBC <- Data$levo_micro_gram_per_liter/1000/Data$DBC_mgC_per_L
```

#Calculate fluxes and yields
```{r}
#discharge normalized levoglucosan concentrations
CoastalFires_CombinedData$Q_norm_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_micro_gram_per_liter / CoastalFires_CombinedData$Discharge_cms)

#flux
CoastalFires_CombinedData$flux_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$levo_micro_gram_per_liter * CoastalFires_CombinedData$Discharge_cms)
CoastalFires_CombinedData$flux_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DOC_mgC_per_L * CoastalFires_CombinedData$Discharge_cms)
CoastalFires_CombinedData$flux_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$DBC_mgC_per_L * CoastalFires_CombinedData$Discharge_cms)

#yield
CoastalFires_CombinedData$yield_levo <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_levo / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DOC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DOC / CoastalFires_CombinedData$Area_square_kilometer)
CoastalFires_CombinedData$yield_DBC <- with(CoastalFires_CombinedData, CoastalFires_CombinedData$flux_DBC / CoastalFires_CombinedData$Area_square_kilometer)
```
#Calculating a "Time Since Fire" metric
```{r}
#Ensuring that the Date column is correctly classified as a date rather than a character
CoastalFires_CombinedData$Date <- as.Date(CoastalFires_CombinedData$Date, format = "%d-%b-%y")

#Calculating the time since fire for all samples and ensuring it is a numeric value
fire_date <- as.Date("2020-08-16")
CoastalFires_CombinedData$time_since_fire_day <- as.numeric(CoastalFires_CombinedData$Date - fire_date)
```

Re-arranged columns and change column names
```{r}
# Specify the desired column order
desired_order <- c("Sample_ID", "IGSN", "Watershed", "Date", "Time","time_since_fire_day", "Latitude", "Longitude", "Location_Change",
                   "Discharge_cms", "Flow_category", "DOC_mgC_per_L", 
                   "levo_micro_gram_per_liter", "std_levo_micro_gram_per_liter", 
                   "DBC_mgC_per_L", "DBC_DOC_percent", "dB6CA_nano_mol_per_L", "dB5CA_nano_mol_per_L", 
                   "dB6CA_B5CA", "Levo_DOC_percent", "Levo_DBC", "Q_norm_levo", "flux_levo", 
                   "flux_DOC", "flux_DBC", "yield_levo", "yield_DOC", 
                   "yield_DBC", "a254_per_m", "SUVA254_L_per_mgC_per_m", 
                   "Area_square_kilometer", "Average_slope_percent", 
                   "Average_precipitation_milli_meter", "Burn_Extent_percent", 
                   "Area_low_severity_percent", "Area_low_to_mod_severity_percent", 
                   "Area_mod_to_high_severity_percent", "Forest_percent", 
                   "Shrub_percent", "Developed_percent", 
                   "Sandstone_and_sedimentary_percent", "Igneous_percent", 
                   "Suficial_Sediments_percent", "Schist_percent")

# Rearrange the columns
CoastalFires_CombinedData <- CoastalFires_CombinedData[, desired_order]
```

Create a csv file of the combined dataset:CoastalFires_CombinedData
```{r}
write.csv(CoastalFires_CombinedData, 
          file = file.path(output_folder, "CoastalFires_CombinedData.csv"), 
          row.names = FALSE)  #Prevents writing row numbers
```