---
title: "Figure 2"
author: "rbb"
date: "2026-07-09"
output: html_document
---

Install necessary packages (run once if necessary)
```{r}
# install.packages("dplyr")
# install.packages("lubridate")
# install.packages("tidyr")
# install.packages("cowplot")
# install.packages("ggplot2")
```

Packages
```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(patchwork)
library(cowplot)
library(grid)
```

Get Data
```{r}
#From CUAHSI Hydroshare
DischargeData <- readr::read_csv("https://www.hydroshare.org/resource/26f08ef7fe344fcb9d3da45160b697fe/data/contents/CoastalFires_DischargeData.csv")

#From GitHub
CombinedData <- readr::read_csv("data/CoastalFires_CombinedData.csv")
```

Exclude samples that were taken at a different location. 
```{r}
#Scott Creek flooded during the events of January 27-28th, 2021. Samples were collected at a downstream lagoon site. However, due to this the data collected does not accurately reflect that at the sample site. As such authors of the associated manuscript have excluded it from all data analysis and figure creation.
CombinedData <- CombinedData %>%
  filter(Location_Change != "Y")
```

Make sure dates are right
```{r}
CombinedData <- CombinedData %>%
  mutate(
    DateTime = as.POSIXct(Date, tz = "America/Los_Angeles") +
      if_else(is.na(Time),
              12*60*60,
              as.numeric(Time)),
    DateTime = format(DateTime, "%Y-%m-%d %H:%M")
  )
```

Get mean error of replicates for error bars
```{r}
mean_levo_sd <- CombinedData %>%
  filter(!is.na(stdev_Levo_micro_mol_C_per_L)) %>%
  summarize(
    mean_sd = mean(stdev_Levo_micro_mol_C_per_L)
  ) %>%
  pull(mean_sd)
mean_levo_sd
```

Simplify Dataframe
```{r}
ConcData <- CombinedData %>%
  select(Watershed, DateTime, Levo_micro_mol_C_per_L, stdev_Levo_micro_mol_C_per_L, B6CA_and_B5CA_micro_mol_C_per_L)

#Fix watershed names to match discharge data
ConcData <- ConcData %>%
  mutate(Watershed = recode(Watershed,
                            "Laguna Creek" = "LAG",
                            "San Lorenzo River" = "SAN",
                            "Pescadero Creek" = "PESC",
                            "Majors Creek" = "MAJ",
                            "Scott Creek" = "SCT"))
#Make sure date is correct format
ConcData <- ConcData %>%
  mutate(
    DateTime = as.POSIXct(DateTime,
                          format = "%Y-%m-%d %H:%M",
                          tz = "UTC")
  )
```

Alter discharge dataframe for easier plotting
```{r}
SAN <- DischargeData %>%
  transmute(
    Watershed = "SAN",
    DateTime = parse_date_time(SAN_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = SAN_Flow_cms
  )

PESC <- DischargeData %>%
  transmute(
    Watershed = "PESC",
    DateTime = parse_date_time(PESC_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = PESC_Flow_cms
  )

LAG <- DischargeData %>%
  transmute(
    Watershed = "LAG",
    DateTime = parse_date_time(LAG_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = LAG_Flow_cms
  )

MAJ <- DischargeData %>%
  transmute(
    Watershed = "MAJ",
    DateTime = parse_date_time(MAJ_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = MAJ_Flow_cms
  )

SCT <- DischargeData %>%
  transmute(
    Watershed = "SCT",
    DateTime = parse_date_time(SCT_DateTime,
                               orders = c("mdy HM", "mdy"),
                               tz = "UTC"),
    Flow = SCT_Flow_cms
  )

FlowData <- bind_rows(SAN, PESC, LAG, MAJ, SCT)

#Remove NAs from FlowData
FlowData <- FlowData %>%
  filter(!is.na(DateTime),
         !is.na(Flow))

#Cut data to prefered date range
FlowData <- FlowData %>%
  filter(as.Date(DateTime) >= as.Date("2020-10-01"),
         as.Date(DateTime) <= as.Date("2021-05-30"))

#check it worked
FlowData %>%
  group_by(Watershed) %>%
  summarize(
    missing_dates = sum(is.na(DateTime)),
    start = min(DateTime),
    end = max(DateTime),
    n=n()
  )
```

Scaling factor
```{r}
# ============================================================
# Prepare flow for plotting
# ============================================================

FlowData <- FlowData %>%
  mutate(
    log_flow = log10(Flow)
  )


# ============================================================
# Common axis scaling for ALL watersheds
# ============================================================

levo_axis_max <- 2.5

# Force common discharge axis:
# 0.01 to 100 m3/s
flow_axis_min <- log10(0.01)
flow_axis_max <- log10(100)

# Convert discharge range into levoglucosan axis range
scale_factor <- levo_axis_max /
  (flow_axis_max - flow_axis_min)

scale_factor
```
Color scheme
```{r}
ws_cols <- c(
  SAN  = "#FBF9A2",
  PESC = "#FFCF55",
  MAJ  = "#FF7410",
  LAG  = "#EA4412",
  SCT  = "#A00000"
)

ws_shapes <- c(
  SAN  = 23,
  PESC = 24,
  MAJ  = 22,
  LAG  = 21,
  SCT  = 25
)
```


Function
```{r}
make_levo_flow_plot <- function(watershed_name,
                                point_fill,
                                point_shape,
                                panel_label,
                                show_x_axis = TRUE) {
  
  levo_plot <- ConcData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.POSIXct(DateTime),
      Levo_error = mean_levo_sd
    )
  
  
  flow_plot <- FlowData %>%
    filter(
      Watershed == watershed_name
    )
  
  
  ggplot() +
    
    geom_line(
      data = flow_plot,
      aes(
        x = DateTime,
        y = (log_flow - flow_axis_min) * scale_factor
      ),
      color = "#156082",
      linewidth = 1.2
    ) +
    
    
    geom_errorbar(
      data = levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4
    ) +
    
    geom_line(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      linetype = "dotted",
      linewidth = 0.5,
      color = "black"
    ) +
    
    geom_point(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = point_shape,
      size = 4,
      fill = point_fill,
      color = "black",
      stroke = 0.8
    ) +
    
    
    scale_y_continuous(
      limits = c(0, levo_axis_max),
      
      name = expression(
        "Levoglucosan ("*mu*"M-C)"
      ),
      
      sec.axis = sec_axis(
        ~ . / scale_factor + flow_axis_min,
        breaks = log10(c(0.01, 0.1, 1, 10, 100)),
        labels = c(
          "0.01",
          "0.1",
          "1",
          "10",
          "100"
        ),
        name = expression(
          "Stream Flow (m"^3*" s"^{-1}*")"
        )
      )
    ) +
    
    
    scale_x_datetime(
      name = NULL
    ) +
    
    annotate(
      "text",
      x = -Inf,
      y = Inf,
      label = panel_label,
      hjust = -0.5,
      vjust = 1.5,
      fontface = "bold",
      size = 5
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid = element_blank()
    )
}
```

Plots
```{r}
SAN_plot <- make_levo_flow_plot(
  "SAN",
  point_fill = ws_cols["SAN"],
  point_shape = ws_shapes["SAN"],
  panel_label = "A)",
  show_x_axis = TRUE
)

PESC_plot <- make_levo_flow_plot(
  "PESC",
  point_fill = ws_cols["PESC"],
  point_shape = ws_shapes["PESC"],
  panel_label = "B)",
  show_x_axis = TRUE
)

MAJ_plot <- make_levo_flow_plot(
  "MAJ",
  point_fill = ws_cols["MAJ"],
  point_shape = ws_shapes["MAJ"],
  panel_label = "C)",
  show_x_axis = TRUE
)

LAG_plot <- make_levo_flow_plot(
  "LAG",
  point_fill = ws_cols["LAG"],
  point_shape = ws_shapes["LAG"],
  panel_label = "D)",
  show_x_axis = TRUE
)

SCT_plot <- make_levo_flow_plot(
  "SCT",
  point_fill = ws_cols["SCT"],
  point_shape = ws_shapes["SCT"],
  panel_label = "E)",
  show_x_axis = TRUE
)
```

Final Plot
```{r}
#combine plots
final_plot <- plot_grid(
  SAN_plot,  LAG_plot,
  PESC_plot, SCT_plot,
  MAJ_plot,  NULL,
  ncol = 2,
  align = "hv",
  axis = "tblr"
)

#Add Date label for x-axis
final_plot <- plot_grid(
  final_plot,
  ggdraw() + draw_label("Date", size = 12),
  ncol = 1,
  rel_heights = c(1, 0.05)
)
```

Save Final Plot
```{r}
ggsave(
  filename = "figures/Fig2_Levo_StreamFlow.tiff",
  plot = final_plot,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 10,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```

Playing with legend
```{r}
legend_df <- data.frame(
  Watershed = c(
    "San Lorenzo River (SAN)",
    "Pescadero Creek (PESC)",
    "Majors Creek (MAJ)",
    "Laguna Creek (LAG)",
    "Scott Creek (SCT)"
  ),
  x = 1:5,
  y = 1
)

legend_cols <- c(
  "San Lorenzo River (SAN)" = "#FBF9A2",
  "Pescadero Creek (PESC)" = "#FFCF55",
  "Majors Creek (MAJ)" = "#FF7410",
  "Laguna Creek (LAG)" = "#EA4412",
  "Scott Creek (SCT)" = "#A00000"
)

legend_shapes <- c(
  "San Lorenzo River (SAN)" = 23,
  "Pescadero Creek (PESC)" = 24,
  "Majors Creek (MAJ)" = 22,
  "Laguna Creek (LAG)" = 21,
  "Scott Creek (SCT)" = 25
)

legend_plot <- ggplot(
  legend_df,
  aes(
    x = x,
    y = y,
    fill = Watershed,
    shape = Watershed
  )
) +

  # watershed symbols
  geom_point(
    size = 5,
    color = "black",
    stroke = 0.8
  ) +

  # discharge line example
  annotate(
    "segment",
    x = 0.4,
    xend = 0.8,
    y = 0.5,
    yend = 0.5,
    color = "#156082",
    linewidth = 1.2
  ) +

  annotate(
    "text",
    x = 1.2,
    y = 0.5,
    label = "Discharge",
    hjust = 0,
    size = 4
  ) +

  scale_fill_manual(
    values = legend_cols,
    name = NULL
  ) +

  scale_shape_manual(
    values = legend_shapes,
    name = NULL
  ) +

  theme_void() +

  theme(
    legend.position = "right"
  )

#extract legend
legend <- cowplot::get_legend(legend_plot)

#final plot with legend
final_with_legend <- ggdraw(final_plot) +

  draw_grob(
    legend,
    x = 0.55,   # horizontal position
    y = 0.05,   # vertical position
    width = 0.35,
    height = 0.3
  )

ggsave(
  filename = "figures/Fig2_wLegend.tiff",
  plot = final_with_legend,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 10,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```

Fig 3 with legend
```{r}
legend_df <- data.frame(
  Watershed = c(
    "San Lorenzo River (SAN)",
    "Pescadero Creek (PESC)",
    "Majors Creek (MAJ)",
    "Laguna Creek (LAG)",
    "Scott Creek (SCT)"
  )
)

# force legend order to match color scheme
legend_df$Watershed <- factor(
  legend_df$Watershed,
  levels = c(
    "San Lorenzo River (SAN)",
    "Pescadero Creek (PESC)",
    "Majors Creek (MAJ)",
    "Laguna Creek (LAG)",
    "Scott Creek (SCT)"
  )
)

legend_cols <- c(
  "San Lorenzo River (SAN)" = "#FBF9A2",
  "Pescadero Creek (PESC)" = "#FFCF55",
  "Majors Creek (MAJ)" = "#FF7410",
  "Laguna Creek (LAG)" = "#EA4412",
  "Scott Creek (SCT)" = "#A00000"
)

legend_shapes <- c(
  "San Lorenzo River (SAN)" = 23,
  "Pescadero Creek (PESC)" = 24,
  "Majors Creek (MAJ)" = 22,
  "Laguna Creek (LAG)" = 21,
  "Scott Creek (SCT)" = 25
)


# create legend plot
watershed_legend_plot <- ggplot(
  legend_df,
  aes(
    x = 1,
    y = Watershed,
    fill = Watershed,
    shape = Watershed
  )
) +
  geom_point(
    size = 4,
    color = "black",
    stroke = 0.8
  ) +
  scale_fill_manual(
    values = legend_cols,
    name = "Watershed"
  ) +
  scale_shape_manual(
    values = legend_shapes,
    name = "Watershed"
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10)
  )


# extract legend
watershed_legend <- get_legend(watershed_legend_plot)

# Add watershed legend as inset
final_with_legend <- ggdraw() +
  
  # original figure
  draw_plot(final_plot) +
  
  # movable watershed legend
  draw_plot(
    watershed_legend,
    x = 0.60,   # move left/right
    y = 0.1,   # move up/down
    width = 0.30,
    height = 0.25
  )


# display
final_with_legend

#save
ggsave(
  filename = "figures/Fig2_wLegend.tiff",
  plot = final_with_legend,
  device = "tiff",
  dpi = 600,
  width = 10,
  height = 10,
  units = "in",
  compression = "lzw",
  bg = "white"
)
```






















Function
```{r}
make_levo_flow_plot <- function(watershed_name) {
  
  levo_plot <- ConcData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.POSIXct(DateTime),
      Levo_error = mean_levo_sd
    )
  
  
  flow_plot <- FlowData %>%
    filter(
      Watershed == watershed_name
    )
  
  
  ggplot() +
    
    geom_line(
      data = flow_plot,
      aes(
        x = DateTime,
        y = (log_flow - flow_axis_min) * scale_factor
      ),
      color = "#156082",
      linewidth = 1.2
    ) +
    
    
    geom_errorbar(
      data = levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4
    ) +
    
    
    geom_point(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = 23,
      size = 4,
      fill = "#FBF9A2",
      color = "black",
      stroke = 0.8
    ) +
    
    
    scale_y_continuous(
      limits = c(0, levo_axis_max),
      
      name = expression(
        "Levoglucosan ("*mu*"mol C L"^{-1}*")"
      ),
      
      sec.axis = sec_axis(
        
        # inverse transformation
        ~ . / scale_factor + flow_axis_min,
        
        # secondary-axis values
        breaks = log10(c(0.01, 0.1, 1, 10, 100)),
        
        labels = c(
          "0.01",
          "0.1",
          "1",
          "10",
          "100"
        ),
        
        name = expression(
          "Discharge (m"^3*" s"^{-1}*")"
        )
      )
    ) +
    
    
    scale_x_datetime(
      name = "Date"
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid.minor = element_blank()
    )
}
```


Five plots
```{r}
SAN_plot <- make_levo_flow_plot("SAN")

PESC_plot <- make_levo_flow_plot("PESC")

LAG_plot <- make_levo_flow_plot("LAG")

MAJ_plot <- make_levo_flow_plot("MAJ")

SCT_plot <- make_levo_flow_plot("SCT")
```


















































Prepare Flow Data and get a scaling factor 
```{r}
# ============================================================
# Prepare flow for plotting
# ============================================================

FlowData <- FlowData %>%
  mutate(
    log_flow = log10(Flow)
  )

# ============================================================
# Common axis scaling for ALL watersheds
# ============================================================

levo_axis_max <- 2.5

# Force common discharge axis:
# 0.01 to 100 m3/s
flow_axis_min <- log10(0.01)
flow_axis_max <- log10(100)

# Convert discharge range into levoglucosan axis range
scale_factor <- levo_axis_max /
  (flow_axis_max - flow_axis_min)

scale_factor
```
Create function for plotting
```{r}
make_levo_flow_plot <- function(watershed_name) {
  
  levo_plot <- ConcData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.POSIXct(DateTime),
      Levo_error = mean_levo_sd
    )
  
  
  flow_plot <- FlowData %>%
    filter(
      Watershed == watershed_name
    )
  
  
  ggplot() +
    
    geom_line(
      data = flow_plot,
      aes(
        x = DateTime,
        y = (log_flow - flow_axis_min) * scale_factor
      ),
      color = "#156082",
      linewidth = 1.2
    ) +
    
    
    geom_errorbar(
      data = levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4
    ) +
    
    
    geom_point(
      data = levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = 23,
      size = 4,
      fill = "#FBF9A2",
      color = "black",
      stroke = 0.8
    ) +
    
    
    scale_y_continuous(
      limits = c(0, levo_axis_max),
      
      name = expression(
        "Levoglucosan ("*mu*"mol C L"^{-1}*")"
      ),
      
      sec.axis = sec_axis(
        
        # inverse transformation
        ~ . / scale_factor + flow_axis_min,
        
        # secondary-axis values
        breaks = log10(c(0.01, 0.1, 1, 10, 100)),
        
        labels = c(
          "0.01",
          "0.1",
          "1",
          "10",
          "100"
        ),
        
        name = expression(
          "Discharge (m"^3*" s"^{-1}*")"
        )
      )
    ) +
    
    
    scale_x_datetime(
      name = "Date"
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid.minor = element_blank()
    )
}
```

Plot each watershed
```{r}

```


Create plot dataset
```{r}
# Calculate mean analytical standard deviation from replicate measurements
mean_levo_sd <- ConcData %>%
  filter(!is.na(stdev_Levo_micro_mol_C_per_L)) %>%
  summarize(
    mean_sd = mean(stdev_Levo_micro_mol_C_per_L, na.rm = TRUE)
  ) %>%
  pull(mean_sd)


# Concentration data for plotting
Levo_plot <- ConcData %>%
  filter(!is.na(Levo_micro_mol_C_per_L)) %>%
  mutate(
    Date = as.Date(DateTime),
    Levo_error = mean_levo_sd
  ) %>%
  filter(Watershed == "SAN")


# Discharge data for plotting
flow_plot <- FlowData %>%
  filter(Watershed == "SAN")

# Scaling factor for secondary axis
# Find scaling factor
log_flow_max <- max(FlowData$log_flow_shift, na.rm = TRUE)
levo_max <- levo_max <- 2.5 #max(Levo_plot$Levo_micro_mol_C_per_L, na.rm = TRUE)

scale_factor <- levo_max / log_flow_max

# Plot
ggplot() +

  # Hydrograph
  geom_line(
    data = flow_plot,
    aes(
      x = DateTime,
      y = log_flow_shift * scale_factor
    ),
    color = "#156082",
    linewidth = 1.2
  )+

  # Levoglucosan error bars
  geom_errorbar(
    data = Levo_plot,
    aes(
      x = Date,
      ymin = Levo_micro_mol_C_per_L - Levo_error,
      ymax = Levo_micro_mol_C_per_L + Levo_error
    ),
    width = 0,
    linewidth = 0.4,
    color = "black"
  ) +

  # Levoglucosan concentrations
  geom_point(
    data = Levo_plot,
    aes(
      x = DateTime,
      y = Levo_micro_mol_C_per_L
    ),
    shape = 23,
    size = 4,
    fill = "#FBF9A2",
    color = "black",
    stroke = 0.8
  ) +

  scale_y_continuous(
    limits = c(0, 2.5),
    name = expression(
      "Levoglucosan ("*mu*"mol C L"^{-1}*")"
    ),
sec.axis = sec_axis(
  ~ . / scale_factor + min_log,
  breaks = (log10(c(0.1, 1, 10, 50)) - min_log) * scale_factor,
  labels = c("0.1", "1", "10", "50"),
  name = expression(
    "Discharge (m"^3*" s"^{-1}*")"
  )
)
    ) +
  
  #Watershed label
  annotate(
   "text",
   x = as.Date("2020-10-01"),
   y = 2.5,
   label = "A. San Lorenzo River",
   hjust = 0,
   fontface = "bold"
  ) +

    scale_x_datetime(
    name = "Date"
  ) +
  
  theme_bw(base_size = 12) +

  theme(
    panel.grid.minor = element_blank()
  )
```
Trying to make a levoglucosan plot function
```{r}
make_levo_plot <- function(
  watershed_name,
  flow_name,
  panel_label,
  point_shape,
  point_fill,
  levo_max = 2.5
) {
  
  # Concentration data
  Levo_plot <- CombinedData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = as.Date(DateTime),
      Levo_error = mean_levo_sd
    )
  
# #Discharge data
 flow_plot <- DailyFlow %>%
   filter(Watershed == flow_name) %>%
   mutate(
     Q_log = log10(Q_mean)
   )
  
# Scaling
q_min <- floor(min(flow_plot$Q_log, na.rm = TRUE))
q_max <- max(flow_plot$Q_log, na.rm = TRUE)

scale_factor <- levo_max / (q_max - q_min)

flow_plot <- flow_plot %>%
  mutate(
    Q_plot = (Q_log - q_min) * scale_factor
  )

# Discharge tick marks
q_ticks <- c(0.01, 0.1, 1, 10)

q_ticks <- q_ticks[
  q_ticks >= min(flow_plot$Q_mean, na.rm = TRUE) &
  q_ticks <= max(flow_plot$Q_mean, na.rm = TRUE)
]
  
  # Plot
  ggplot() +
    
    # Hydrograph
    geom_line(
      data = flow_plot,
      aes(
        x = Date,
        y = Q_plot
        # y = (log10(Q_mean) - q_low) * scale_factor
      ),
      color = "#156082",
      linewidth = 1
    ) +
    
    # Levoglucosan connecting line
    geom_line(
      data = Levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      color = "grey30",
      linewidth = 0.45,
      linetype = "dashed",
      alpha = 0.7
    ) +
    # Error bars
    geom_errorbar(
      data = Levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4,
      color = "black"
    ) +
    
    # Points
    geom_point(
      data = Levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = point_shape,
      size = 3,
      fill = point_fill,
      color = "black",
      stroke = 0.8
    ) +
    
    # Axes
scale_y_continuous(
  limits = c(0, levo_max),
  name = expression(
    "Levoglucosan ("*mu*"mol C L"^{-1}*")"
  ),
  sec.axis = sec_axis(
    ~ (. / scale_factor + q_min) |> (\(x) 10^x)(),
    breaks = (log10(q_ticks) - q_min) * scale_factor,
    labels = q_ticks,
    name = expression(
      "Discharge (m"^3*" s"^{-1}*")"
    )
  )
      ) +
      # sec.axis = sec_axis(
      #   ~ (. / scale_factor) + q_low,
      #   breaks = log10(c(0.01, 0.1, 1, 10, 100)),
      #   labels = c("0.01", "0.1", "1", "10", "100"),
      #   name = expression(
      #     "Discharge (m"^3*" s"^{-1}*")"
      #   )
      # )

    
    # Label
    annotate(
      "text",
      x = as.Date("2020-10-01"),
      y = 2.4,
      label = panel_label,
      hjust = 0,
      fontface = "bold",
      size = 5
    ) +
    
    #Scale x-axis
    scale_x_date(
      limits = as.Date(c("2020-10-01", "2021-05-02")),
      date_breaks = "1 month",
      date_labels = "%b"
    ) +
    
    theme_bw(base_size = 12) +
    
    theme(
      panel.grid = element_blank()
    )
}
  
```

Plot with function
```{r}
plot_SAN <- make_levo_plot(
  "San Lorenzo River",
  "SAN",
  "A)",
  point_shape = 23,
  point_fill = "#FBF9A2"
)


plot_PESC <- make_levo_plot(
  "Pescadero Creek",
  "PESC",
  "B)",
  point_shape = 24,
  point_fill = "#FFCF55"
)


plot_MAJ <- make_levo_plot(
  "Majors Creek",
  "MAJ",
  "C)",
  point_shape = 22,
  point_fill = "#FF7410"
)


plot_LAG <- make_levo_plot(
  "Laguna Creek",
  "LAG",
  "D)",
  point_shape = 21,
  point_fill = "#EA4412"
) 


plot_SCT <- make_levo_plot(
  "Scott Creek",
  "SCT",
  "E)",
  point_shape = 25,
  point_fill = "#A00000"
) 

#To remove the X-axis label from some of the plots
plot_SAN <- plot_SAN +
  theme(axis.title.x = element_blank())

plot_PESC <- plot_PESC +
  theme(axis.title.x = element_blank())

plot_LAG <- plot_LAG +
  theme(axis.title.x = element_blank())
```

Trying to plot with patchwork
```{r}
final_fig <- (
  plot_SAN /
  plot_PESC /
  plot_MAJ
) |
(
  plot_LAG /
  plot_SCT /
  plot_spacer()
)

final_fig <- final_fig +
  plot_layout(
    widths = c(1,1)
  ) &
  theme(
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    plot.tag.position = c(0.5,1.1),
    plot.tag = element_text(
      face = "bold",
      size = 12
    )
  )
```

Save
```{r}
ggsave(
  "Figure2_Levo_Hydrographs.png",
  final_fig,
  width = 9,
  height = 9,
  units = "in",
  dpi = 600
)
```

Log-transform discharge data for plotting
```{r}
Daily_flow_log <-
```

Function for plotting levo & discharg
```{r}
make_levo_plot <- function(
  watershed_name,
  flow_name,
  panel_label,
  point_shape,
  point_fill,
  levo_max = 2.5
) {
  
  # -------------------------
  # Levoglucosan data
  # -------------------------
  Levo_plot <- CombinedData %>%
    filter(
      Watershed == watershed_name,
      !is.na(Levo_micro_mol_C_per_L)
    ) %>%
    mutate(
      Date = DateTime,
      Levo_error = mean_levo_sd
    )
  
  
# -------------------------
# Discharge data
# -------------------------
flow_plot <- FlowData %>%
  filter(Watershed == flow_name) %>%
  mutate(
    Date = DateTime
  )

# Actual discharge limits
q_min <- min(flow_plot$Flow, na.rm = TRUE)
q_max <- max(flow_plot$Flow, na.rm = TRUE)

# Scale discharge to levoglucosan axis
scale_factor <- levo_max / (q_max - q_min)

flow_plot <- flow_plot %>%
  mutate(
    Q_scaled = (Flow - q_min) * scale_factor
  )
  
  # -------------------------
  # Plot
  # -------------------------
  ggplot() +
    
    # Hydrograph
geom_line(
  data = flow_plot,
  aes(
    x = Date,
    y = Q_scaled
  ),
  color = "#156082",
  linewidth = 1
) +
    
    # Levoglucosan connecting line
    geom_line(
      data = Levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      color = "grey30",
      linewidth = 0.45,
      linetype = "dashed",
      alpha = 0.7
    ) +
    
    # Error bars
    geom_errorbar(
      data = Levo_plot,
      aes(
        x = Date,
        ymin = Levo_micro_mol_C_per_L - Levo_error,
        ymax = Levo_micro_mol_C_per_L + Levo_error
      ),
      width = 0,
      linewidth = 0.4,
      color = "black"
    ) +
    
    # Levoglucosan points
    geom_point(
      data = Levo_plot,
      aes(
        x = Date,
        y = Levo_micro_mol_C_per_L
      ),
      shape = point_shape,
      size = 3,
      fill = point_fill,
      color = "black",
      stroke = 0.8
    ) +
    
    # Axes
scale_y_continuous(
  limits = c(0, levo_max),
  name = expression(
    "Levoglucosan ("*mu*"mol C L"^{-1}*")"
  ),
  sec.axis = sec_axis(
    trans = ~ . / scale_factor + q_min,
    name = expression(
      "Discharge (m"^3*" s"^{-1}*")"
    )
  )
) +
    
    # Panel label
    annotate(
      "text",
      x = as.Date("2020-10-01"),
      y = 2.4,
      label = panel_label,
      hjust = 0,
      fontface = "bold",
      size = 5
    ) +
    
    # Date axis
    scale_x_datetime(
      limits = as.POSIXct(c("2020-10-01", "2021-05-02")),
      date_breaks = "1 month",
      date_labels = "%b"
    ) +
    
    theme_bw(base_size = 12) +
    theme(
      panel.grid = element_blank()
    )
}
```

Scratch plots
```{r}
plot_SAN <- make_levo_plot(
  "San Lorenzo River",
  "SAN",
  "A)",
  point_shape = 23,
  point_fill = "#FBF9A2"
)


plot_PESC <- make_levo_plot(
  "Pescadero Creek",
  "PESC",
  "B)",
  point_shape = 24,
  point_fill = "#FFCF55"
)


plot_MAJ <- make_levo_plot(
  "Majors Creek",
  "MAJ",
  "C)",
  point_shape = 22,
  point_fill = "#FF7410"
)


plot_LAG <- make_levo_plot(
  "Laguna Creek",
  "LAG",
  "D)",
  point_shape = 21,
  point_fill = "#EA4412"
)


plot_SCT <- make_levo_plot(
  "Scott Creek",
  "SCT",
  "E)",
  point_shape = 25,
  point_fill = "#A00000"
)

plot_SAN <- plot_SAN +
  theme(axis.title.x = element_blank())

plot_PESC <- plot_PESC +
  theme(axis.title.x = element_blank())

plot_LAG <- plot_LAG +
  theme(axis.title.x = element_blank())
```



