---
title: "CoastalFires_Dataset_Wrangling"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
Before starting, find and download necessary files:

Levoglucosan data from ESS-DIVE:
  CoastalFires_LevoglucosanData.csv

Black carbon data from CUAHSI Hydroshare:
Barton, R., C. Richardson, E. Pae, M. Montalvo, M. Redmond, M. Zimmer, S. Wagner (2025). Hydrology, rather than wildfire burn extent, determines post-fire organic and black carbon export from mountain rivers in central coastal California [Dataset], HydroShare, https://www.hydroshare.org/resource/421c0226bb38460c8393d67fe0c4f802/
  Organic and Black carbon data: CoastalFires_BiogeochemData.csv
  Discharge data: CoastalFires_DischargeData.csv
  Watershed Metadata: CoastalFires_WatershedData.csv
  

Install package if necessary
```{r}
install.packages("tidyverse")
```

Load library
```{r}
library(tidyverse)
```

Set directory for input files and working directory
```{r}
#Input_Files folder should contains the following files: CoastalFires_LevoglucosanData.csv, CoastalFires_BiogeochemData.csv, CoastalFires_DischargeData.csv, CoastalFires_WatershedData.csv
input_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/Input_Files" # Change this to match your directory. Make sure it ends in "/Input_Files"

#working folder should contain: ReadMe.txt, CoastalFires_DataFrameWrangling.Rmd, and CoastalFires_DataAnalysis.Rmd
working_folder <- "C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires/"  # Change this to match your directory.
```

Load datafiles from Input_Files folder: levoglucosan data, discharge data, watershed and burn severity data, and previously published organic carbon and black carbon data
```{r}
#From ESS-DIVE
LevoData <- readr::read_csv(file.path(input_folder, "CoastalFires_LevoglucosanData.csv"))

#From CUAHSI Hydroshare
BlackCarbonData <- readr::read_csv(file.path(input_folder, "CoastalFires_BiogeochemData.csv"))
WatershedData <- readr::read_csv(file.path(input_folder, "CoastalFires_WatershedData.csv"))
DischargeData <- readr::read_csv(file.path(input_folder, "CoastalFires_DischargeData.csv"))
``` 

Combine LevoData, WatershedData, and BlackCarbonData data frames to one centralized dataframe: CombinedData
```{r}
#combine data frames
CombinedData <- left_join(LevoData, BlackCarbonData, by = "Sample_ID")

#add watershed metadata and remove duplicated columns
CombinedData <- left_join(CombinedData, WatershedData, by = "Watershed")
CombinedData <- CombinedData %>% select(-Lat, -Long) #delete extra location columns
```

Adjust the black carbon data frame to remove unnecessary columns
```{r}
#Adjust the published data frame to make it more applicable to current study. 

#Remove unnecessary columns. The columns that are being removed are associated with the particulate phase (i.e., particulate organic carbon and particulate black carbon). As levoglucosan concentrations were determined only for the dissolved phase, the following data analysis will also only focus on the dissolved phase. 
BlackCarbonData <- BlackCarbonData %>%
  select(-TSS_mg_per_L, -POC_mgC_per_L, -`PBC_mgC_per_L`, -PBC_TSS_percent, 
         -PBC_POC_percent, -pB6CA_nano_mol_per_L, -pB5CA_nano_mol_per_L, -pB6CA_B5CA)
```

Re-arranged columns and change column names
```{r}
# Specify the desired column order
desired_order <- c("Sample_ID", "Watershed", "Date", "Time", "Latitude", "Longitude", "Location_Change",
                   "Discharge_cms", "Flow_category", "DOC_mgC_per_L", 
                   "levo_micro_gram_per_liter", "std_levo_micro_gram_per_liter", 
                   "DBC_mgC_per_L", "DBC_DOC_percent", "dB6CA_nano_mol_per_L", "dB5CA_nano_mol_per_L", 
                   "dB6CA_B5CA", "a254_per_m", "SUVA254_L_per_mgC_per_m", 
                   "Area_square_kilometer", "Average_slope_percent", 
                   "Average_precipitation_milli_meter", "Burn_Extent_percent", 
                   "Area_low_severity_percent", "Area_low_to_mod_severity_percent", 
                   "Area_mod_to_high_severity_percent", "Forest_percent", 
                   "Shrub_percent", "Developed_percent", 
                   "Sandstone_and_sedimentary_percent", "Igneous_percent", 
                   "Suficial_Sediments_percent", "Schist_percent")

# Rearrange the columns
CombinedData <- CombinedData[, desired_order]
```

Create a csv file of the combined dataset:CoastalFires_CombinedData
```{r}
write.csv(CombinedData, 
          file = file.path(working_folder, "CoastalFires_CombinedData.csv"), 
          row.names = FALSE)  #Prevents writing row numbers
```