---
title: "Figure1"
output: html_document
---

add info here

Install package if necessary
```{r}
install.packages("arcgislayers")
install.packages("sf")
install.packages("tmap")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("terra")
```

load package(s)
```{r}
library(arcgislayers)
library(sf)
library(tmap)
library(ggplot2)
library(dplyr)
library(terra)
```

find and download shapefiles/rasters
```{r}
#California
  #download as zip file
    temp <- tempdir() #get temp directory to download zip to 
    url <- "https://data.ca.gov/dataset/e212e397-1277-4df3-8c22-40721b095f33/resource/3db1e426-fb51-44f5-82d5-a54d7c6e188b/download/ca_state.zip" #location of zip file 
    download.file(url, file.path(temp, "CA.zip"), mode="wb")
    #unzip file
    unzip(zipfile=file.path(temp, "CA.zip"), exdir=temp)
    #load data 
    ca <- st_read(file.path(temp,"CA_State.shp"))
    
#Burn severity from MTBS, downloaded on May-26-2026. Filters - Region: California, Year: 2020
  burn_severity <- rast("gis_data/mtbs_2020/mtbs_CONUS_2020.tif")
  
#Sites
  sites <- 
```

test to visualize the fires in CA
```{r}
ca <- st_transform(ca, crs(burn_severity)) #ensure same projection
#crop
fires_ca <- crop(burn_severity, vect(ca))
fires_ca <- mask(fires_ca, vect(ca))

#QUICK test map
plot(fires_ca)
plot(st_geometry(ca), add = TRUE, border = "black", lwd = 1)
```



Read spatial layers
```{r}
ca <- st_read("data/shapefiles/california.shp")
burn_severity <- rast("data/fire/burn_severity.tif")
fire_perimeter <- st_read("data/fire/fire_perimeter.shp")
watersheds <- st_read("data/watersheds/watersheds.shp")
sites <- st_read("data/sites/sampling_locations.shp")
```

```{r}
workflow

# VECTOR
watersheds <- st_read(...)
sites <- st_read(...)
fire_perim <- st_read(...)

# RASTER
burn <- rast(...)

# PROJECT
target_crs <- "EPSG:3310"

watersheds <- st_transform(watersheds, target_crs)
sites <- st_transform(sites, target_crs)
fire_perim <- st_transform(fire_perim, target_crs)

burn <- project(burn, target_crs)
```

Reproject layers
```{r}
target_crs <- 3310   # California Albers

ca <- st_transform(ca, target_crs)
burn_severity <- project(burn_severity, "EPSG:3310")
fire_perimeter <- st_transform(fire_perimeter, target_crs)
watersheds <- st_transform(watersheds, target_crs)
sites <- st_transform(sites, target_crs)
```

Crop to study area
```{r}
study_extent <- st_bbox(watersheds)
```

Visualize as a map
```{r}
tmap_mode("plot")

map1 <- 
  tm_shape(ca_crop) +
  tm_polygons(col = "gray95", border.col = "gray70") +

  tm_shape(burn_crop) +
  tm_raster(
    palette = "-YlOrRd",
    title = "Burn Severity",
    alpha = 0.8
  ) +

  tm_shape(fire_perimeter) +
  tm_borders(col = "black", lwd = 2) +

  tm_shape(watersheds) +
  tm_borders(col = "dodgerblue4", lwd = 1.5) +

  tm_shape(sites) +
  tm_symbols(
    col = "cyan",
    size = 0.12,
    border.col = "black"
  ) +

  tm_layout(
    frame = FALSE,
    legend.outside = TRUE,
    legend.text.size = 0.8,
    legend.title.size = 1,
    fontfamily = "sans"
  ) +

  tm_compass(
    type = "arrow",
    position = c("left", "top")
  ) +

  tm_scale_bar(
    position = c("left", "bottom")
  )

map1
```

Export high quality figure
```{r}
tmap_save(
  map1,
  "figures/study_site_map.png",
  width = 8,
  height = 6,
  units = "in",
  dpi = 600
)
```













    
    
#step 0: load libraries------ 
library(sbtools)

#step 1: get wildfire data ------ 
  #go to MTBS (https://burnseverity.cr.usgs.gov/viewer/?product=MTBS) to download historic fire data 
  #select by watershed and select McKenzie in Oregon

  #!!!! change to make name of file
  #mtbs_loc <- file.path(fs::path_home(), "Downloads/", "Fire_data_bundles_cXE2AL4en0OeTkOTbc7c.zip")

  #clean wildfire data 
    clean_mtbs("data/input/MTBS-wildfires-mckenzie-basin.zip", save_loc="data/spatial-data/mtbs")
    
  #make a shapefile with all fire boundries
    burn_bd <- list.files("data/spatial-data/mtbs", recursive = TRUE, full.names = TRUE, pattern ="burn_bndy.shp")
    comb <- lapply(burn_bd, read_sf)
    comb <- do.call(rbind, comb)
    write_sf(comb, "data/spatial-data/mtbs/all-fires-burn-bndy.shp")
    
```

Install package if necessary
```{r}
#install.packages("tidyverse")
```
