---
title: "hekp"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document
---
#Load Necessary Libraries
```{r}
library("vegan")
library(ggplot2)
library(dplyr)
```

#Set Working Directory and load in data frame
```{r}
setwd("C:/Users/barto/OneDrive/Documents/GitHub/CoastalFires")
FigureData <- readr::read_csv("CoastalFires_FigureData.csv")
#Normalize discharge to watershed area
FigureData$Q_norm <- FigureData$Q_cms / FigureData$Area_square_kilometer
```

#NMDS for PyC yields
```{r}
# Prepare data frames for NMDS
analyte_data <- FigureData[, c("avg_levo_ppb", "DOC_mgC_per_L", "DBC_mgC_per_L", "DBC_DOC_perc", "dB6CA_B5CA", "Levo_DBC", "Levo_DOC_perc")] #choosing only the yield variables from the dataset

#Scale analyte_data to be compatible with a distance matrix
analyte_data <- scale(analyte_data) # Standardize to z-scores
analyte_data <- analyte_data - min(analyte_data) # Shift to make all values positive

#Calculate distance matrix to prepare data for NMDS
dist_matrix <- vegdist(analyte_data, method = "bray") #compute Bray-Curtis distance matrix

#Run NMDS
nmds_result <- metaMDS(dist_matrix, k = 2, trymax = 100)
print(nmds_result)
```
#Environmental Fitting
```{r}
# Prepare data for environmental fitting
env_data <- FigureData[, c("Q_norm", "Area_square_kilometer", "Burn_Extent_percent", "Area_low_severity_percent", "time_since_fire_day")]  # choosing environmental variables that will later be showcased as vectors: "Average_slope_percent", "Area_low_to_mod_severity_percent", "Area_mod_to_high_severity_percent"

# Run envfit
envfit_result <- envfit(nmds_result, env_data, permutations = 999)  # Fit environmental variables to the NMDS ordination
print(envfit_result)  # Check the result to ensure it's valid

# Extract vector data from envfit
envfit_vectors <- as.data.frame(scores(envfit_result, display = "vectors"))
envfit_vectors$variable <- rownames(envfit_vectors)  # Add variable names

# Extract p-values and add to the data frame
envfit_vectors$pval <- envfit_result$vectors$pvals  # Extract p-values for vectors

# Categorize vectors based on p-value significance
envfit_vectors$significance <- ifelse(envfit_vectors$pval < 0.05, "significant", "not_significant")
```
#Extract analyte points for plotting
```{r}
analyte_points <- envfit(nmds_result, analyte_data, permutations = 999)$vectors
analyte_points_df <- as.data.frame(analyte_points$arrows)
analyte_points_df$Analyte <- rownames(analyte_points$arrows)
colnames(analyte_points_df)[1:2] <- c("NMDS1", "NMDS2")
```

#Extract NMDS results and rename analytes and vectors for plotting
```{r}
# Extract NMDS scores for plotting
nmds_scores <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_scores$Site <- FigureData$time_since_fire_day  # Add Site info for coloring

# Rename variables in envfit_vectors
envfit_vectors$variable <- recode(envfit_vectors$variable,
                                  Q_norm = "Q",
                                  time_since_fire_day = "Time Post Fire",
                                  Burn_Extent_percent = "BE%",
                                  Area_low_severity_percent = "Low BE%",
                                  Area_square_kilometer = "Area",
                                  ) # Adjust names as needed

# Rename analytes in analyte_points_df
analyte_points_df$Analyte <- recode(analyte_points_df$Analyte,
                                  avg_levo_ppb = "Levo",
                                  DOC_mgC_per_L = "DOC",
                                  DBC_mgC_per_L = "DBC",
                                  Levo_DBC = "Levo:DBC",
                                  Levo_DOC_perc = "Levo:DOC",
                                  DBC_DOC_perc = "DBC:DOC",
                                  dB6CA_B5CA = "B6CA:B5CA") # Replace with desired names "DBC_DOC_perc", "dB6CA_B5CA", "Levo_DBC", "Levo_DOC_perc"
```

#Plot NMDS
```{r}
# Multiply arrow lengths
envfit_vectors_long <- envfit_vectors %>%
  mutate(NMDS1 = NMDS1 * 1.5,  # Multiply xend by 1.5
         NMDS2 = NMDS2 * 1.5)  # Multiply yend by 1.5

ggplot() +
  # NMDS points, colored by the continuous variable `Site`
  geom_point(data = nmds_scores, aes(x = NMDS1, y = NMDS2, color = Site), size = 3) +
  
  # Vectors for envfit results
  geom_segment(data = envfit_vectors_long,
               aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2, linetype = significance),
               arrow = arrow(length = unit(0.3, "cm")), size = 1, color = "black") +
  
  # Text labels for vectors
  geom_text(data = envfit_vectors_long,
            aes(x = NMDS1, y = NMDS2, label = variable),
            vjust = -0.5, hjust = 1, color = "black") +
  
  # Additional points for analyte data
  geom_point(data = analyte_points_df, aes(x = NMDS1, y = NMDS2), 
             shape = 21, fill = "hotpink", size = 5) +
  
  # Text labels for analyte points
  geom_text(data = analyte_points_df, aes(x = NMDS1, y = NMDS2, label = Analyte),
            vjust = -0.5, color = "black", size = 4) +
  
  # Continuous color scale for `Site`
  scale_color_gradient(low = "#00FFFF", high = "navy", name = "Time Since Fire", 
                       limits = c(50, 300)) +
  
  # Discrete linetype scale for `significance`
  scale_linetype_manual(values = c("significant" = "solid", "not_significant" = "dashed"), 
                        name = "Vector Significance") +
  
  labs(title = "NMDS of Analyte Yields with Environmental Vectors and Points",
       x = "NMDS1",
       y = "NMDS2") +
  
  # Adjust the x-axis limits
  scale_x_continuous(limits = c(-1.1, NA)) +  # Set minimum x to -1.1, keep upper limit auto
  
  # Customize theme for gridlines and axis appearance
  theme_minimal() +
  theme(
    panel.grid = element_blank(),                 # Remove gridlines
    axis.line = element_line(color = "black"),   # Add black axis lines
    axis.ticks = element_line(color = "black"),  # Add black axis ticks
    axis.ticks.length = unit(0.2, "cm")          # Set tick marks outside
  )




```

ggplot() +
  # NMDS points, colored by the continuous variable `Site`
  geom_point(data = nmds_scores, aes(x = NMDS1, y = NMDS2, color = Site), size = 3) +
  
  # Vectors for envfit results
  geom_segment(data = envfit_vectors,
               aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2, linetype = significance),
               arrow = arrow(length = unit(0.3, "cm")), color = "black") +
  
  # Text labels for vectors
  geom_text(data = envfit_vectors,
            aes(x = NMDS1, y = NMDS2, label = variable),
            vjust = -0.5, hjust = 1, color = "black") +
  
  # Continuous color scale for `Site`
  scale_color_gradient(low = "#00FFFF", high = "navy", name = "Time Since Fire") +
  
  # Discrete linetype scale for `significance`
  scale_linetype_manual(values = c("significant" = "solid", "not_significant" = "dashed"), 
                        name = "Vector Significance") +
  
  labs(title = "NMDS of Analyte Yields with Environmental Vectors",
       x = "NMDS1",
       y = "NMDS2") +
  
  theme_minimal()

```


#Plot PyC Yield NMDS results
#```{r}
# Extract NMDS scores for plotting
nmds_scores <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_scores$Site <- FigureData$time_since_fire_day  # Add Site info for coloring

analyte_points <- envfit(nmds_result, analyte_data, permutations = 999)$vectors
analyte_points_df <- as.data.frame(analyte_points$arrows)
analyte_points_df$Analyte <- rownames(analyte_points$arrows)
colnames(analyte_points_df)[1:2] <- c("NMDS1", "NMDS2")

# Extract vector data from envfit
envfit_vectors <- as.data.frame(scores(envfit_result, display = "vectors"))
envfit_vectors$variable <- rownames(envfit_vectors)  # Add variable names

# Extract p-values and add to the data frame
envfit_vectors$pval <- envfit_result$vectors$pvals  # Extract p-values for vectors


# Categorize vectors based on p-value significance
envfit_vectors$significance <- ifelse(envfit_vectors$pval < 0.05, "significant", "not_significant")

ggplot() +
  # NMDS points, colored by the continuous variable `Site`
  geom_point(data = nmds_scores, aes(x = NMDS1, y = NMDS2, color = Site), size = 3) +
  
  # Vectors for envfit results
  geom_segment(data = envfit_vectors,
               aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2, linetype = significance),
               arrow = arrow(length = unit(0.3, "cm")), color = "black") +
  
  # Text labels for vectors
  geom_text(data = envfit_vectors,
            aes(x = NMDS1, y = NMDS2, label = variable),
            vjust = -0.5, hjust = 1, color = "black") +
  
  # Additional points for analyte data
  geom_point(data = analyte_points_df, aes(x = NMDS1, y = NMDS2), 
             shape = 21, fill = "red", size = 3) +
  
  # Text labels for analyte points
  geom_text(data = analyte_points_df, aes(x = NMDS1, y = NMDS2, label = Analyte),
            vjust = -0.5, color = "red", size = 3) +
  
  # Continuous color scale for `Site`
  scale_color_gradient(low = "#00FFFF", high = "navy", name = "Time Since Fire") +
  
  # Discrete linetype scale for `significance`
  scale_linetype_manual(values = c("significant" = "solid", "not_significant" = "dashed"), 
                        name = "Vector Significance") +
  
  labs(title = "NMDS of Analyte Yields with Environmental Vectors and Points",
       x = "NMDS1",
       y = "NMDS2") +
  
  theme_minimal()

```


