---
title: "TestingNMDS"
author: "rbb"
date: "`r Sys.Date()`"
output: html_document

```{r}
# Load necessary libraries
library(vegan)   # for NMDS and distance matrix calculation
library(ggplot2) # for plotting
library(ggrepel) # for labeling with ggplot
```

```{r}
# Use iris dataset as a sample dataset
data(iris)
head(iris)
```

# Prepare data frame for NMDS
# Remove the species column since we are not focusing on it for NMDS
iris_data <- iris[, 1:4]

# Standardize the data
iris_data_scaled <- scale(iris_data)

# Calculate the Bray-Curtis distance matrix
dist_matrix <- vegdist(iris_data_scaled, method = "bray")

# Run NMDS
nmds_result <- metaMDS(dist_matrix, k = 2, trymax = 100)

# Extract NMDS coordinates for the samples (MDS1 and MDS2)
nmds_points <- as.data.frame(scores(nmds_result, display = "sites"))

# Extract the NMDS loadings (vectors for the variables)
loadings <- scores(nmds_result, display = "species")

# Create the NMDS plot with ggplot
ggplot(nmds_points, aes(x = MDS1, y = MDS2)) +
  geom_point(aes(color = iris$Species), size = 3) + # Plot points colored by species
  geom_segment(data = loadings, aes(x = 0, y = 0, xend = MDS1, yend = MDS2), 
               arrow = arrow(type = "closed", length = unit(0.2, "inches")), 
               color = "blue") + # Add vectors
  geom_text_repel(data = loadings, aes(x = MDS1, y = MDS2, label = rownames(loadings)), 
                  size = 4, color = "red", box.padding = 0.5) + # Add labels for vectors
  labs(title = "NMDS with Vectors and Sample Points", 
       x = "NMDS Dimension 1", 
       y = "NMDS Dimension 2") +
  theme_minimal() + 
  theme(legend.position = "bottom")
```
